---
title: "Basics - Quick Start Guide"
description: "2. Basics - Quick Start Guide"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Basics - Quick Start Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

```{r developernote, eval=FALSE, echo= FALSE, include = FALSE}
#  *>>>>>>>>>> Developer note: vignettes need to be tested/edited/rebuilt regularly <<<<<<<<<<<*
#    - **See ?pkgdown::build_site** and script in EJAM/data-raw/- EJAM uses the pkgdown R package to build help and articles/ vignettes as web pages
```

```{r SETUP_default_eval_or_not, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(eval = FALSE)
# https://r-pkgs.org/vignettes.html
```

```{r libraryEJAM, eval = TRUE, echo= FALSE, include= FALSE, warning=FALSE, message=FALSE}
# rm(list = ls()); golem::detach_all_attached(); devtools::load_all()

if (!exists("blockgroupstats")) {library(EJAM)} # use installed version only if pkg not yet attached

dataload_dynamic('all') # varnames = all  currently means all defined by .arrow_ds_names

indexblocks()
```

# A Brief Intro to Using EJAM in RStudio

This document is about analysts or coders using the EJAM R package in RStudio.
After you [install the EJAM R package](installing.html), this
document explains how you can run an EJAM analysis and view results right away
using R.

This document is not about [using EJAM as a web
application](webapp.html), but you can launch a local web app
after installing the EJAM R package.

## Load EJAM

To start using EJAM in RStudio/R, you first attach/load the R package using
`library()` or `require()`.

```{r library, eval=FALSE}
library(EJAM)
```

## Analyze Places with `ejamit()`

To quickly try EJAM in RStudio:

```{r ejamit100, message=FALSE, warning=FALSE, echo=TRUE, eval=FALSE}
# EJAM analysis of 100 places, for everyone within 3 miles
out <- ejamit(testpoints_100, radius = 3)

pts <- sitepoints_from_any(c("30.97740, -83.36900", "32.51581, -86.37732"))
out2 <- ejamit(pts, radius = 2)
```

To quickly try EJAM with an example input file (spreadsheet with latitude and
longitude of each point)

```{r, eval=FALSE}
myfile <- system.file("testdata/latlon/testpoints_10.xlsx", package = "EJAM")

out <- ejamit(myfile, radius = 3)
```

If you already have your own spreadsheet of point locations to analyze, then in
RStudio you can just use the `ejamit()` function without specifying the
locations or radius -- EJAM will prompt you to select the file and a radius.

```{r notevaluated, eval=FALSE}
out <- ejamit()
```

Note: The file should be an Excel file or .csv file and the first sheet (tab)
needs to be just a table of one header row (with at least two columns named lat
and lon), and one row per site (point). No extra rows, no merged cells, etc.

If you need examples of spreadsheets (and other input files you can try), you
can find the ones installed with the EJAM package in your local folder, like
this in the RStudio console:

```{r systemfile, eval=FALSE, echo=TRUE}
## See where the folder is and see what files are there:
testdata()

# or see just the latlon files:
testdata("latlon", quiet=T)
# dir(system.file("testdata/latlon", package = "EJAM"))
```

## Pick a Radius

You can specify the radius in miles. EJAM will analyze all residents within that
many miles of each point (site).

```{r RADIUS, eval = TRUE}
radius <- 3 # radius (in miles).  5 km = 3.106856 miles, 10 km = 6.2 miles
```

Converting between miles and kilometers -- If you know you want to analyze for 5
kilometers, for example, you can turn it into miles.

```{r convert_units, eval=TRUE}
5000 / meters_per_mile
convert_units(5, 'km', 'miles')
```

## Map your sites before analyzing them

This creates an interactive map. Click a point on the map to see a popup with
details about that point.

```{r eval = FALSE}
# input to EJAM
pts <- testpoints_100
mapfast(pts)
```

## Map results with `ejam2map()`

This also creates an interactive map. Click a point on the map to see a popup
with details about people near that point.

```{r map100, eval = TRUE}
out <- testoutput_ejamit_100pts_1miles 
ejam2map(out, launch_browser = FALSE) 
```

## Report via `ejam2report()`

`ejam2report()` turns the analysis results into a multisite report or community
report that can be viewed in a browser or saved as an HTML file (with
interactive map linking to each site's report) or saved as a PDF (better page
breaks for printing).

```{r ejam2report1_show_commands, eval=FALSE, echo = TRUE}

out <- testoutput_ejamit_100pts_1miles

# Create a multisite report for all sites:
ejam2report(out)

# Save the multisite report as a PDF:
ejam2report(out, fileextension = "pdf")

# Save the multisite report as an interactive HTML file:
ejam2report(out, fileextension = "html")

# Create a community report for one site:
ejam2report(out, sitenumber = 1, analysis_title = "Site #1")
```

See complete examples:

- [Community Report for Charles County, Maryland (single
  site)](https://api.ejanalysis.com/report?fips=24029)

- [Interactive HTML Multisite Report for a 10-site analysis of point locations,
  using a one-mile radius](https://github.com/Public-Environmental-Data-Partners/EJAM/blob/main/inst/testdata/examples_of_output/testoutput_ejam2report_10pts_1miles.html)
- [Multisite Report for 100 point locations, using a one-mile
  radius](https://github.com/Public-Environmental-Data-Partners/EJAM/blob/main/inst/testdata/examples_of_output/testoutput_ejam2report_100pts_1miles.pdf)

More Multisite Report PDF examples cover [selected
counties](https://github.com/Public-Environmental-Data-Partners/EJAM/blob/main/inst/testdata/examples_of_output/testoutput_ejam2report_fips_counties.pdf)
and [uploaded
shapes](https://github.com/Public-Environmental-Data-Partners/EJAM/blob/main/inst/testdata/examples_of_output/testoutput_ejam2report_shapes_2.pdf).


## Table of Results in RStudio console

As an alternative to the report provided by `ejam2report()`, this gives you a
quick, simple list of results for all the indicators:

```{r ejam2table_tall, eval=FALSE}
ejam2table_tall(out)
ejam2table_tall(out, sitenumber = 1)
```

## Barplot

```{r ejam2barplot, eval=FALSE, echo=TRUE}
out <- testoutput_ejamit_100pts_1miles

# Check long list of indicators for any that are elevated

ejam2barplot(out,
             varnames = names_these_ratio_to_avg,
             main = "Envt & Demog Indicators at Selected Sites Compared to US Averages")

ejam2barplot(out,
             varnames = names_these_ratio_to_state_avg,
             main = "Envt & Demog Indicators at Selected Sites Compared to State Averages")


# residential population data only

# vs nationwide avg
ejam2barplot(testoutput_ejamit_100pts_1miles)

# vs statewide avg
ejam2barplot(testoutput_ejamit_1000pts_1miles,
             varnames = c(names_d_ratio_to_state_avg, names_d_subgroups_ratio_to_state_avg),
             main = "Residential population group percentages at Selected Sites Compared to State Averages")


# Environmental only

ejam2barplot(testoutput_ejamit_100pts_1miles,
             varnames = c(names_e_ratio_to_avg),  # , names_e_ratio_to_state_avg),
             main = "Environmental Indicators at Selected Sites Compared to Averages")
```

```{r}
# see more examples at ?ejam2barplot
```

To see the underlying US or State average values these barplots compare against, use `usastats_means()` / `statestats_means()` (or `usastats_query()` / `statestats_query()` for percentiles too) -- see the "Using EJAM for Analysis in R" article.

## Special Areas and Features (`ejam2areafeatures()`)

EJAM also summarizes what percent of the residents analyzed have certain features
in their blockgroup (a school, hospital, or place of worship) or live in a
blockgroup that overlaps certain special types of areas (Tribal areas,
nonattainment areas, impaired waters, CEJST or IRA disadvantaged communities,
housing burden communities, food deserts, transportation disadvantaged
communities), compared to the share of all US residents -- and the share of all
residents of the state(s) analyzed -- who have those features or overlaps.
This info also appears in the Report, and in the "Area Features" tab
of the spreadsheet from `ejam2excel()`.

```{r ejam2areafeatures, eval=FALSE, echo=TRUE}
out <- testoutput_ejamit_100pts_1miles

# table of the stats, incl. ratios to the US and State percentages
ejam2areafeatures(out)

# barplot of ratios to the US percentages
ejam2barplot_areafeatures(out)

# barplot of ratios to the State percentages
# (for multistate analyses, the baseline is the average among all the residents
#  at these sites, using the statewide percentage in each resident's state)
ejam2barplot_areafeatures(out, vs = "state")
```

## View Results Spreadsheet via `ejam2excel()` (to Launch Excel)

```{r table_xls_from_ejam-launchexcel, eval=FALSE}
out <- testoutput_ejamit_100pts_1miles
ejam2excel(out, launchexcel = T, save_now = F)
```

### Save Results as a Spreadsheet file

```{r table_xls_from_ejam-savenow, eval=FALSE}
ejam2excel(out, save_now = T)
```

<!--  REMOVE FROM HERE UNTIL FIXED FOR RMARKDOWN VIGNETTE 

![Excel output showing 1 row per site and 1 column per indicator such as total
population nearby](images/image-xlsx.png){width="100%"}

REMOVED FROM HERE -->

--------------------------------------------------------------------------------

## More about points

### Use one point

```{r one point, eval=FALSE}
pts <- data.frame(lon = -92.380556, lat = 31.316944)
```

### Use a few points

```{r ptstypedin, eval = TRUE}
pts <- sitepoints_from_any(c(
  "34.8799123, -92.1",
  "30.2906971, -91.8",
  "30,         -95"
))
## or
pts  <- data.frame(
  lon = c(-92.1,      -91.8), 
  lat = c(34.8799123, 30.2906971)
)

pts
```

### Create a random sample of points representative of the average facility, average resident, or average area

You can create a set of random points with function `testpoints_n()` that can be
weighted to represent the average resident, average regulated facility, average
point on a map weighted by square meters, etc. See more details in the
documentation of the function `testpoints_n()`.

Create random test data points in States of LA and TX

```{r mapfast_testpoints_n, eval = TRUE, fig.height=6, fig.width=6}
# p1k <- testpoints_n(1000)
# mapfast(p1k)

mapfast(testpoints_n(300, ST = c('LA','TX'), weighting = 'bg'), radius = 0.1) 
# weighting = "frs" better represents regulated facilities,
# but requires loading the (large) frs dataset
```

# Documentation of Functions and Data

-   [README](`r paste0(EJAM::url_package("docs"), "/index.html")`){.uri target="_blank" rel="noreferrer noopener"}
-   [Function Reference Document](`r paste0(EJAM::url_package("docs"), "/reference/index.html")`){.uri target="_blank"
rel="noreferrer noopener"}
-   In RStudio, see `?EJAM`

```{r helpEJAM, eval=FALSE, include=TRUE}
?EJAM
# or 
help("EJAM", package='EJAM')

?ejamit()
```
