---
title: "Defaults and Custom Settings for the Web App"
description: "How EJAM app defaults, launch parameters, bookmarks, and Advanced Settings differ"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Defaults and Custom Settings for the Web App}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 100
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# What this article covers

EJAM has several ways to customize the Shiny app. They act at different times and are not
interchangeable:

| Mechanism | Best for | Scope |
|---|---|---|
| Arguments to `ejamapp()` | Configuring a local app or a deployment | Fixed for that app process |
| App launch parameters such as `?fips=` | Opening an already-running app with sites or a radius | One visitor's session |
| A Shiny bookmark | Restoring Shiny input controls | One saved session state |
| The Advanced Settings tab | Changing the controls exposed there | Current session, after launch |
| EJAM API request parameters | Requesting a report or data without the Shiny app | One API request |

This distinction is important. For example, adding an arbitrary `?name=value` to the app URL does
not turn it into an `ejamapp(name = value)` argument. The app recognizes only the launch parameters
listed below. Likewise, an `ejamapp()` argument is not bookmarkable unless it becomes a Shiny
`input$` control.

For development-branch documentation, use:

- [this article on the development site](https://docs.ejanalysis.com/dev/articles/dev-app-settings.html);
- [the development `ejamapp()` reference](https://docs.ejanalysis.com/dev/reference/ejamapp.html);
- [Accessing the Web App](webapp.html) for the user-facing overview; and
- [Using the EJAM API](dev-api.html) for API endpoints and parameters.

# Configure an app with `ejamapp()`

Run `ejamapp()` from R to launch a local app. The root `app.R` used for the hosted public app calls
`ejamapp(isPublic = TRUE)`. A plain local `ejamapp()` call defaults to the fuller analyst
configuration because `isPublic` is not set to `TRUE`.

The authoritative parameter examples are in `?ejamapp`. Keep calls small and use the documented
convenience names:

```{r ejamapp-examples}
## A local app preloaded with two counties and a 3-mile radius
EJAM::ejamapp(
  fips = c("10001", "10003"),
  radius = 3,
  default_show_advanced_settings = TRUE
)

## Start with a detailed NAICS selection
EJAM::ejamapp(
  analysis_title = "Custom NAICS analysis",
  naics = "562211"
)
```

Use `radius`, not the internal setting name `radius_default`, in new `ejamapp()` calls.
`ejamapp()` normalizes `radius` (and its synonym `buffer`) to that internal setting.

## Supported convenience arguments

`ejamapp()` accepts `...`, then normalizes these documented aliases before collecting the app
defaults:

| Argument supplied | Internal setting or data |
|---|---|
| `radius` or `buffer` | `radius_default` |
| `pts`, or `lat` plus `lon` | `sitepoints` |
| `shp` or `shape` | `shapefile` |
| `fips` | FIPS upload data |
| `naics`, `sic`, or `mact` | The corresponding `default_*` selection |
| `analysis_title` | `default_standard_analysis_title` |
| `report_title` | `report_title_multisite` |
| `default_upload_dropdown` | Backward-compatible alias for `default_site_method` |

Supplying sites also selects the matching site-selection method. For example, `fips=` selects the
FIPS upload path, `shapefile=` selects the polygon upload path, and `naics=` selects the dropdown
NAICS path.

Because `...` can accept any name, a misspelled or unused name is not guaranteed to produce an
error. Use names documented by `?ejamapp`, or an exact setting name that the UI or server reads
through `global_or_param()`.

## How deploy-time defaults are collected

At launch, `ejamapp()`:

1. normalizes its convenience arguments;
2. calls `get_global_defaults_or_user_options()`;
3. keeps user-supplied values ahead of same-named defaults from
   `global_defaults_package.R`, `global_defaults_shiny.R`, and
   `global_defaults_shiny_public.R`;
4. stores the resulting list as golem options; and
5. creates the app with `app_ui`, `app_server`, and the requested bookmarking mode.

`global_defaults_shiny_public.R` uses `isPublic` to choose the simpler public configuration. Most
code reads the collected value with `global_or_param("setting_name")`.

Package maintainers may change a default in the appropriate `inst/global_defaults_*.R` source file.
For a particular local run or deployment, prefer an `ejamapp()` argument instead of editing an
installed package file.

# Launch an already-running EJAM app

An app launch URL can preload sites for one visitor without changing the deployed app's defaults.
Use `url_ejamapp()` to build these links. It uses the query-preserving app alias recorded in
`DESCRIPTION`, rather than hardcoding a deployment hostname.

```{r app-launch-links, eval = TRUE}
EJAM::url_ejamapp(
  lat = c(33, 34),
  lon = c(-112, -114),
  radius = 3
)

EJAM::url_ejamapp(fips = c("10001", "10003"), buffer = 0)
```

The app recognizes:

- `?lat=33,34&lon=-112,-114` for points;
- `?fips=10001,10003` for FIPS codes;
- `?shape=<URL-encoded inline GeoJSON>` for a polygon or FeatureCollection;
- `?radius=3` or `?buffer=3` for the analysis radius in miles;
- `?handoff=<token>` for a site collection stored temporarily by the EJAM API; and
- `?advanced=TRUE` (or `?show_advanced_settings=TRUE`) to show the Advanced Settings tab.

`advanced` accepts `1`, `true`, `t`, `yes`, or `y`; the corresponding false values hide the tab.
An unrecognized value is ignored. `url_ejamapp()` builds the site, radius, and handoff parameters.
Append the Advanced Settings parameter when needed:

```{r advanced-launch-link, eval = TRUE}
paste0(EJAM::url_ejamapp(), "?advanced=TRUE")

paste0(
  EJAM::url_ejamapp(fips = c("10001", "10003")),
  "&advanced=TRUE"
)
```

## Launch-parameter rules and limits

- A handoff token is handled instead of direct site parameters.
- Otherwise the app loads one place type, in this order: valid `lat`/`lon` points, then FIPS,
  then an inline GeoJSON shape.
- Point latitude and longitude vectors must have equal lengths and contain no missing values.
- `shape=` must contain inline GeoJSON. A URL or local file path is deliberately rejected.
- A single radius applies to all supplied sites. Radius zero is valid for FIPS and polygon
  analyses; a point radius is clamped to the point slider's allowed range.
- Launch-provided sites and radius take precedence over corresponding `ejamapp()`/deployment
  defaults at startup. The user can still change controls afterward.

Large polygon collections can exceed normal URL limits. EJScreen's "Send to EJAM" workflow sends
those places to the EJAM API `POST /handoff` endpoint, receives a short-lived token, and opens the
app with `?handoff=<token>`. The app retrieves the collection from `GET /handoff/<token>`.
See [Using the EJAM API](dev-api.html) and the
[EJAM-API README](https://github.com/Public-Environmental-Data-Partners/EJAM-API#readme).

# Shiny bookmarks

Bookmarks restore Shiny `input$` values. They are separate from the compact launch parameters
above.

`ejamapp(enableBookmarking = "url")` is the default. When bookmarking is enabled, use the app's
**Bookmark** button to generate the URL; do not construct `?_inputs_&...` URLs by hand. Input IDs
and encoded formats can change.

- URL bookmarking stores input values in the URL but cannot carry uploaded files.
- Server bookmarking stores state on the server and can preserve uploads, subject to the Shiny
  server's configuration and retention.
- `enableBookmarking = "disable"` disables the feature.

A bookmark captures inputs, not every app option. Values such as a deployment logo or an option
that never becomes an input are not restored by a bookmark. Use the compact launch parameters for
points, FIPS, polygons, handoff tokens, and radius rather than trying to encode uploaded site data
as bookmark inputs.

# Advanced Settings

The Advanced Settings tab exposes a subset of app settings as Shiny inputs. It does not expose
every `global_defaults_*` value or every `ejamapp()` argument.

- For a local or purpose-built deployment, use
  `ejamapp(default_show_advanced_settings = TRUE)`.
- To let users reveal or hide it from the About tab, use
  `ejamapp(default_can_show_advanced_settings = TRUE)`.
- For one visitor to a running app, use `?advanced=TRUE`. This also enables the tab in a public
  deployment where it is normally unavailable.

The UI is built before reactive `input$` values exist. A normal static control therefore gets its
initial `selected=` or `value=` from `global_or_param()`, never from `input$...`. When an Advanced
Settings control changes a main control later, `app_server.R` applies it with an
`update*Input()` function. Use `renderUI()` only when the widget's structure itself depends on
reactive state.

Advanced Settings includes experimental and specialist controls. Changing a cap in the tab cannot
exceed an absolute `maxmax_*` limit set by the deployment.

# Links from EJAM to EJScreen

`url_ejscreenmap()` builds the reverse kind of deep link: it opens the EJScreen map with a place
drawn or selected. This is not an EJAM app-default mechanism, so its full parameter vocabulary is
maintained in `?url_ejscreenmap` and the EJScreen repository rather than duplicated here.

```{r ejscreen-links, eval = TRUE}
## One EJScreen URL containing both points
EJAM::url_ejscreenmap(
  lat = c(39, 39.7),
  lon = c(-75.5, -75.6),
  radius = 3,
  combined = TRUE
)
```

The helper supports points, supported Census FIPS types, small polygon outlines, explicit ZIP
codes, and `wherestr` place searches. The destination EJScreen deployment must include the deep-link
support added in
[Public-Environmental-Data-Partners/EJScreen#70](https://github.com/Public-Environmental-Data-Partners/EJScreen/pull/70);
older deployments ignore the newer selection parameters.

# API parameters are a separate interface

`ejamapi()` and `url_ejamapi()` request reports or data from the REST API. They are not alternate
ways to set arbitrary Shiny app controls. The API and app deep links share core site names such as
`lat`, `lon`, `fips`, `shape`, `radius`, and `buffer`, but each interface has additional parameters
and different return behavior.

See [Using the EJAM API](dev-api.html), `?ejamapi`, `?url_ejamapi`, and the
[EJAM-API README](https://github.com/Public-Environmental-Data-Partners/EJAM-API#readme) for the
current endpoint-specific contract.

# Developer maintenance checks

Do not paste a snapshot of every global default or `input$` name into this article. Those lists
become obsolete whenever the app changes. Inspect the source of truth instead:

```{r inspect-settings}
## Consolidated defaults from the installed package
defaults <- EJAM:::get_global_defaults_or_user_options()
sort(names(defaults))

## Approximate input$ references in the current server source
EJAM:::input_names_listing("R/app_server.R")
```

When changing this system, check these files together:

- `R/ejamapp.R` for convenience arguments and launch-time normalization;
- `R/utils_get_global_defaults_or_user_options.R` and `R/utils_global_or_param.R`;
- `inst/global_defaults_package.R`, `inst/global_defaults_shiny.R`, and
  `inst/global_defaults_shiny_public.R`;
- `R/app_ui.R` for static controls and the Advanced Settings tab;
- `R/app_server.R` for launch-URL parsing, reactive controls, and `update*Input()` synchronization;
- `R/url_ejamapp.R` for app-link generation; and
- `tests/testthat/test-webapp-ui_and_server.R` and
  `tests/testthat/test-URL_FUNCTIONS_part1.R` for the corresponding contracts.

If a new launch parameter is added, update the parser, URL builder where applicable, tests,
`?ejamapp`/`?url_ejamapp`, this article, and the shorter cross-references in
[Accessing the Web App](webapp.html) and [Using the EJAM API](dev-api.html).
