---
title: "Updating EJScreen Datasets Annually (via the Pipeline)"
description: "Annual staged workflow for updating EJScreen/EJAM demographic, environmental, and related datasets"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Updating EJScreen Datasets Annually (via the Pipeline)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

```{r, eval=TRUE, echo=FALSE, include=FALSE}

code_reponame_url        <- EJAM::url_package(type = "code", get_full_url = TRUE)
code_reponame_with_owner <- EJAM::url_package(type = "code", get_full_url = FALSE)
code_reponame <- gsub(".*/", "", code_reponame_with_owner)

data_reponame_url        <- EJAM::url_package(type = "data", get_full_url = TRUE)
data_reponame_with_owner <- EJAM::url_package(type = "data", get_full_url = FALSE)
data_reponame <- gsub(".*/", "", data_reponame_with_owner)

docs_reponame_url        <- EJAM::url_package(type = "docs", get_full_url = TRUE)

```

This document describes the annual staged pipeline for updating the
blockgroup-level EJScreen/EJAM datasets, especially the objects historically
called `blockgroupstats`, `bgej`, `usastats`, and `statestats`.

For general dataset maintenance outside this pipeline, such as FRS-related
tables, NAICS/SIC tables, block-level files, Arrow releases, and package-data
publication mechanics, see [Updating and Managing the Datasets Used by
EJAM](dev-update-datasets.html).

## Scope

The pipeline is designed to make each major data step explicit, saved, and
rerunnable. The main entry point is `calc_ejscreen_dataset()`. The recommended
annual runner is:

```{r, eval=FALSE}
source("data-raw/run_ejscreen_pipeline_annual.R")
```

That wrapper builds a validated annual pipeline config and then delegates to the
package pipeline runner, `EJAM:::run_ejscreen_pipeline()`. The repository also
provides smaller runner wrappers for common rerun modes:

```{r, eval=FALSE}
# Recheck saved outputs and prior-version comparisons without rebuilding stages.
source("data-raw/run_ejscreen_pipeline_validation_only.R")

# Recreate EJScreen-facing export files from saved pipeline stages.
source("data-raw/run_ejscreen_pipeline_exports_only.R")

# Run a release-oriented recipe after outputs and settings have been reviewed.
source("data-raw/run_ejscreen_pipeline_release.R")
```

These wrappers build a validated pipeline config first and then call the same
underlying package runner. They are meant for review and release-maintenance
passes after the main annual run has already created the needed intermediate
stages. The long-standing compatibility script,
`data-raw/run_ejscreen_dataset_pipeline.R`, remains available for older
interactive workflows that set environment variables and then call `source()`.
The release wrapper also keeps package-data replacement opt-in; set
`EJAM_REPLACE_PACKAGE_DATA = "TRUE"` explicitly only when reviewed outputs
should replace package `.rda` objects.

The runner writes pipeline checkpoints, by default as CSV files. It can also
write secondary formats such as `.rda` files for the same table stages when
`EJAM_STAGE_FORMATS` includes those formats. It does not, by itself, replace
every installed package dataset in `data/*.rda`. Replacing package data objects
with EJAM metadata helpers is a separate release step after the pipeline
outputs have been reviewed. `bgej` is the main exception: the accepted annual
`bgej` table should be republished as `bgej.arrow` in the `ejamdata` release
tag recorded in DESCRIPTION as `ejamdata_required_tag`, rather than saved as
`data/bgej.rda`.

## Pipeline Stages

The annual workflow creates or reads these stages:

1. `bg_acs_raw`: raw ACS table-based summary file data downloaded from the
   Census Bureau. This is saved before EJAM renaming or formula calculations.
   The default raw ACS storage is a folder of per-table files, with a manifest.

2. `bg_islandareas_raw`: optional raw 2020 Island Areas Census DHC tables for
   American Samoa, Guam, the Commonwealth of the Northern Mariana Islands, and
   the U.S. Virgin Islands. The default EJScreen-compatible release path does
   not need this stage, because AS/GU/MP/VI row IDs, area fields, and available
   environmental fields come from the archived EPA EJScreen ACS2022 reference
   named by `EJAM_ISLANDAREAS_REFERENCE_PATH`. Puerto Rico is not part of this
   stage because it is already covered by ACS.

3. `bg_islandareas_demographics`: optional transformed 2020 Island Areas Census
   DHC demographics. This is saved as a separate checkpoint for review and
   possible supplemental use. The Island Areas source is not ACS 5-year data,
   and the legacy EPA/EJScreen file with AS/GU/MP/VI rows had no usable ACS
   demographic values for those rows. Therefore the default EJScreen-compatible
   pipeline does not use these DHC demographics in `bg_acsdata` or downstream
   outputs. Set `EJAM_USE_ISLANDAREAS_DEMOGRAPHICS = "TRUE"` only when
   intentionally creating a mixed-source supplemental dataset.

4. `bg_acsdata`: ACS-derived blockgroup indicators calculated from
   `bg_acs_raw`, including demographic indicators and `pctpre1960`. When the
   Island Areas stage is enabled, AS/GU/MP/VI rows are appended from the
   archived EPA reference with no DHC-derived demographic values by default.
   The separate `bg_islandareas_demographics` file keeps the available Island
   Areas Census values available for review without changing the
   EJScreen-compatible demographic calculations.

5. `bg_envirodata`: blockgroup environmental indicators used for EJ indexes.
   This normally comes from a separate environmental-data workflow. For draft
   builds, it can be provisionally reused from the current package data.

6. `bg_extra_indicators`: other blockgroup indicators that are not ACS and not
   the main EJ environmental indicators, such as health, life expectancy, and
   related context variables. These can also be provisionally reused from the
   current package data.

7. `bg_geodata`: Census/TIGER block group geography attributes. This stage
   stores `bgfips`, square-meter `arealand` and `areawater`, optional internal
   point fields `intptlat` and `intptlon`, and a compatibility-only `area`
   column. The pipeline uses Census TIGER/Line block group shapefiles by default
   because their `ALAND` and `AWATER` values best match the legacy EJScreen
   tables; Census TIGERweb remains available as a lighter fallback source.
   TIGER/Line zip files are cached locally using `EJAM_TIGER_BG_CACHE_DIR`, or
   the EJAM user cache folder when that variable is unset, so later reruns can
   reuse the state files.
   By default, the pipeline requests geography only for blockgroups found in
   the ACS tabulated rows for that vintage.

8. `blockgroupstats`: combined blockgroup table with ACS indicators,
   environmental indicators, extra indicators, and geography fields.

9. `usastats_acs`, `statestats_acs`, `usastats_envirodata`, and
   `statestats_envirodata`: percentile lookup tables for ACS and environmental
   inputs.

10. `bgej`: blockgroup EJ index values calculated from demographic indexes and
   environmental percentiles.

11. `usastats_ej` and `statestats_ej`: percentile lookup tables for EJ index
   columns.

12. `usastats` and `statestats`: combined lookup tables used by EJAM.

13. `ejscreen_export`: EJScreen-ready export combining `blockgroupstats` and
    `bgej`, applying EJScreen-style names from `map_headernames`, and adding
    map helper fields where possible.

14. `ejscreen_export_statepct`: EJScreen-ready export matching EPA's
    StatePct convention, where state raw scores and state percentiles are
    written into the generic EJScreen field names.

15. `ejscreen_us_pctile_lookup` and `ejscreen_state_pctile_lookup`:
    optional EJScreen-style percentile lookup CSVs created from `usastats` and
    `statestats`. These use EJScreen field names and add `std` rows to match
    EPA lookup tables such as `EJScreen_2024_BG_National_Lookup.csv` and
    `EJScreen_2024_BG_State_Lookup.csv`. The annual pipeline does not create
    these by default because the live EJScreen app maps from the blockgroup
    exports and reports are served through EJAM-API/EJAM.

16. `ejscreen_dataset_creator_input`: optional smaller input table for EPA's
    Python `ejscreen-dataset-creator-2.3` workflow. Enable it with
    `EJAM_INCLUDE_EJSCREEN_DATASET_CREATOR_INPUT = "TRUE"`.

The runner also writes `pipeline_validation_summary.csv` and
`pipeline_run_manifest.csv`. If prior-version validation is requested, it writes
`prior_validation_summary.csv` and per-stage prior-validation details. If the
EJScreen export is requested, it writes `ejscreen_export_schema_report.csv`.
For ACS 2018-2022 replication runs, the runner can also compare
`ejscreen_export` to the EPA-style
`EJSCREEN_2024_BG_with_AS_CNMI_GU_VI.csv` reference file and write
`prior_validation_ejscreen_export_vs_epa_2024_acs2022*` reports.

The run manifest records the package version, Git branch and SHA, ACS vintage,
pipeline location, primary stage format, selected run settings, and whether
provisional environmental or extra-indicator inputs were reused.

## Related Data Update Groups

EJAM data objects do not all change on the same schedule. For annual work, use
these update groups to decide what must be rebuilt, validated, or only checked.

1. **Facility Data Updates** include FRS-related and facility-code datasets such
   as `frs`, `frs_by_programid`, `frs_by_naics`, `frs_by_sic`,
   `frs_by_mact`, `frsprogramcodes`, `epa_programs`, NAICS, SIC, and MACT
   lookup tables. These can be refreshed when EPA facility data are updated,
   independently of the annual EJScreen-style pipeline.

2. **EJSCREEN Annual Data Update** includes the main pipeline stages:
   `bg_acs_raw`, `bg_acsdata`, `bg_envirodata`, `bg_extra_indicators`,
   `bg_geodata`, `blockgroupstats`, `bgej`, `usastats`, `statestats`, and
   `ejscreen_export`. It also includes supporting objects that may need review
   or regeneration, such as `map_headernames`, `names_*`, `namez`,
   `tables_ejscreen_acs`, `formulas_ejscreen_acs`,
   `formulas_ejscreen_acs_disability`, `formulas_ejscreen_demog_index`,
   `avg.in.us`, `high_pctiles_tied_with_min`, and `testoutput*`.
   In practice, `dataload_dynamic()` uses the `ejamdata_required_tag` field in
   DESCRIPTION to find the required `ejamdata` release tag, such as `v3`,
   for Arrow assets. This tag usually matches the EJAM package version for an
   annual data release, but can differ for patch releases. `bgej.arrow` is part
   of this annual release bundle and is additionally checked against the
   installed `blockgroupstats` blockgroup universe.

3. **Blockgroup Geography Updates** include blockgroup-keyed geography files
   and crosswalks such as `bgid2fips`, `blockwts`, `bgpts`, and
   `bg_cenpop2020`. The annual pipeline always checks whether these are still
   compatible with the current blockgroup universe, but they only need to be
   regenerated when blockgroup FIPS, EJAM `bgid`, internal points, or
   blockgroup-to-block relationships change. `bg_cenpop2020` requires special
   care because it is tied to Census 2020 geography. Related state/place
   geography objects such as `states_shapefile`, `stateinfo2`, and
   `censusplaces` should also be checked when their FIPS or boundaries change.
   For the v3 Island Area decision and live EJSCREEN layer availability
   notes, see [Island Areas in EJAM v3](island-areas.html).

4. **Block Geography Updates** include block-level geometry/index files such as
   `blockpoints`, `quaddata`, and `blockid2fips`. These change only when block
   FIPS or block internal-point geography changes.

The runner writes `dynamic_geography_arrow_report.csv` to summarize whether
category 3 and 4 Arrow files cover the current `blockgroupstats` blockgroups
and whether block-level IDs line up across `blockwts`, `blockpoints`,
`quaddata`, and `blockid2fips`.

## ACS Geography Universe

For a given ACS 5-year release, use the Census geography vintage associated
with the ACS end year. For example, ACS 2020-2024 should use 2024 Census/TIGER
or TIGERweb geography attributes. The release pipeline prefers the downloadable
Census TIGER/Line block group shapefiles for `arealand` and `areawater`, with
TIGERweb as a fallback. Those geography sources can occasionally include
block groups that are valid geography features but that are not present in the
ACS tabulated summary-file tables used by EJAM.

That mismatch is unusual but real. In the draft ACS 2020-2024 build, the
Census/TIGER block group geography source included 39 Suffolk County, New York
block groups that were not present in the relevant ACS block group or tract
tables downloaded for the pipeline. Including those geography-only rows would
expand `blockgroupstats` beyond the ACS data universe and create rows with no
ACS-derived indicators.

For that reason, the default pipeline setting is
`EJAM_BLOCKGROUP_UNIVERSE_SOURCE = "acs"`. Under that setting, `bg_acsdata`
defines the final blockgroup universe, `bg_geodata` is downloaded or subset to
those `bgfips` values, and environmental or extra-indicator inputs cannot add
extra rows to the final `blockgroupstats`. The alternative setting
`EJAM_BLOCKGROUP_UNIVERSE_SOURCE = "union"` is kept only for diagnostic or
special-purpose runs where the maintainer intentionally wants to retain
blockgroups found only in other inputs.

The pipeline uses the packaged `formulas_ejscreen_acs` object for ACS-derived
indicator formulas and sorts formula rows by dependency before calculating
them. The old `data-raw/archived_datacreate_formulas_ejscreen_acs_notes.R`
file is reference material only. It is not the current formula rebuild
workflow.

## Storage

The default local pipeline folder is:

```{r, eval=FALSE}
data-raw/pipeline_outputs/ejscreen_acs_2024
```

`data-raw/pipeline_outputs/` is ignored by Git because checkpoint files can be
large. The repository also has build-ignore rules for pipeline outputs and
Arrow data files, so release artifacts should normally be stored outside the
package source tree.

The pipeline can also use AWS S3. S3 support uses the AWS CLI, so `aws` must be
installed and configured before running an S3-backed pipeline. By default the
runner uses `EJAM_STAGE_FORMAT = "csv"` for loading/review and
`EJAM_STAGE_FORMATS = "csv,rda"` for saving major table stages in both formats.
Small summary and manifest files are written as CSV.

```{r, eval=FALSE}
Sys.setenv(
  AWS_PROFILE = "ejam",
  AWS_REGION = "us-east-1",
  EJAM_PIPELINE_DIR = "s3://pedp-data-preserved/ejscreen-data-processing/pipeline/ejscreen_acs_2024",
  EJAM_PIPELINE_STORAGE = "s3"
)
```

For local testing, use a local directory:

```{r, eval=FALSE}
Sys.setenv(
  EJAM_PIPELINE_DIR = file.path(
    getwd(),
    "data-raw",
    "pipeline_outputs",
    "ejscreen_acs_2024"
  ),
  EJAM_PIPELINE_STORAGE = "local"
)
```

## Key Settings

The preferred interface is a validated config object, usually built by one of
the recipe helpers such as `pipeline_config_annual()` or
`pipeline_config_release()`. Environment variables remain supported for
RStudio/source-based workflows and GitHub Actions. The main settings are:

| Variable | Purpose |
|---|---|
| `EJAM_PIPELINE_YR` | ACS 5-year end year, such as `"2024"` for ACS 2020-2024. |
| `EJAM_PIPELINE_DIR` | Local folder or `s3://...` pipeline location. |
| `EJAM_PIPELINE_STORAGE` | `"auto"`, `"local"`, or `"s3"`. |
| `EJAM_STAGE_FORMAT` | Primary stage format used for loading/validation, usually `"csv"`. |
| `EJAM_STAGE_FORMATS` | Comma-separated formats to save for major table stages, usually `"csv,rda"`. |
| `EJAM_BLOCKGROUP_UNIVERSE_SOURCE` | `"acs"` uses the ACS tabulated blockgroup rows as the final universe. `"union"` also keeps rows found only in environmental or extra-indicator inputs. |
| `EJAM_TRACT_WEIGHT_SOURCE` | `"decennial2020"` uses 2020 Decennial Census population weights to apportion tract-only ACS tables to blockgroups, matching legacy EJSCREEN. `"acs"` uses same-vintage ACS blockgroup population weights. |
| `EJAM_DECENNIAL_BGWTS_CACHE` | Optional local `.rds` cache path for 2020 Decennial blockgroup-to-tract weights. If unset, EJAM uses a user cache folder. |
| `EJAM_REFRESH_DECENNIAL_BGWTS` | `"TRUE"` to ignore and overwrite the cached decennial blockgroup weights. |
| `EJAM_TIGER_BG_CACHE_DIR` | Optional local folder for downloaded Census TIGER/Line block group zip files. If unset, EJAM uses a durable user cache folder. |
| `AWS_PROFILE`, `AWS_REGION` | Used by the AWS CLI for S3-backed runs. |
| `CENSUS_API_KEY` | **Required.** Used by ACS/Census helpers for data *and metadata*. tidycensus (>= 1.8) errors without a key, even for `load_variables()`. Set once with `tidycensus::census_api_key("YOUR KEY", install = TRUE)`, then restart R. See `?tidycensus::census_api_key`. |
| `EJAM_FORCE_ACS` | `"TRUE"` to redownload raw ACS and rebuild ACS stages. |
| `EJAM_FORCE_BG_ACSDATA` | `"TRUE"` to rebuild `bg_acsdata` from saved raw ACS. |
| `EJAM_FORCE_BG_GEODATA` | `"TRUE"` to redownload/rebuild the Census/TIGER `bg_geodata` stage. |
| `EJAM_ACS_DOWNLOAD_TIMEOUT` | Download timeout in seconds. Useful for large ACS tables. |
| `EJAM_ACS_DOWNLOAD_RETRIES` | Number of retry attempts for ACS downloads. |
| `EJAM_INCLUDE_ISLANDAREAS_DATA` | `"TRUE"` to append AS/GU/MP/VI rows. The annual/release runner defaults to `"TRUE"` unless explicitly set otherwise. |
| `EJAM_ISLANDAREAS_REFERENCE_PATH` | Archived EPA EJScreen ACS2022 reference CSV used for Island Areas row IDs, area fields, and available environmental fields. |
| `EJAM_USE_ISLANDAREAS_DEMOGRAPHICS` | `"TRUE"` only for an intentional mixed-source supplemental dataset that uses 2020 Island Areas Census DHC demographics in `bg_acsdata`. The default EJScreen-compatible path is `"FALSE"`. |
| `EJAM_USE_PROVISIONAL_BG_ENVIRODATA` | `"FALSE"` to require a supplied `bg_envirodata` stage file, such as `bg_envirodata.csv` under the default stage format. |
| `EJAM_INCLUDE_EJSCREEN_EXPORT` | `"TRUE"` to create the `ejscreen_export` stage, such as `ejscreen_export.csv` under the default stage format. |
| `EJAM_INCLUDE_EJSCREEN_EXPORT_STATEPCT` | `"TRUE"` to create the `ejscreen_export_statepct` stage. |
| `EJAM_INCLUDE_EJSCREEN_PCTILE_LOOKUP_EXPORTS` | `"TRUE"` only when intentionally refreshing EJScreen-style lookup CSVs. These are not created by default. |
| `EJAM_INCLUDE_EJSCREEN_DATASET_CREATOR_INPUT` | `"TRUE"` to also create the smaller `ejscreen_dataset_creator_input` stage for EPA's Python dataset-creator workflow. |
| `EJAM_VALIDATE_VS_PRIOR` | `"TRUE"` to write prior-version comparison files. |
| `EJAM_PRIOR_PIPELINE_DIR` | Prior pipeline folder/S3 prefix to compare against. |
| `EJAM_PRIOR_PACKAGE_REF` | Optional Git ref/tag/SHA for prior package data comparison. |
| `EJAM_EJSCREEN_EXPORT_REFERENCE_PATH` | Optional EPA-style EJScreen export CSV to compare against `ejscreen_export`. For S3-backed 2022 runs, the runner defaults to the preserved `EJSCREEN_2024_BG_with_AS_CNMI_GU_VI.csv` file, which uses ACS 2018-2022 data despite its 2024 filename. |
| `EJAM_VALIDATE_EJSCREEN_EXPORT_REFERENCE` | `"TRUE"` to write `prior_validation_ejscreen_export_vs_epa_2024_acs2022.csv`, `*_summary.csv`, and `*_summary.txt` when a reference export path is available. |

To see what the runner will use:

```{r, eval=FALSE}
Sys.getenv(c(
  "EJAM_PIPELINE_YR",
  "EJAM_PIPELINE_DIR",
  "EJAM_PIPELINE_STORAGE",
  "EJAM_STAGE_FORMAT",
  "EJAM_STAGE_FORMATS",
  "EJAM_BLOCKGROUP_UNIVERSE_SOURCE",
  "EJAM_TRACT_WEIGHT_SOURCE",
  "EJAM_DECENNIAL_BGWTS_CACHE",
  "EJAM_REFRESH_DECENNIAL_BGWTS",
  "EJAM_TIGER_BG_CACHE_DIR",
  "AWS_PROFILE",
  "AWS_REGION",
  # "CENSUS_API_KEY",
  "EJAM_FORCE_ACS",
  "EJAM_FORCE_BG_ACSDATA",
  "EJAM_FORCE_BG_GEODATA",
  "EJAM_ACS_DOWNLOAD_TIMEOUT",
  "EJAM_ACS_DOWNLOAD_RETRIES",
  "EJAM_INCLUDE_ISLANDAREAS_DATA",
  "EJAM_USE_ISLANDAREAS_DEMOGRAPHICS",
  "EJAM_USE_PROVISIONAL_BG_ENVIRODATA",
  "EJAM_INCLUDE_EJSCREEN_EXPORT",
  "EJAM_INCLUDE_EJSCREEN_EXPORT_STATEPCT",
  "EJAM_INCLUDE_EJSCREEN_PCTILE_LOOKUP_EXPORTS",
  "EJAM_INCLUDE_EJSCREEN_DATASET_CREATOR_INPUT",
  "EJAM_VALIDATE_VS_PRIOR",
  "EJAM_PRIOR_PIPELINE_YR",
  "EJAM_PRIOR_PIPELINE_DIR",
  "EJAM_PRIOR_PACKAGE_REF",
  "EJAM_EJSCREEN_EXPORT_REFERENCE_PATH",
  "EJAM_VALIDATE_EJSCREEN_EXPORT_REFERENCE"
))
```

For ACS 2022 and later, Connecticut ACS tract FIPS use planning-region county
equivalents while 2020 Decennial blockgroup FIPS use the older county
equivalents. The pipeline detects that no Connecticut tract FIPS overlap in the
decennial weight table and uses same-vintage ACS blockgroup population weights
for Connecticut only. In normal package use, the decennial weights are created
from packaged `bg_cenpop2020` data. If that data is unavailable, EJAM falls back
to `tidycensus::get_decennial()` and caches the downloaded weights locally.

## New or renamed indicators

### map_headernames

If new indicators are being used compared with prior version of the datasets and package, 
`map_headernames` may need metadata rows for those new indicators, including
the variable name (`rname`), `longname`, calculation type, calculation weight,
rounding information, EJScreen export names, and `varlist` groups such as
`names_e` and `names_d`. The editable source for release work is
`data-raw/map_headernames.csv`. Edit that CSV directly, then source
`data-raw/datacreate_map_headernames.R` to validate and save
`data/map_headernames.rda`. Older spreadsheet workflows are obsolete and
should not be used to regenerate this object.

### names_*

Much of the code depends on the `varlist` info, so `names_e`, `names_d`, and
related `names_*` objects need to be updated when `map_headernames$varlist`
changes. The script in `data-raw/datacreate_names_of_indicators.R` uses
`map_headernames$varlist` to update those data objects.


## Run a Fresh ACS Update

> **Prerequisite: a Census API key.** As of tidycensus (>= 1.8) the Census Bureau
> requires an API key for all requests, including the `load_variables()` metadata lookup
> EJAM uses to resolve ACS table geography. tidycensus now *errors* (no longer warns)
> without one. Get a free key at <https://api.census.gov/data/key_signup.html> and set it
> once with `tidycensus::census_api_key("YOUR KEY", install = TRUE)`, then restart R.
> (The bulk ACS download via `ACSdownload::get_acs_new()` reads Census summary files, but
> the metadata step still needs the key.)

Start from a clean branch. For ACS 2020-2024 using local checkpoints:

```{r, eval=FALSE}
yr <- 2024

cfg <- EJAM:::pipeline_config_annual(
  yr = yr,
  pipeline_dir = file.path(
    getwd(),
    "data-raw",
    "pipeline_outputs",
    paste0("ejscreen_acs_", yr)
  ),
  pipeline_storage = "local",
  force_acs = TRUE,
  force_bg_acsdata = TRUE,
  force_bg_geodata = TRUE
)

pipeline_run <- EJAM:::run_ejscreen_pipeline(cfg)
```

For the S3-backed pipeline:

```{r, eval=FALSE}
yr <- 2024

cfg <- EJAM:::pipeline_config_annual(
  yr = yr,
  pipeline_root = "s3://pedp-data-preserved/ejscreen-data-processing/pipeline",
  pipeline_storage = "s3",
  force_acs = TRUE,
  force_bg_acsdata = TRUE,
  force_bg_geodata = TRUE,
  aws_profile = "ejam",
  aws_region = "us-east-1"
)

pipeline_run <- EJAM:::run_ejscreen_pipeline(cfg)
```

The runner prints the resolved settings before it starts the pipeline. Review
those settings carefully, especially the year, storage backend, stage formats,
force flags, Island Areas settings, provisional-input flags, and prior
validation target.

For ACS2024/v3 and later annual/release runs, AS/GU/MP/VI rows are added by
default at the blockgroup dataset, EJSCREEN export, and map-data visibility
level, in the same general style as the legacy EPA/EJScreen export. The default
source for Island Areas row IDs, area fields, and available environmental
fields is the archived EPA EJScreen ACS2022 reference named by
`EJAM_ISLANDAREAS_REFERENCE_PATH`. Keep the DHC demographics out of
`bg_acsdata` unless you are intentionally creating a mixed-source supplemental
dataset. To make the default explicit:

```{r, eval=FALSE}
cfg <- EJAM:::pipeline_config_annual(
  yr = 2024,
  include_islandareas_data = TRUE,
  use_islandareas_demographics = FALSE
)

pipeline_run <- EJAM:::run_ejscreen_pipeline(cfg)
```

Set `EJAM_USE_ISLANDAREAS_DEMOGRAPHICS = "TRUE"` only for an intentional
mixed-source supplemental dataset. That uses 2020 Island Areas Census DHC
demographic values in `bg_acsdata`, which is useful for review but is not the
default EJScreen replication path.
Set `EJAM_INCLUDE_ISLANDAREAS_DATA = "FALSE"` only when deliberately creating a
States/DC/PR-only run for comparison or debugging.

## Rerun From Saved ACS Data

If raw ACS has already been downloaded, rerun downstream ACS calculations
without redownloading:

```{r, eval=FALSE}
cfg <- EJAM:::pipeline_config_annual(
  yr = 2024,
  force_acs = FALSE,
  force_bg_acsdata = TRUE
)

pipeline_run <- EJAM:::run_ejscreen_pipeline(cfg)
```

If both raw ACS and `bg_acsdata` should be reused, leave both force flags false:

```{r, eval=FALSE}
cfg <- EJAM:::pipeline_config_annual(
  yr = 2024,
  force_acs = FALSE,
  force_bg_acsdata = FALSE
)

pipeline_run <- EJAM:::run_ejscreen_pipeline(cfg)
```

If `bg_geodata` has already been created for the same ACS/TIGER vintage and
same blockgroup universe, leave `EJAM_FORCE_BG_GEODATA` false. Set it to
`"TRUE"` when changing vintages or when you want to refresh the Census
TIGER/Line area and internal-point attributes.
Even with `EJAM_FORCE_BG_GEODATA = "TRUE"`, already-downloaded TIGER/Line state
zip files are reused from `EJAM_TIGER_BG_CACHE_DIR` when present and valid.

```{r, eval=FALSE}
cfg <- EJAM:::pipeline_config_annual(
  yr = 2024,
  force_acs = FALSE,
  force_bg_acsdata = FALSE,
  force_bg_geodata = TRUE
)

pipeline_run <- EJAM:::run_ejscreen_pipeline(cfg)
```

## Supplying Updated Environmental Data

The environmental stage is intentionally separate from the ACS stage. When
updated environmental indicators are available, save them in the pipeline
folder as the stage file for `bg_envirodata`. With the default stage format,
that file is `bg_envirodata.csv`.

For a local pipeline:

```{r, eval=FALSE}
file.path(pipeline_dir, "bg_envirodata.csv")
```

For an S3 pipeline:

```{r, eval=FALSE}
s3://pedp-data-preserved/ejscreen-data-processing/pipeline/ejscreen_acs_2024/bg_envirodata.csv
```

The file must include `bgfips` and the environmental indicators used for EJ
indexes. It should also include `pctpre1960`. The environmental-data workflow
may create `pctpre1960` by reading the saved `bg_acsdata` stage.

Environmental indicator missing values should be preserved as missing values.
Do not convert `NA` values to zero unless the source explicitly reports a valid
zero score. This is especially important for the drinking-water
non-compliance indicator. EJAM versions through v2.32.8.001 converted missing
EPA `DWATER` values to `drinking = 0` in `blockgroupstats`; later EJAM releases
should preserve the distinction between missing/no valid score and a valid zero
score.

To force the runner to stop unless `bg_envirodata.csv` has been supplied:

```{r, eval=FALSE}
cfg <- EJAM:::pipeline_config_annual(
  yr = 2024,
  use_provisional_bg_envirodata = FALSE
)

pipeline_run <- EJAM:::run_ejscreen_pipeline(cfg)
```

After replacing `bg_envirodata.csv`, rerun without forcing ACS:

```{r, eval=FALSE}
cfg <- EJAM:::pipeline_config_annual(
  yr = 2024,
  force_acs = FALSE,
  force_bg_acsdata = FALSE,
  force_bg_geodata = FALSE,
  use_provisional_bg_envirodata = FALSE
)

pipeline_run <- EJAM:::run_ejscreen_pipeline(cfg)
```

This reuses the saved ACS stages and regenerates downstream
`blockgroupstats`, `bgej`, `usastats`, `statestats`, `ejscreen_export`,
and `ejscreen_export_statepct`. EJScreen-style lookup exports are created only
when `include_ejscreen_pctile_lookup_exports = TRUE` in the config, or
`EJAM_INCLUDE_EJSCREEN_PCTILE_LOOKUP_EXPORTS = "TRUE"` in compatibility-runner
workflows.

## Extra Indicators

Some `blockgroupstats` columns are not ACS indicators and are not the main
environmental indicators. Examples include life expectancy, health indicators,
facility-count context variables, climate-related fields, and other columns
grouped in `map_headernames$varlist`.

The pipeline makes these explicit in the `bg_extra_indicators` stage, usually
`bg_extra_indicators.csv` under the default stage format. If an updated table is
not supplied, the runner currently creates a provisional version from the
packaged `EJAM::blockgroupstats`. That is useful for testing the ACS update,
but final release review should document clearly any reuse of older non-ACS
data.

## Provisional Draft Builds

For early pipeline testing, it is acceptable to reuse existing environmental and
extra indicators:

```{r, eval=FALSE}
cfg <- EJAM:::pipeline_config_annual(
  yr = 2024,
  use_provisional_bg_envirodata = TRUE
)

pipeline_run <- EJAM:::run_ejscreen_pipeline(cfg)
```

The runner writes source-note text files next to provisional stages, such as
`bg_envirodata_SOURCE.txt` and `bg_extra_indicators_SOURCE.txt`. Final release
review should confirm whether any provisional stage remains.

## Reviewing Outputs

Start with the validation summary:

```{r, eval=FALSE}
library(data.table)

pipeline_dir <- "data-raw/pipeline_outputs/ejscreen_acs_2024"

validation <- fread(file.path(pipeline_dir, "pipeline_validation_summary.csv"))
validation[, .(stage, rows, columns, errors, warnings)]
validation[nzchar(errors)]
validation[nzchar(warnings)]
```

There should be no validation errors. Warnings should be understood and either
fixed or explicitly accepted for a draft build.

Then inspect the main outputs. The example below assumes the default CSV stage
format:

```{r, eval=FALSE}
bg_acsdata      <- fread(file.path(pipeline_dir, "bg_acsdata.csv"))
blockgroupstats <- fread(file.path(pipeline_dir, "blockgroupstats.csv"))
bgej            <- fread(file.path(pipeline_dir, "bgej.csv"))
usastats        <- fread(file.path(pipeline_dir, "usastats.csv"))
statestats      <- fread(file.path(pipeline_dir, "statestats.csv"))
bg_geodata      <- fread(file.path(pipeline_dir, "bg_geodata.csv"))

nrow(blockgroupstats)
nrow(bgej)
names(blockgroupstats)
names(bgej)

island_prefixes <- c("60", "66", "69", "78")
for (stage in list(bg_acsdata, blockgroupstats, bgej)) {
  print(stage[substr(bgfips, 1, 2) %in% island_prefixes, .N, by = ST])
}
```

Useful checks include:

- expected FIPS/geography columns are present and typed as character;
- row counts are plausible for the ACS vintage and geography coverage;
- under the default `EJAM_BLOCKGROUP_UNIVERSE_SOURCE = "acs"`,
  `blockgroupstats`, `bgej`, and `bg_geodata` have the same `bgfips` values as
  `bg_acsdata`;
- key ACS indicators are non-missing for most populated blockgroups;
- percentage/rate variables are in the expected range;
- `blockgroupstats` and `bgej` join cleanly by `bgfips`;
- `bg_geodata` has one row per `bgfips` and non-missing, nonnegative
  `arealand` and `areawater`;
- lookup tables include `REGION`, `PCTILE`, `0`, `100`, and `mean`;
- `usastats` has one region, `"USA"`;
- `statestats` has expected state/territory regions.
- if optional EJScreen-style lookup exports were requested, they include
  `PCTILE`, `REGION`, `0`, `100`, `mean`, and `std`, with EJScreen field names
  rather than EJAM `rname` columns.
- for ACS2024/v3, Island Areas AS/GU/MP/VI are present in `bg_acsdata`,
  `blockgroupstats`, `bgej`, `ejscreen_export`, and `ejscreen_export_statepct`
  unless `EJAM_INCLUDE_ISLANDAREAS_DATA = "FALSE"`.

For the default EJScreen-compatible path, Island Areas demographic columns in
`bg_acsdata` and downstream stages are expected to be `NA` because the archived
EPA/EJScreen reference file with AS/GU/MP/VI rows had no usable ACS demographic
values for those rows. The separate `bg_islandareas_demographics` checkpoint is
optional and contains available 2020 Island Areas Census DHC values for review.
Those values are used in `bg_acsdata` only when
`EJAM_USE_ISLANDAREAS_DEMOGRAPHICS = "TRUE"`. Available EPA environmental and
area fields for AS/GU/MP/VI are retained from the reference where supplied.
This visibility-level support does not mean radius/buffer or block-weighted
polygon analysis works in Island Areas. Island Area blocks are not added to
`blockwts`, `blockpoints`, `bgid2fips`, `blockid2fips`, `quaddata`, or related
helper files for this release path, so analyses there should return no-data
results rather than block-weighted estimates.

For example:

```{r, eval=FALSE}
stopifnot("0" %in% as.character(usastats$PCTILE))
stopifnot("100" %in% as.character(usastats$PCTILE))
stopifnot("mean" %in% as.character(usastats$PCTILE))

stopifnot("0" %in% as.character(statestats$PCTILE))
stopifnot("100" %in% as.character(statestats$PCTILE))
stopifnot("mean" %in% as.character(statestats$PCTILE))

if (file.exists(file.path(pipeline_dir, "ejscreen_us_pctile_lookup.csv"))) {
  ejscreen_us_pctile_lookup <- fread(file.path(pipeline_dir, "ejscreen_us_pctile_lookup.csv"))
  stopifnot("std" %in% as.character(ejscreen_us_pctile_lookup$PCTILE))
  stopifnot(all(c("PCTILE", "REGION", "DEMOGIDX_2", "LOWINCPCT", "D2_PM25") %in%
                  names(ejscreen_us_pctile_lookup)))
}

if (file.exists(file.path(pipeline_dir, "ejscreen_state_pctile_lookup.csv"))) {
  ejscreen_state_pctile_lookup <- fread(file.path(pipeline_dir, "ejscreen_state_pctile_lookup.csv"))
  stopifnot("std" %in% as.character(ejscreen_state_pctile_lookup$PCTILE))
}

stopifnot(!anyDuplicated(blockgroupstats$bgfips))
stopifnot(!anyDuplicated(bgej$bgfips))
stopifnot(!anyDuplicated(bg_geodata$bgfips))
stopifnot(setequal(blockgroupstats$bgfips, bg_acsdata$bgfips))
stopifnot(setequal(bgej$bgfips, bg_acsdata$bgfips))
stopifnot(setequal(bg_geodata$bgfips, bg_acsdata$bgfips))
stopifnot(all(bg_geodata$arealand >= 0, na.rm = TRUE))
stopifnot(all(bg_geodata$areawater >= 0, na.rm = TRUE))
```

Also review the run manifest:

```{r, eval=FALSE}
manifest <- fread(file.path(pipeline_dir, "pipeline_run_manifest.csv"))
manifest[key %in% c(
  "package_version",
  "git_sha",
  "git_branch",
  "git_dirty",
  "acs_version",
  "stage_format",
  "setting_EJAM_STAGE_FORMATS",
  "used_provisional_bg_envirodata",
  "used_provisional_bg_extra_indicators"
)]
```

For S3-backed runs, read the same files from S3 using EJAM's pipeline input
helpers or the AWS CLI.

## Slow Stages

The longest stages are usually:

- fresh `bg_acs_raw` downloads, because several ACS table-based summary files
  are large;
- S3 writes of large raw and derived stage files, especially when saving both
  `.csv` and `.rda`;
- `bg_acsdata`, because it applies ACS formulas and tract-to-blockgroup
  calculations;
- `bg_geodata`, because it downloads and reads Census TIGER/Line block group
  shapefiles for every state, the District of Columbia, Puerto Rico, and
  optional Island Areas, with TIGERweb available as a fallback. Reusing
  `EJAM_TIGER_BG_CACHE_DIR` makes later `bg_geodata` rebuilds much faster
  because the state or territory zip files do not need to be downloaded again;
- final `blockgroupstats`/`bgej`/statistics/export calculations;
- prior-version validation when comparing large blockgroup tables.

When debugging pipeline speed, check the console timestamps and consider
rerunning without forcing earlier stages once their saved outputs are known to
be current.

## Reviewing the EJScreen Export Schema

> **EJSCREEN web-app layers (issue #395).** The two block-group export files here
> *are* the web app's main percentile layers: `ejscreen_export.csv` is the **US
> percentiles** block-group layer and `ejscreen_export_statepct.csv` is the
> **State percentiles** layer. Their columns match the **235-field** EJSCREEN
> FeatureServer schema exactly (verified via `ejscreen_feature_server_fields()`;
> `OBJECTID`/`Shape__*` are added by ArcGIS on upload), and
> `ejscreen_export_schema_report.csv` validates each run.

### EJSCREEN web-app extra layers produced by the ejscreen_export stage

When `EJAM_INCLUDE_EJSCREEN_EXPORT` is true, the `ejscreen_export` stage now also
writes the remaining web-app outputs (issue #395), as CSV stage files:

- **ACS "additional demographics" by geography** — `acs_by_blockgroup`,
  `acs_by_tract`, `acs_by_county`, `acs_by_state`, built by
  [calc_acs_by_geography()] (count columns summed; percentage columns recomputed
  as denominator-weighted means via [calctype()]/[calcweight()], the same rules
  [doaggregate()] uses). Zero-denominator block groups (no demographic basis) get
  `NA`, which is expected. The block-group level matches the ACS values already in
  `blockgroupstats`/`ejscreen_export`.
- **Four threshold layers** — `ejscreen_threshold_us_ejindexes`,
  `ejscreen_threshold_us_supplemental`, `ejscreen_threshold_state_ejindexes`,
  `ejscreen_threshold_state_supplemental`, built by
  [calc_ejscreen_threshold_layers_from_exports()] from the per-block-group
  `P_D2_*`/`P_D5_*` percentile ranks already in the national and state exports.
  Each adds the `P1`..`P100` hit-count columns.

This does **not** include the full EPA "extensive ACS" (~951 extra demographic
breakdowns); only the indicators EJAM already computes are aggregated. A future
enhancement can have this stage emit an EJSCREEN-formatted "extensive ACS" table
built from extra indicators **supplied as an input in the S3 folder**, rather than
newly calculated here.

### Replication of the web-app layers against EPA

For an **ACS 2018-2022** build, compare these new outputs to EPA's published
EJScreen values. EPA exposes them as ArcGIS **FeatureServers** (the
`EJScreen_Census` service's `by_Tract`/`by_County`/`by_State` layers, and the
EJ-index threshold layers), **not** as archived CSVs in the S3 pipeline folders —
so the replication fetches the reference live and uses
[ejscreen_compare_geography_to_epa()] (which repairs FIPS leading zeros with
[fips_lead_zero()], maps EPA names with `fixcolnames(., "acs", "r")`, joins on the
full GEOID, and reports correlation + median difference per indicator). This
verified an **exact** match for the ACS indicators EJAM carries.

```{r, eval=FALSE}
# EPA tract reference (live feature server; no S3 archive). STCNTR = full GEOID.
fs <- paste0("https://services.arcgis.com/EXyRv0dqed53BmG2/arcgis/rest/services/",
             "EJScreen_Census/FeatureServer/5/query",
             "?where=1%3D1&outFields=STCNTR,TOTALPOP,LOWINC,PCT_LOWINC,PCT_MINORITY",
             "&returnGeometry=false&resultRecordCount=2000&f=json")
epa_tracts <- jsonlite::fromJSON(fs)$features$attributes
acs_tract  <- outputs$acs_by_tract   # the pipeline stage output

EJAM:::ejscreen_compare_geography_to_epa(
  ejam_table = acs_tract, reference = epa_tracts,
  ejam_fips_col = "tractfips", reference_fips_col = "STCNTR"
)
```

If `EJAM_INCLUDE_EJSCREEN_EXPORT` is true, the runner writes the export stage
and its schema report. With the default stage format, those files are:

```{r, eval=FALSE}
ejscreen_export.csv
ejscreen_export_schema_report.csv
```

Use the schema report as a field-by-field checklist:

```{r, eval=FALSE}
schema <- fread(file.path(pipeline_dir, "ejscreen_export_schema_report.csv"))

schema[, .N, by = status]
schema[status == "missing_expected"]
schema[status == "missing_expected", .N, by = field_type][order(-N)]
schema[status == "unexpected_output"]
```

Each missing expected field should be classified as one of:

1. a field EJScreen needs and the export must add;
2. a metadata mapping issue in `map_headernames`;
3. a deliberately deferred field that is not needed for the current export.

For release, the ideal schema report has no `missing_expected` rows for the
FeatureServer fields required by the EJScreen app.

For 2022 replication checks, also review the EPA-reference comparison reports
when they are present:

```{r, eval=FALSE}
prior_validation_ejscreen_export_vs_epa_2024_acs2022.csv
prior_validation_ejscreen_export_vs_epa_2024_acs2022_summary.csv
prior_validation_ejscreen_export_vs_epa_2024_acs2022_summary.txt
```

Those reports use `ID` as a character field so leading zeroes in block group
FIPS are preserved. They are meant to explain differences between EJAM's
current export and the EPA-style reference export, not to force EJAM to
replicate legacy behavior where the new pipeline has intentionally corrected a
formula or missing-value rule.

## Replacing Package Data

Pipeline stage files are review artifacts. After they are accepted, update the
package data objects deliberately. The runner currently has an interactive
helper path for replacing `blockgroupstats`, but it does not automatically
replace every final package dataset. Work on the vintage's branch
(`ACS2024` / `ACS2023` / `ACS2022`) with the package version and metadata fields in
`DESCRIPTION` already set, because the metadata helpers stamp datasets from
`DESCRIPTION` via `desc::description$new("DESCRIPTION")` and
`ejam_package_version_current()`.

### Read first: the `bgid` type must be `double`

Island Areas (AS/GU/MP/VI) get FIPS-derived `bgid` values that overflow 32-bit
integer (they range to ~7.8e9; 416 such rows), so the pipeline stores `bgid` as
**character**. The package geography Arrow files and the engine historically used
**integer** `bgid`, and a character-vs-integer join fails with
`Incompatible join types: x.bgid (character) and i.bgid (integer)`.

**Store `bgid` as `double` everywhere** — in `blockgroupstats`, `bgej`, `blockwts`,
`bgid2fips`, and the engine's `sites2blocks`. Doubles represent integers exactly up
to 2^53 (~9.0e15), far above the largest `bgid`, and `bgid` is only ever a join key
(never indexed or used in arithmetic), so there is no precision loss and it is far
cheaper than character.

> **Do not coerce `bgid` to integer.** That turns the 416 Island-Area `bgid`s into
> `NA` ("NAs introduced by coercion to integer range") and silently drops those
> block groups.

### Avoid the load-time Arrow-download trap

`.onAttach()` calls `download_latest_arrow_data()` for the `ejamdata_required_tag`
in `DESCRIPTION`. If you have already bumped that tag to the new release but have
not published the matching `ejamdata` release yet, attaching the package —
including `devtools::load_all()` — will try to fetch a tag that does not exist and
fail or hang.

Keep `ejamdata_required_tag` at the **previously installed** tag during the data
work (so it matches `data/ejamdata_version.txt` and the present local `.arrow`
files, and `download_latest_arrow_data()` short-circuits with no network), and flip
it to the new tag only after publishing (see
[Updating and Managing the Datasets Used by EJAM](dev-update-datasets.html)).
Always use `devtools::load_all()` rather than an installed `library(EJAM)` here, so
`ejam_package_version_current()` reads the **source** `DESCRIPTION` (the new
`3.YYYY.0`), not a stale installed version.

A release update should explicitly replace at least:

- `blockgroupstats`;
- `usastats`;
- `statestats`;
- any related lookup or metadata objects that changed.

Use the established EJAM metadata helpers before saving package `.rda` data.
When reviewing CSV stages manually, a typical pattern is:

```{r, eval=FALSE}
library(arrow); library(data.table)

# blockgroupstats: coerce character bgid -> double, then stamp + save data/blockgroupstats.rda
blockgroupstats <- fread(file.path(pipeline_dir, "blockgroupstats.csv"))  # or load the .rda stage
setDT(blockgroupstats)[, bgid := as.numeric(bgid)]
EJAM:::metadata_add_and_use_this("blockgroupstats")

# usastats / statestats: no bgid; just stamp + save
usastats        <- fread(file.path(pipeline_dir, "usastats.csv"))
statestats      <- fread(file.path(pipeline_dir, "statestats.csv"))
EJAM:::metadata_add_and_use_this("usastats")
EJAM:::metadata_add_and_use_this("statestats")

# avg.in.us: the mean-row slice of indicator columns from the NEW usastats
longlist  <- unique(c(names_e, names_d, names_d_subgroups_nh, names_d_subgroups_alone,
                      names_d_language, names_d_extra))
longlist  <- longlist[longlist %in% names(usastats)]
avg.in.us <- usastats[usastats$PCTILE == "mean", longlist]
EJAM:::metadata_add_and_use_this("avg.in.us")
```

Confirm the exact metadata values before saving, especially ACS vintage,
Census vintage, EJSCREEN/EJAM version, data source notes, and provisional reuse
notes. `bgej` should be checked and saved through the pipeline stage files, then
published as `bgej.arrow` in the `ejamdata` release tag recorded in
DESCRIPTION as `ejamdata_required_tag`, such as `v3.2024.0` for the matching EJAM v3 release. A
local `data/bgej.arrow` copy can be useful for testing from source, but it is
ignored for package builds and should not be treated as normal package data.

The v3 `blockgroupstats` is the full EJScreen schema — about 404 columns and
~242,983 rows including Island Areas — much wider than the 173-column v2.5.0
release. Confirm this against `pipeline_validation_summary.csv` rather than treating
it as an anomaly.

### Convert `bgid` to `double` in the Arrow datasets

Only three Arrow files carry `bgid`: `bgej` (string from the stage), `blockwts`
(int32 — this feeds `sites2blocks$bgid` in `getblocksnearby()`), and `bgid2fips`
(int32). `blockpoints`, `quaddata`, `blockid2fips`, and the `frs*` files have no
`bgid`, so they carry over unchanged.

```{r, eval=FALSE}
# bgej from the accepted stage file, bgid -> double
e <- new.env(); load(file.path(pipeline_dir, "bgej.rda"), envir = e)
bgej <- e$bgej; setDT(bgej)[, bgid := as.numeric(bgid)]
arrow::write_feather(bgej, "data/bgej.arrow")

# geography Arrow files
bw <- as.data.table(arrow::read_feather("data/blockwts.arrow")); bw[, bgid := as.numeric(bgid)]
arrow::write_feather(bw, "data/blockwts.arrow")
bf <- as.data.table(arrow::read_feather("data/bgid2fips.arrow")); bf[, bgid := as.numeric(bgid)]
arrow::write_feather(bf, "data/bgid2fips.arrow")
```

`data/*.arrow` files are gitignored build artifacts (published to `ejamdata`, not
committed). Known gap: the current `bgid2fips` and `blockwts` come from the
integer-era geography build, so they exclude Island Areas (e.g. `bgid2fips`
~242,355 rows vs `blockgroupstats` ~242,983). Normal queries are unaffected
(`doaggregate()` reads `bgfips` from `blockgroupstats`), but rebuilding the
geography Arrow files to include Island-Area `bgid`s is a clean follow-up now that
the key is `double`.

### Regenerate `testoutput_*`

Regenerate the `testoutput_*` fixtures **after** the data swap and with the new
`bgej`/`blockwts` present locally, so the engine runs on the new vintage —
otherwise the outputs silently encode the old vintage. The driver
`data-raw/datacreate_testpoints_testoutputs.R` defines
`pkg_update_testpoints_testoutputs()` and then auto-calls it with heavy defaults, so
define the function only and call it with conservative flags:

```{r, eval=FALSE}
# define the function without the script's heavy trailing auto-call
for (ex in parse("data-raw/datacreate_testpoints_testoutputs.R"))
  if (is.call(ex) && length(ex) >= 3 &&
      identical(ex[[2]], as.name("pkg_update_testpoints_testoutputs"))) eval(ex, globalenv())

pkg_update_testpoints_testoutputs(
  do_load_all = TRUE, nvalues = c(10, 100, 1000),
  recreating_getblocksnearby = TRUE, resaving_getblocksnearby_rda = TRUE,
  recreating_doaggregate_output = TRUE, resaving_doaggregate_rda = TRUE,
  recreating_ejamit_output = TRUE, resaving_ejamit_rda = TRUE,
  resaving_ejam2excel = FALSE, resaving_ejam2report = FALSE, resaving_ejam2report_pdf = FALSE)

# FIPS + shapes fixtures (simple ejamit() calls), then stamp each
testoutput_ejamit_fips_counties <- ejamit(fips = testinput_fips_counties)
testoutput_ejamit_fips_cities   <- ejamit(fips = testinput_fips_cities)
testoutput_ejamit_shapes_2      <- ejamit(shapefile = testinput_shapes_2, radius = 0)
EJAM:::metadata_add_and_use_this("testoutput_ejamit_fips_counties")
EJAM:::metadata_add_and_use_this("testoutput_ejamit_fips_cities")
EJAM:::metadata_add_and_use_this("testoutput_ejamit_shapes_2")
```

The `nvalues = 10000` case and the Excel/HTML/PDF example outputs are heavy (and the
PDF needs pandoc/LaTeX); they are not the `testoutput_*` `.rda` fixtures, so they can
be skipped here.

Then run `EJAM:::metadata_check()` and `EJAM:::metadata_check_print()` from the
current source package to confirm that package datasets with metadata-style
attributes have the expected EJAM version, ACS version, release dates, and save
dates. Atomic name-vector objects such as many `names_*` datasets do not need
metadata attributes.

### Restamp remaining metadata and verify

Bump `ejam_package_version` on every other metadata-bearing `.rda`
(FRS/geography/reference/`names_*`); atomic name vectors are skipped automatically.

```{r, eval=FALSE}
mc    <- as.data.frame(EJAM:::metadata_check(which = "ejam_package_version"))
stale <- mc$item[!is.na(mc$ejam_package_version) & mc$ejam_package_version != "3.YYYY.0"]
EJAM:::metadata_update_attr(x = stale[!grepl("^testoutput", stale)])
EJAM:::metadata_check_print()   # expect 0 stale
```

Confirm the join key and a real run before committing, then restore
`ejamdata_required_tag` to the new tag, reinstall the package, and rerun
release-critical tests and `R CMD check`:

```{r, eval=FALSE}
class(blockgroupstats$bgid)                 # "numeric"
EJAM:::ejamdata_required_tag()              # the new tag, after you restore it
out <- ejamit(testpoints_10, radius = 1, quiet = TRUE)  # no "Incompatible join types" error
```

The regenerated `testoutput_*` fixtures should make the comparison tests pass,
because the engine output and the fixtures came from the same vintage build.

### Pipeline provisions so future builds emit `double` `bgid`

So a fresh pipeline run produces `double` `bgid` without manual coercion, the
pipeline source carries these provisions:

- `R/calc_ejscreen_blockgroupstats.R` — coerce `bgid` to numeric before saving
  `blockgroupstats`;
- `R/calc_ejscreen_stats.R` — coerce `bgej$bgid` to double after `calc_bgej()`;
- `R/download_bg_islandareas_raw.R` — set the Island-Areas `bgid` from `bgfips` as
  numeric (the root cause);
- `R/getblocksnearby_from_fips.R` — empty-result templates use `bgid = numeric(0)`;
  the FIPS path gets `bgid` from the now-`double` `bgid2fips` join.

## Comparing Two Vintages Side by Side

While testing an annual update it is useful to run two vintages/releases at once
(for example the new `ACS2024` / `v3.2024.0` build and the prior `ACS2023` /
`v3.2023.0` build) and compare the same sites in each. Each EJAM release pins its
own `ejamdata_required_tag` in DESCRIPTION, so each pulls the matching vintage of
the large Arrow datasets (`bgej`, etc.) from its tagged `ejamdata` release.

The cleanest approach is one **git worktree per vintage**, so each checkout has its
own `data/` folder and Arrow cache and there is no cross-vintage collision:

```{r, eval=FALSE}
# from the main checkout, create a worktree per release branch/tag:
# (run in a terminal)
# git worktree add ../EJAM-2024 ACS2024
# git worktree add ../EJAM-2023 ACS2023

# then, in a SEPARATE R session per folder:
devtools::load_all(".")             # first load downloads this vintage's bgej (~90 MB)
EJAM:::ejamdata_required_tag()      # confirm: "v3.2024.0" (or v3.2023.0, ...)
attr(blockgroupstats, "acs_version")# confirm: "2020-2024" (or "2019-2023", ...)
out_new <- ejamit(testpoints_10, radius = 1)   # run the SAME input in each session
# compare out_new$results_overall across the two sessions / vintages
```

If instead you stay in **one checkout and switch branches**, the downloaded
`data/bgej.arrow` is gitignored and persists across `git switch`, while the version
marker `data/ejamdata_version.txt` is tracked and changes with the branch. To be
sure you are testing the right vintage's EJ indexes, remove the cached file so it is
re-downloaded for the new vintage:

```{r, eval=FALSE}
# after: git switch ACS2023
file.remove("data/bgej.arrow")   # force re-download of the v3.2023.0 bgej
devtools::load_all(".")
```

As a safeguard, `dataload_dynamic_validate_bgej()` compares the loaded `bgej` to the
package's `blockgroupstats` and drops it if they do not match, so a stale `bgej` is
rejected rather than silently used.

## Release Checklist

Before releasing a new annual dataset build:

1. Run the pipeline with the intended ACS year and storage backend.
2. Confirm whether `bg_envirodata` and `bg_extra_indicators` are updated or
   provisional.
3. Review `pipeline_run_manifest.csv`, including package version, Git SHA, ACS
   vintage, run settings, and provisional-input flags.
4. Review `pipeline_validation_summary.csv`.
5. Review row counts, missingness, ranges, and joins for final tables,
   including the `bg_geodata` area fields.
6. Review `ejscreen_export_schema_report.csv`.
7. Replace package `.rda` datasets only after the stage files are accepted.
8. Update dataset metadata and documentation.
9. Rebuild documentation and pkgdown.
10. Reinstall the package and run focused tests plus package checks.
11. Publish large artifacts through the chosen storage path, such as S3 or the
    data repository release process.

The general release and large-data publication steps are covered in [Updating
and Managing the Datasets Used by EJAM](dev-update-datasets.html) and [Updating
the Package as a New Release](dev-update-package.html).
