---
title: "Defaults and Custom Settings for the Web App"
description: "Defaults and Custom Settings for the Web App"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Defaults and Custom Settings for the Web App}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 100
---

```{r developernote, eval=FALSE, echo= FALSE, include = FALSE}
#  *>>>>>>>>>> Developer note: vignettes need to be tested/edited/rebuilt regularly <<<<<<<<<<<*
#    - **See ?pkgdown::build_site** and script in EJAM/data-raw/- EJAM uses the pkgdown R package to build help and articles/ vignettes as web pages
```

```{r SETUP_default_eval_or_not, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(eval = FALSE)
# https://r-pkgs.org/vignettes.html
```

```{r libraryEJAM, eval = TRUE, echo= FALSE, include= FALSE}
# rm(list = ls()); golem::detach_all_attached(); devtools::load_all()

if (!exists("blockgroupstats")) {library(EJAM)} # use installed version only if pkg not yet attached

# dataload_dynamic('all') # varnames = all  currently means all defined by .arrow_ds_names
# 
# indexblocks()
```


# NOTES ON GLOBAL DEFAULTS, SHINY INPUTS, BOOKMARKS, AND FUNCTION PARAMETERS

This document provides technical notes on how defaults and custom settings are defined for the EJAM web app.

## HOW SETTINGS GET PASSED TO THE WEB APP

Many defaults are defined in files like `global_defaults_*.R`. They can be changed there, but also can be passed to `ejamapp()` aka `run_app()` as parameters, to override those global default settings for the duration of the app. Using `ejamapp()` parameters allows experienced analysts to customize how the web app works without having to edit any files. For example, one can change default settings for the radius, the type of sites to analyze, name of the analysis, etc., or to override caps such as limits on file upload size or number of points:

```{r run-app1, eval=FALSE}
ejamapp(
  radius_default=3.1,
  default_max_miles=31,
  default_max_mb_upload=100
)
```

These custom parameters are "advanced" features for knowledgeable users, and not all the possible changes to these have been tested,
so anyone using them should be careful to understand the details of how this work and confirm it is doing what is expected. 

*For example, the radius is defined like this:* 

1. if a user in RStudio launches the app with something like `ejamapp(radius_default=3.1)`, the setting `radius_default` is defined, and will override what it says in `global_defaults_shiny.R`

2. the files `global_defaults_*.R` are where the settings like `radius_default` are defined and stored as appOptions.

3. that option/setting is checked in the advanced tab UI, to provide the initially selected value of `input$radius_default` in a bookmarkable input control found in the advanced tab. In this case the option and input used the same exact name, "radius_default" but that is not always the case for other settings.

4. in the advanced tab, `input$radius_default` can then be adjusted by a user or just left alone. It provides the initially selected value of `input$radius_now` (which is for the slider UI, but happens to be defined in the server not ui in this case).

5. the radius slider bar will change the value of `input$radius_now` when a user moves the radius slider bar in the UI.

6. the value of `input$radius_now` is what the app uses in calculations when the start analysis button is clicked.

*A setting can be named differently in three places:*

Sometimes there are two or three separate names that affect the same
setting. You only need to know about the first one unless
you are editing URL-encoded bookmarks or editing source code!

1) The global default constant that is the initial value of x,
and is normally defined in a `global_defaults_*.R` file installed with the EJAM package, but
also can be provided as a parameter to `ejamapp()`.

2) for the subset of settings that can be adjusted in the Advanced tab (not all can),
there is a second variable: the input$ one can adjust in that tab.

3) the name of the input$ finally used in the main part of the web app.

The first of these three cannot be saved as a bookmark, but the second or third
can be bookmarked since it is a shiny app input$ variable.

*Steps in EJAM shiny app launch, and the use of params/ defaults/ advanced tab/ bookmarked inputs/etc.:*

### app.R and isPublic=T

If the app is on a server or one uses the RStudio button "Run", app.R is used.
app.R is normally how the app is launched if it is a hosted shiny app on a server.
app.R mostly just does library(EJAM) and ejamapp(isPublic = TRUE).
isPublic is special because it is checked by the global_defaults_shiny_public.R file
and some settings depend on isPublic being T or F.

### ejamapp() parameters

If the app is launched directly via ejamapp() by a user in RStudio console, it skips app.R
and a user can provide arguments as parameters in ejamapp(...). See `?ejamapp`

### global_defaults_*.R files

Many defaults/ settings used by the web app are defined in the global_defaults_ files
(global_defaults_package.R, global_defaults_shiny.R, global_defaults_shiny_public.R).
Some settings are needed even if the shiny app is never launched, so they are stored for the package 
as soon as the package is attached, in a list object called `global_defaults_package` in the global environment, e.g., `global_defaults_package$report_title`. (When the app is launched, those and many other global_defaults from the files `global_defaults_*.R`, including radius_default, will be stored in the shiny app).
Note the parameter isPublic is special because it is checked by the global_defaults_shiny_public.R file
and some settings depend on isPublic being T or F.

### any ejamapp() parameters will override global_defaults_*.R settings

Just before launching the app, `ejamapp()` calls `get_global_defaults_or_user_options()` which
gets the global defaults and then overrides/replaces them with
any user-specified versions of those options given as arguments in `ejamapp()`.

### app launch

`ejamapp()` launches the app -- it uses a function called `golem::with_golem_options()`, 
which defines the `app` object using `shiny::shinyApp()`, but then
stores all the global defaults and any user-provided parameters as `app$appOptions$golem_options`, 
just before launching the app. Those options can be retrieved in app or server later, 
via `EJAM:::global_or_param()` which relies on `golem::get_golem_options()`.

### global env?

Then app_ui() and app_server() can check for defaults/params using `EJAM:::global_or_param("radius_default")` 
which uses `golem::get_golem_options()` to check for a named option stored in `app$appOptions$golem_options`.
Usually these are used as initial/default values for input$ variables in the app.
However, note if an expected setting is not found (ie not set by user in ejamapp() and not set in any global_defaults_ file)
`global_or_param()` also looks in global env, as a last resort!! That might be a problem if ui or server looks for a
setting that we failed to provide in global_defaults files and that variable name happens to be in the global env !

### Advanced Tab

Then, SOME BUT NOT ALL global_defaults_ settings are in the Advanced Tab (in app_ui)
where they typically provide the default for an input$ but that default can be overridden
there by someone using the app's Advanced Tab,
so what ui or server ends up using could be an input$ that is adjustable in the Advanced Tab
and therefore can be bookmarked (since inputs can be bookmarked unless specifically excluded from that).

### web app reactives and interactive inputs

The shiny app uses some settings as initial values for
parameters that can get changed by the user via sliders, radio buttons, text boxes, dropdowns, etc.
and some selections/inputs can get changed by reactive observers, etc.

### Bookmarks (URL or server-based) for saved input$ state

If the app is launched by a bookmark (e.g.URL-encoded bookmark),
any inputs encoded in the URL will override the above and the app will change to
a state where the inputs are as specified by the bookmark.

Bookmarks have limitations though: 
Notably, uploaded files can only be bookmarked
if enableBookmarking="server" not enableBookmarking="url".
Also, bookmarks normally just save the state
of UI inputs, not other reactive values (unless the server code specifies otherwise by 
using the `shiny::onBookmark()` and `shiny::onRestore()` functions to make 
bookmarks store other information, beyond just inputs).
Some important global defaults affect the app but not by providing
initial values for inputs in the shiny app, so some useful information 
including some that is controlled by global defaults or user parameters in ejamapp()
cannot be bookmarked because it is stored in reactives
but not as shiny app inputs. 

--------------
--------------

## EXAMPLES OF USING CUSTOM SETTINGS 


### Launch local app (the R user's computer) with custom settings

```{r run-app, eval=FALSE}

# Type out the custom settings each time you use ejamapp()

ejamapp(
  default_standard_analysis_title="Custom NAICS Analysis",
  default_upload_dropdown="dropdown",
  default_selected_type_of_site_category="NAICS",
  default_naics_digits_shown="detailed",
  default_naics="562211",
  radius_default=3.1,
  default_ui_show_advanced_settings=1
)
```





<!--   not finished 


Functions to help launch with custom parameters

```{r run-app-with-params, eval=FALSE}

# Define helper functions -   *************** UNFINISHED DRAFT  -   *************** 

# draft function generator that gives you a custom version of ejamapp()
make_run_custom = function(params = NULL) {
  if (is.list(params)) {params <- unlist(params)}
  txt <- paste0(
    "ejamapp(", paste0(paste0(names(params), " = '", params, "'"), collapse=", "), ")"
  )
  myapp = function() {
    EJAM:::source_this_codetext(txt)  ### ***(((((((((((((( ACTUALLY THIS HAS TO BE EVALUATED BEFOREHAND - FIX THIS  ))))))))))))))
  }
  return(myapp)
}

# function to launch ejamapp() with parameters saved as vector or list
run_with_paramlist = function(params = NULL) {
  fn = make_run_custom(params)
  fn()
}
```

Launch using parameters that are saved in a named vector or list

```{r run-with-paramlist, eval=FALSE}
myparams <- c(
  default_standard_analysis_title="Custom NAICS Analysis",
  default_upload_dropdown="dropdown",
  default_selected_type_of_site_category="NAICS",
  default_naics_digits_shown="detailed",
  default_naics="562211",
  radius_default=3.1,
  default_ui_show_advanced_settings=1
)

run_with_paramlist(myparams)
```

Save a custom reusable function you can use to launch with preset parameters

```{r custom-app, eval=FALSE}

myparams <- c(
  default_standard_analysis_title="Custom NAICS Analysis",
  default_upload_dropdown="dropdown",
  default_selected_type_of_site_category="NAICS",
  default_naics_digits_shown="detailed",
  default_naics="562211",
  radius_default=3.1,
  default_ui_show_advanced_settings=1
)

# create your reusable customized app launcher
myapp = make_run_custom(myparams)
# save(myapp, file = "myapp.rda")

# use that later to launch app
# load("myapp.rda")
myapp()
```

-->




### Launch hosted app (online, URL) with custom settings

If the web app is running locally or hosted on a server,
launching it via URL-encoded bookmark can provide it any of the input$ variables as a starting state.

This approach cannot provide settings that are not input$ reactives, like some
of the defaults in global_defaults_ that are not turned into inputs in the app,
such as "app_logo_aboutpage" set in global_defaults_package.R and used in app_ui.R
as the src parameter for img().

Also note that url-encoded bookmarked inputs can launch the app
and pass various input$ settings to the app, which would override
the default values of those input$ settings provided by the global_defaults_*.R
and/or `ejamapp()` parameters handled by the utility function
get_global_defaults_or_user_options().

**Example that will launch app with a custom title and radius**

Note this tries to specify the NAICS method, but at least initially the draft code handling this could not specify the NAICS code itself or start analysis via URL bookmark:

Here is an example using the URL of one copy of the hosted app:

*** THIS EXAMPLE WILL WORK only with VERSION 2.32.6 or later:

Locally hosted app example

```{r, eval=FALSE}

# In 1 R session, launch locally with advanced tab disabled
myport <- 5432
ejamapp(isPublic=TRUE, options = list(port=myport))

# Then you can open a second R session and in that new R session
# paste/use the code below to relaunch the app
# with a parameter that enables the advanced tab:

params <- c('ui_show_advanced_settings=1')
urlbase <- paste0("http://127.0.0.1:", myport, "/?_inputs_&")
urlx <- URLencode(paste0(urlbase, paste0(params, collapse = "&")))
browseURL(urlx)

```


Using a custom URL to go to a web app with custom settings

```{r browseURL, eval=FALSE}
urlbase <- "https://ejam.policyinnovation.info/?_inputs_&"

myparams <- c(
  standard_analysis_title="Custom NAICS Analysis",
  default_ss_choose_method="dropdown", 
  ss_choose_method_drop="NAICS", 
  naics_digits_shown="detailed", 
  # ss_select_naics=313,  
  radius_default=3.1,
  ui_show_advanced_settings=1
)

params <- c('standard_analysis_title="Custom NAICS Analysis"',
            'ss_choose_method="dropdown"',   # like ejamapp( default_upload_dropdown="dropdown" )
            'ss_choose_method_drop="NAICS"', # see ?ejamapp()
            'naics_digits_shown="detailed"', # like ejamapp( default_naics_digits_shown = "detailed")
            # 'ss_select_naics=313',           # like ejamapp(default_naics=313)
            'radius_default=3.1',             # like ejamapp(radius_default = 3.1)
            'ui_show_advanced_settings=1')

urlx <- URLencode(paste0(urlbase, paste0(params, collapse = "&")))
browseURL(urlx)

# URL tried, with fips picked
# http://127.0.0.1:4200/?_inputs_&standard_analysis_title=%22Custom%20COUNTIES%20Analysis%22&pickermoduleid-all_regions_button=%22TRUE%22&pickermoduleid-all_states_button=%22TRUE%22&ui_show_advanced_settings=1&ss_choose_method=%22dropdown%22&fipspicker_done_button=2&ss_choose_method_drop=%22FIPS_PLACE%22&pickermoduleid-fips_type2pick=%22Counties%22&pickermoduleid-counties_picked=%2202110%22

```


--------------
--------------

## COMPILED LISTS OF ALL SETTINGS (FUNCTION PARAMETERS, GLOBAL_DEFAULTS_, AND SHINY INPUT$ VARS)

### > global_defaults_xyz - compiled full list

```{r global-d, eval=TRUE}
gdefs = EJAM:::get_global_defaults_or_user_options()
gdefnames = sort(unique(names(gdefs)))
cbind(global_defaults = gdefnames)
```

### > function params - compiled full list
```{r, eval=TRUE}
params = formals(ejamit) 
# params = c(formals(getblocksnearby), formals(getblocksnearbyviaQuadTree), formals(get_blockpoints_in_shape), formals(getblocksnearby_from_fips), formals(ejamit), formals(doaggregate))
paramnames = sort(unique(names(params)))
cbind(params = paramnames)
```

### > shiny inputs - compiled full list

Compilation of input$  variables used in the shiny app


<!--   needs updating   and  compare to  EJAM:::input_names_listing()    -->


```{r inputs, eval=FALSE}
# getwd() must be the source pkg root folder
getwd()
lines_with_inputnames = trimws(grep("input\\$", gsub("^(.*)(input\\$.*\\))(.*)", "\\2", readLines("./R/app_server.R") ), value=T))
x = gsub("^(.*)(input\\$.*)", "\\2", lines_with_inputnames)
x = gsub("]|)|,|}", " ", x)
x = gsub("(input\\$)([^ ]*)(.*)", "\\1\\2", x)
# cbind(x, lines_with_inputnames)
inputnames = sort(unique(x)); rm(x,lines_with_inputnames)
inputnames = grep("\\.$|\\$$", inputnames, value = TRUE, invert = TRUE)
# cbind(inputnames)
# cbind(part_of_a_named_list = inputnames[grepl("\\$.*\\$", inputnames)])
inputnames =  inputnames[!grepl("\\$.*\\$", inputnames)] # they all also showed up without the extra $xyz part

# What names are used in the shiny app input$ reactives?

cbind(shiny_inputs = inputnames)
inputnames = gsub("^input\\$", "", inputnames)
dput(inputnames)
```

```{r inputnames, eval=TRUE}

inputnames = c("add_naics_subcategories", "all_tabs", "allow_median_in_barplot_indicators", 
               "an_map_clusters",
               "an_thresh_comp2", "an_threshgroup1", "an_threshgroup2", "an_threshnames2", 
               "analysis_title", 
               "avoidorphans",
               "back_to_site_sel", "back_to_site_sel2", "bt_get_results",
               "calculate_ratios", "checkbox1", 
               "coauthor_emails", "coauthor_names", "conclusion1", "conclusion2", "conclusion3", 
               "Custom_title_for_bar_plot_of_indicators",
               
               "default_ss_choose_method", 
               
               "demog_high_at_what_share_of_sites", "demog_how_elevated", 
               "envt_high_at_what_share_of_sites", "envt_how_elevated",
               "epa_program_help",
               "extra_demog", "extratable_hide_missing_rows_for", 
               "extratable_show_ratios_in_report", "extratable_title", "extratable_title_top_row", 
               
               "facilities_studied", "facilities_studied_enter", "fips_help", 
               "fipspicker_done_button",
               "format1pager", "frs_help", "fundingsource", 
               "go", "in_areas_where_enter", "in_the_x_zone", "in_the_x_zone_enter", 
               "include_ejindexes", "latlon_help", "latlontypedin_submit_button", 
               
               "max_mb_upload", "max_miles", "maxradius",
               "max_pts_map", "max_pts_run", "max_pts_select", "max_pts_upload", "max_shapes_map",
               
               "naics_digits_shown", 
               "need_proximityscore", "ok2plot", "plotkind_1pager", 
               
               "radius_default", "radius_default_shapefile", "radius_now", "radius_units", 
               
               "results_tabs", "return_to_results",
               
               "rg_author_email", "rg_author_name",
               "rg_enter_miles", "rg_enter_sites", "rg_zonetype", "risks_are_x", 
               "shiny.testmode", 
               "show_data_preview", "show_ratios_in_report", "shp_help", "source_of_latlons", 
               
               "ss_choose_method", "ss_choose_method_drop", "ss_choose_method_upload", 
               "ss_select_mact", "ss_select_naics", "ss_select_program", "ss_select_sic", 
               "ss_upload_fips", "ss_upload_frs", "ss_upload_latlon", "ss_upload_program", 
               "ss_upload_shp", 
               
               "standard_analysis_title",
               "subgroups_type",
               
               "summ_bar_data", "summ_bar_ind", "summ_bar_stat", "summ_hist_bins", 
               "summ_hist_data", "summ_hist_distn", "summ_hist_ind",
               
               "testing", 
               "ui_hide_advanced_settings", "ui_show_advanced_settings", "y"
)
```

### > ALL OVERLAPS/ DIFFS (compiled func params, global defaults, shiny inputs)

```{r overlaps, echo=TRUE}

gdefnames_noprefix = gsub("default_", "", gdefnames)

cat("How many are named the same way in shiny inputs, global_defaults_, and function parameters?\n\n")

length(
  intersect(inputnames, gdefnames_noprefix) # about 28
  )
length(
  intersect(paramnames, gdefnames_noprefix)
  )
length(
  intersect(paramnames, inputnames)
  )
# in all 3 lists
print(
  intersect(paramnames, intersect(inputnames, gdefnames_noprefix))
)

cat("Unique names in shiny inputs, as named \n")
length(
  setdiff(inputnames, c(paramnames, gdefnames_noprefix))
)

cat("Unique names in global_defaults_ (not function parameters or inputs, as named) \n")
length(
  setdiff(gdefnames_noprefix, c(paramnames, inputnames))
)

cat("Unique names in function parameters \n")
length(
  setdiff(paramnames, c(gdefnames_noprefix, inputnames))
)

```

--------------
--------------

## ANNOTATED LISTS OF SOME SETTINGS


### > function params - grouped by function ####

Function parameters whose defaults are or could be included in advanced tab, and allowed by `ejamapp()`.

```{r func-params-by-func, eval=TRUE}

EJAM:::args2(getblocksnearby)

EJAM:::args2(getblocksnearbyviaQuadTree)

EJAM:::args2(get_blockpoints_in_shape)

EJAM:::args2(getblocksnearby_from_fips)

# sitepoints can be a param in ejamapp()
# shapefile  can be a param in ejamapp()
# fips

# radius_donut_lower_edge
# quiet = FALSE

```

### > function params - annotated list of most

```{r funcparams-annotated, eval=TRUE}
# > cbind(formals(ejamit))
EJAM:::args2(ejamit)

# sitepoints               NULL            -- Now can be passed to ejamapp()
# shapefile                NULL           ----   shapefile now can be  passed to ejamapp()
# fips                     NULL            ----   fips  will be enabled as param

# radius                   3   for getblocksnearby(), shiny radius_default is in global_defaults_
# radius_donut_lower_edge  0         na
# maxradius                31.07     for getblocksnearby(), shiny default is set in global
# avoidorphans             FALSE     for getblocksnearby(), shiny default is set in global
# quadtree                 NULL      na

# countcols                NULL      for doaggregate() ** shiny default NOT specified here
# popmeancols              NULL      for doaggregate() ** shiny default NOT specified here
# calculatedcols           NULL      for doaggregate() ** shiny default NOT specified here
# calctype_maxbg           NULL      for doaggregate() ** shiny default NOT specified here
# calctype_minbg           NULL      for doaggregate() ** shiny default NOT specified here

# subgroups_type           "nh"      for doaggregate(), shiny default is set in global
# include_ejindexes        TRUE      for doaggregate(), shiny default is set in global
# calculate_ratios         TRUE      for doaggregate(), shiny default is set in global
# extra_demog              TRUE      for doaggregate(), shiny default is set in global
# need_proximityscore      FALSE     for doaggregate(), shiny default is set in global
# infer_sitepoints         FALSE     for doaggregate() ** shiny default NOT specified here
# need_blockwt             TRUE      for getblocksnearby_from_fips(), shiny default not specified here, but by function defaults

# thresholds          expression  for batch.summarize() - set in global_ but named differently
# threshnames         expression  for batch.summarize() - set in global_ but named differently
# threshgroups        expression  for batch.summarize() - set in global_ but named differently

# updateProgress           NULL      ** shiny default NOT specified here
# updateProgress_getblocks NULL      ** shiny default NOT specified here
# in_shiny                 FALSE     for ejamit(), build_community_report(), and related functions
# quiet                    TRUE      for getblocksnearby() and batch.summarize(), na

# testing                  FALSE    for doaggregate(), shiny default is set in global
# silentinteractive        FALSE    for doaggregate() etc. na
# called_by_ejamit         TRUE     for doaggregate(), na
# showdrinkingwater        TRUE     for doaggregate(), na
# showpctowned             TRUE     for doaggregate(), na
#   download_fips_bounds_to_calc_areas FALSE
# ...                      ?   for getblocksnearby(), like use_unadjusted_distance ** shiny default NOT specified here
```


### > shiny inputs - categorized (partial) ####

EJAM shiny app inputs that can be URL-encoded in a bookmark to reopen the app with those settings
Note the URL cuts off  at about 4096 characters

#### inputs: tabs and nav and help buttons

```{r, eval=FALSE}

all_tabs="See Results"   # or # all_tabs="Advanced Settings"
results_tabs="Community Report"
details_subtabs="Site-by-Site Table"

show_advanced_settings="TRUE" # ??
can_show_advanced_settings="TRUE"

# buttons:
ui_show_advanced_settings=1 #button
ui_hide_advanced_settings=0 #button
show_data_preview=0
back_to_site_sel2=0
return_to_results=0
latlon_help=0
shp_help=0
fips_help=0
frs_help=0
epa_program_help=0
```

#### inputs: start analysis button

```{r, eval=FALSE}
bt_get_results=0
```

#### inputs: site selection (buttons/dropdowns)

```{r, eval=FALSE}

default_ss_choose_method="upload" # or "dropdown"
ss_choose_method="upload"         # or "dropdown"
ss_choose_method_upload="latlon"   # see global_defaults_*.R for options
ss_choose_method_drop="FIPS_PLACE" # see global_defaults_*.R for options
ss_upload_shp=null
ss_upload_fips=null
ss_upload_frs=null
ss_upload_program=null
ss_select_program="CAMDBS"
ss_select_naics=null
ss_select_sic=null
ss_select_mact="AA"
naics_digits_shown="basic"
add_naics_subcategories="TRUE"

shinyjs-resettable-ss_upload_latlon={}
shinyjs-resettable-ss_upload_shp={}
shinyjs-resettable-ss_upload_frs={}
shinyjs-resettable-ss_upload_program={}
shinyjs-resettable-ss_upload_fips={}

pickermoduleid-fips_type2pick="Counties"
pickermoduleid-states_picked=null
pickermoduleid-counties_picked=null
pickermoduleid-cities_picked=null
pickermoduleid-regions_picked=["1","2","3","4","5","6","7","8","9","10"]
pickermoduleid-all_regions_button=true
pickermoduleid-all_states_button=true
pickermoduleid-all_counties_button=false
pickermoduleid-all_cities_button=false
pickermoduleid-reset_button=0
fipspicker_done_button=0
```

#### inputs: radius and caps

```{r, eval=FALSE}
radius_now=3.1  # input from slider but not the way to bookmark /url-encode the radius
radius_default=3.1  # works as parameter to ejamapp() or as bookmarked input$ 
radius_default_shapefile=0
max_miles=10
maxradius=31.06856
max_pts_upload=5000
max_pts_select=5000
max_pts_map=5000
max_pts_showtable=1000
max_pts_run=10000
max_shapes_map=159
max_mb_upload=50
```

#### inputs: more ejamit() parameters, summary report and excel output settings, etc.

```{r, eval=FALSE}
standard_analysis_title="CUSTOM Analysis" # shown on webpage of app and on reports
analysis_title="testname"
Custom_title_for_bar_plot_of_indicators=""
include_ejindexes="TRUE"
calculate_ratios=true
include_averages=true
show_ratios_in_report="TRUE"
extratable_show_ratios_in_report="TRUE"
more3="a"
avoidorphans="FALSE"
need_proximityscore="FALSE"
plotkind_1pager="bar"
format1pager="html"
ok2plot=true
allow_median_in_barplot_indicators="FALSE"
summ_hist_distn="People" # histograms
summ_hist_data="pctile"
summ_hist_bins=10

extra_demog="TRUE"
include_extraindicators=true
extratable_title=""
extratable_title_top_row="ADDITIONAL INFORMATION"
extratable_list_of_sections="pctpoor"
# extratable_hide_missing_rows_for=["pcthisp","pctnhba","pctnhaa","pctnhaiana","pctnhnhpia","pctnhotheralone","pctnhmulti","pctnhwa","pctlan_ie","pctlan_api","pctlan_other","pctlan_english","pctlan_spanish","pctlan_french",
# "pctlan_rus_pol_slav","p_chinese","p_korean","pctlan_other_ie","pctlan_vietnamese","pctlan_other_asian","pctlan_arabic","pctlan_nonenglish",
# "pctspanish_li","pctie_li","pctapi_li","pctother_li","pctmale","pctfemale","pctdisability","lowlifex","rateheartdisease","rateasthma","ratecancer","pctunder5","pctunder18","pctover64","occupiedunits","lifexyears",
# "percapincome","pctownedunits","pctpoor","sitecount_avg","sitecount_unique","sitecount_max","distance_min_avgperson","distance_min","count.NPL","count.TSDF","num_waterdis","num_airpoll","num_brownfield","num_tri","num_school",
# "num_hospital","num_church","yesno_tribal","yesno_airnonatt","yesno_impwaters","yesno_cejstdis","yesno_iradis","pctflood","pctfire","pctfire30","pctflood30","yesno_houseburden","yesno_transdis","yesno_fooddesert","pctnobroadband",
# "pctnohealthinsurance","pop","nonmins","age25up","hhlds","unemployedbase","pre1960","builtunits","povknownratio"]

# thresholds - counting how many of a certain indicator type are at/above some threshold
# see help docs on ejamapp() for examples of using these (but they are named a bit differently as parameters there)
an_thresh_comp1=80
an_thresh_comp2=80
an_threshgroup1="EJ-US-or-ST"
an_threshgroup2="Supp-US-or-ST"
an_threshnames1=["pctile.EJ.DISPARITY.dpm.eo","pctile.EJ.DISPARITY.drinking.eo","pctile.EJ.DISPARITY.no2.eo","pctile.EJ.DISPARITY.o3.eo","pctile.EJ.DISPARITY.pctpre1960.eo","pctile.EJ.DISPARITY.pm.eo","pctile.EJ.DISPARITY.proximity.npdes.eo",
                 "pctile.EJ.DISPARITY.proximity.npl.eo","pctile.EJ.DISPARITY.proximity.rmp.eo","pctile.EJ.DISPARITY.proximity.tsdf.eo","pctile.EJ.DISPARITY.rsei.eo","pctile.EJ.DISPARITY.traffic.score.eo","pctile.EJ.DISPARITY.ust.eo",
                 "state.pctile.EJ.DISPARITY.dpm.eo","state.pctile.EJ.DISPARITY.drinking.eo","state.pctile.EJ.DISPARITY.no2.eo","state.pctile.EJ.DISPARITY.o3.eo","state.pctile.EJ.DISPARITY.pctpre1960.eo","state.pctile.EJ.DISPARITY.pm.eo","state.pctile.EJ.DISPARITY.proximity.npdes.eo",
                 "state.pctile.EJ.DISPARITY.proximity.npl.eo","state.pctile.EJ.DISPARITY.proximity.rmp.eo","state.pctile.EJ.DISPARITY.proximity.tsdf.eo","state.pctile.EJ.DISPARITY.rsei.eo","state.pctile.EJ.DISPARITY.traffic.score.eo","state.pctile.EJ.DISPARITY.ust.eo"]
an_threshnames2=["pctile.EJ.DISPARITY.dpm.supp","pctile.EJ.DISPARITY.drinking.supp","pctile.EJ.DISPARITY.no2.supp","pctile.EJ.DISPARITY.o3.supp","pctile.EJ.DISPARITY.pctpre1960.supp","pctile.EJ.DISPARITY.pm.supp","pctile.EJ.DISPARITY.proximity.npdes.supp",
                 "pctile.EJ.DISPARITY.proximity.npl.supp","pctile.EJ.DISPARITY.proximity.rmp.supp","pctile.EJ.DISPARITY.proximity.tsdf.supp","pctile.EJ.DISPARITY.rsei.supp","pctile.EJ.DISPARITY.traffic.score.supp","pctile.EJ.DISPARITY.ust.supp","state.pctile.EJ.DISPARITY.dpm.supp",
                 "state.pctile.EJ.DISPARITY.drinking.supp","state.pctile.EJ.DISPARITY.no2.supp","state.pctile.EJ.DISPARITY.o3.supp","state.pctile.EJ.DISPARITY.pctpre1960.supp","state.pctile.EJ.DISPARITY.pm.supp","state.pctile.EJ.DISPARITY.proximity.npdes.supp",
                 "state.pctile.EJ.DISPARITY.proximity.npl.supp","state.pctile.EJ.DISPARITY.proximity.rmp.supp","state.pctile.EJ.DISPARITY.proximity.tsdf.supp","state.pctile.EJ.DISPARITY.rsei.supp","state.pctile.EJ.DISPARITY.traffic.score.supp","state.pctile.EJ.DISPARITY.ust.supp"]

```

#### inputs: maps

```{r, eval=FALSE}
an_leaf_map_zoom=5
an_leaf_map_bounds={"north":45.61403741135093,"east":-68.02734375000001,"south":28.18824364185031,"west":-123.2666015625}
an_leaf_map_center={"lng":-95.64786455000001,"lat":37.40467238129763}
an_leaf_map_groups="circles"
circleweight_in=4
```

#### inputs: test mode

```{r, eval=FALSE}
testing="FALSE"
shiny.testmode="FALSE"
print_uploaded_points_to_log=true
```

#### inputs: draft / reserved for a long word/pdf report

```{r, eval=FALSE}

# authors

rg_author_name="FirstName LastName"
rg_author_email="author@email.org"
rg_add_coauthors=false
coauthor_names=""
coauthor_emails=""
fundingsource=""

# methods & data

acs_version=""
ejscreen_version=""
subgroups_type="nh"

## sites selected for analysis

prefix_filenames=""
in_the_x_zone="area"
facilities_studied="rule"
facilities_studied_enter=""
rg_enter_sites="facilities in the _____ source category"
source_of_latlons=""
in_areas_where_enter=""
in_the_x_zone_enter="in "
rg_zonetype="zone_is_named_x"
within_x_miles_of="near the"
in_areas_where="in areas with"
risks_are_x="risk is at or above 1 per million (lifetime individual cancer risk due to inhalation of air toxics from this source category)"

# results

demog_how_elevated=""
envt_how_elevated=""
demog_high_at_what_share_of_sites="some of these sites, just as it varies nationwide"
envt_high_at_what_share_of_sites="some of these sites, just as it varies nationwide"
conclusion1=""
conclusion2=""
conclusion3=""

```

--------------
--------------
