---
title: "Updating the Package as a New Release"
description: "Steps for incrementing package version and GitHub releases"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Updating the Package as a New Release}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---


```{r, eval=TRUE, echo=FALSE, include=FALSE}

code_reponame_with_owner <- EJAM:::repo_from_desc()           # "OWNER/REPONAME"
code_reponame <- gsub(".*/", "", code_reponame_with_owner)    #       "REPONAME"
code_reponame_url <- EJAM:::repo_from_desc(get_full_url = T)   
# "https://github.com/OWNER/REPONAME"

docs_reponame_with_owner <- EJAM:::repo_from_desc('github.io')# "OWNER/REPONAME"
docs_reponame <- gsub(".*/", "", docs_reponame_with_owner)    #       "REPONAME"
docs_reponame_url <- EJAM:::repo_from_desc('github.io', get_full_url = T) 
# "https://OWNER.github.io/REPONAME"

data_reponame_with_owner <- desc::desc(file = system.file(
  "DESCRIPTION", package = "EJAM"))$get("ejam_data_repo")     # "OWNER/REPONAME"
data_reponame <- gsub(".*/", "", data_reponame_with_owner)    #       "REPONAME"
data_reponame_url <- paste0("https://github.com/", data_reponame_with_owner) 
# "https://github.com/OWNER/REPONAME"
```

This document outlines the general development process for updating EJAM (for a new release of the package) or relocating it to a different repository (repo name and/or github user name), or changing the full name of the tool. Also see the separate vignettes on [updating datasets](dev-update-datasets.html) and [updating documentation](dev-update-documentation.html).

## If relevant, carefully change owners and names of repositories etc.

- The full name/title of the web tool is used in several places and all of those should be reading the full name from where it is defined. The app title is stored in the DESCRIPTION file, and is available for vignettes, functions, etc. directly via `as.vector(desc::desc_get("Title"))`. After the package is attached, the name as potentially modified via global_defaults_package.R or parameters to `ejamapp()` is available as `EJAM:::global_or_param(".app_title")`

- The name of the *package* (e.g., "`r code_reponame`") is assumed to be the same as the name of the github *repository*, and there are many places where the package name is mentioned in vignettes/ articles, etc. Changing just the repo name is feasible, as described here, but you would have to be careful about this distinction. You would have to globally find in files all hardcoded references to the old names, urls, etc. since there still are places where it does not check in DESCRIPTION for the repo name or assumes repo and package names are identical.

If the package were renamed -- or especially if a new name were used for the repository or the owner of the repository changed -- several files would need to be updated for everything to work correctly. This is also true of the data repository.

- The name of the github repository storing the package code should be recorded as part of the URL parameter in the DESCRIPTION file in the root folder of the source package.

- The name of the repo (or other domain) where webpages of package documentation are published also should be recorded as part of the URL parameter in the DESCRIPTION file in the root folder of the source package. The website providing documentation of the R package is created via code in `EJAM/data-raw/datacreate_0_UPDATE_ALL_DOCUMENTATION_pkgdown.R`, relying on the R package `{pkgdown}` and hosted on github pages (or other domain).

- The name of the repo where datasets are stored should be recorded as the ejam_data_repo parameter in the DESCRIPTION file in the root folder of the source package.


## Update the package release

-   Update EJAM version number in *DESCRIPTION* in development branch (e.g. to 2.4.1) (& tying it to EJScreen version numbers if relevant). Search in files globally too, as there still are places where the version number is not read from DESCRIPTION, like the files *_pkgdown.yml* and *golem-config.yml* and *CITATION.cff* where the version must be updated too!

-   Update `NEWS.md` in development branch listing changes made. Use the numbering x.y.z (same numbering as was put in DESCRIPTION).

-   Merge development into main branch

-   Update the EJAM release/ create a new release on github.com, using the `NEWS.md` as the changelog, tagging to the main branch. 

## Test the Installation Process

To ensure that, after changes are made, the package can still be installed by users with various operating systems and versions of R, a workflow file, `.github/workflows/installation_test.yaml`, is triggered by any pushes to the `r code_reponame_with_owner` main branch. This file tests installation with the following matrix of options:

1.  **OS**: Latest Ubuntu, Windows, macOS

2.  **R version**: 4.3 and 4.4

3.  **Install Method**: `remotes::install_url()` and `remotes::install_github()`

But also note users may need to set up a personal access token (PAT) if they want to install from github - [see article on installing for info on setting up a PAT](installing.html).

### Update the Test Installation Script

You may want to update this test script in the future for a couple reasons:

-   Remove older versions of R to test on
-   Add new system libraries required by newly added packages


## obsolete step: Make updates public before a new release (from when a nonpublic repo was used for development at EPA)

```{r, echo=FALSE, eval=TRUE, include=FALSE}
# helper to get the owner and name of the code repo and the documentation repo as specified in the DESCRIPTION file
owner_repo <- EJAM:::repo_from_desc()      # had been "USEPA/EJAM-open"
repo <- gsub(".*/", "", owner_repo)        # had been       "EJAM-open"

```
If the development repo being used is internal or private, not public, then to make the package installable, the updates need to be pushed to a public repo. These would be the steps:

1. If not already done, add `r owner_repo` (or the appropriate public repo) as a remote for the internal repo main branch:

`    git remote add `r repo` git@github.com:`r owner_repo`.git`

2.  Push to the main branch of that public repo:

`    git push `r repo` main:main`

3.  If you want to squash commits

    a.  If you want to add a new squash commit, do this from main in the internal repo:

`    git checkout -b squash-temp`
`    git reset --soft #SHA OF PREVIOUS MAIN COMMIT`
`    git commit -m "Creating squashed commit"`
`    git push --force `r repo` squash-temp:main`

    b.  If you just want `r repo` to end with a single squash commit, do this from main in the internal repo:

`    git checkout –orphan squash-temp`
`    git add .`
`    git commit -m "Publishing to public repo"`
`    git push --force `r repo` squash-temp:main`
`    git -M main`

4.  Update `r repo` release using a similar process as updating an internal repo release.
