---
title: "Updating Documentation"
description: "Steps for updating package documentation and the pkgdown site"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Updating Documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

```{r, eval=TRUE, echo=FALSE, include=FALSE}

code_reponame_url        <- EJAM::url_package(type = "code", get_full_url = TRUE)  # "https://github.com/OWNER/REPONAME"
code_reponame_with_owner <- EJAM::url_package(type = "code", get_full_url = FALSE) # "OWNER/REPONAME"
code_reponame <- gsub(".*/", "", code_reponame_with_owner)    #       "REPONAME"

data_reponame_url        <- EJAM::url_package(type = "data", get_full_url = TRUE)  # "https://github.com/OWNER/REPONAME"
data_reponame_with_owner <- EJAM::url_package(type = "data", get_full_url = FALSE) # "OWNER/REPONAME"
data_reponame <- gsub(".*/", "", data_reponame_with_owner)    #       "REPONAME"

docs_reponame_url        <- EJAM::url_package(type = "docs", get_full_url = TRUE)  # "https://OWNER.github.io/REPONAME"
# only the full URL is useful in this case

```

This document describes how to update the documentation that supports the
package, including R help pages, dataset documentation, articles, app help
materials, and the pkgdown website.

## Updating package documentation

The R package is documented in web pages and in standard R help files viewable
in RStudio. Documentation generally needs to be updated before a new package
release. Updates may be needed in the following places:

1.  [Package: overall package-wide information and documents]

2.  [Functions: reference docs explaining the R functions]

3.  [Datasets: reference docs explaining the datasets] (in EJAM/R/data\_\*.R
    files)

4.  [Articles: vignettes helping R users with the package] (in
    `EJAM/vignettes/*.Rmd`)

5.  [User Guide: help for people using the web app] (in `EJAM/inst/app/www/` as
    a PDF file)

6.  [Spell-checking package-wide]

7.  [Website: web pages providing all of the above in HTML format], via the
    `{pkgdown}` package

Each is explained below:

### Package: overall package-wide information and documents

-   Package-wide documentation files can be edited directly and include the
    following: `DESCRIPTION`, `README.Rmd` (for the README, **edit
    `README.Rmd`, not the generated `README.md` file**),
    `R/EJAM-package.R` (for `?EJAM-package`),
    `NEWS.md` used to make [NEWS](../news/index.html), 
    `CONTRIBUTING.md` used to make [CONTRIBUTING](../CONTRIBUTING.html),
    [LICENSE](../LICENSE.html) based on `inst/LICENSE.md`,
    [CITATION](../authors.html#citation) based on `inst/CITATION`, 
    CITATION.cff, and others.

-   Some of these are used by and shown by the GitHub repository, notably
    the `README.md` that is generated from the `README.Rmd` file and appears on
    the GitHub repository as
    [README](`r paste0(code_reponame_url, "?tab=readme-ov-file#readme")`)
    and in the pkgdown site as [README](../index.html), and the `DESCRIPTION`
    file that stores the package version number, repository URLs, and related
    package metadata.

### Functions: reference docs explaining the R functions

-   `{roxygen2}` package tags define the documentation in the `R/*.R` files that
    contain the source code and documentation of functions and datasets.

-   While many functions or objects have their own `.R` files, some `.R` files
    contain code and documentation for multiple functions, so not every function
    has a file of its own.

-   Some utility/helper functions, mostly not exported, are stored in files
    prefixed with `utils_`, such as `R/utils_indexblocks.R` documenting `indexblocks()`
    (but not all helper functions are in files named that way).

-   It can be useful to check generated help files with
    `tools::checkDocFiles(".")` after running `devtools::document()`.

### Datasets: reference docs explaining the datasets

-   All datasets are documented in files named with a prefix of `data_`,
    such as `R/data_blockgroupstats.R` which documents the `?blockgroupstats`
    dataset.

-   Most dataset documentation is created or updated alongside the code that
    creates or updates those datasets. For annual EJScreen-style datasets, use
    the staged pipeline described in [Updating EJScreen Datasets Annually (via
    the Pipeline)](dev-update-ejscreen-datasets-yearly.html). For other
    package datasets, see [Updating and Managing the Datasets Used by
    EJAM](dev-update-datasets.html) and the relevant `data-raw/datacreate_*.R`
    scripts.

-   Many `R/*.R` files for datasets are created automatically, not edited by
    hand, through the `dataset_documenter()` function as used in the scripts
    within `data-raw/datacreate_*.R`, while others are updated by editing the `.R`
    file directly.

### Articles: vignettes helping R users with the package

-   Articles (called vignettes here) are written and edited as `.Rmd` files in
    the `vignettes` folder. You can test one using
    `pkgdown::build_article()` before rebuilding the whole pkgdown site.

-   New articles or renamed files also require edits to `_pkgdown.yml`.

### User Guide: help for people using the web app

-   If you change the UI, new screenshots may be needed for a new version of
    the User Guide. The User Guide has historically been stored as a PDF
    document built from a Word document.

### Spell-checking package-wide

-   To run a spell check and review terms flagged as possible problems:

```{r, eval=FALSE}
    x <- spelling::spell_check_package()
    x <- data.frame(
      frq = sapply(x$found, function(z) {length(unlist(z))}), 
      word = x$word)
    x <- x[order(x$frq), ]
    rownames(x) <- NULL
    x
```

-   Then use Ctrl-Shift-F to search all files for a flagged term and check
    whether it is actually a problem.

-   To update `inst/WORDLIST`, the list of words ignored by spell checking:

```{r, eval=FALSE}
    # 1. Add obvious dataset names documented to the WORDLIST
    datanames = gsub('\\.R$', '', gsub('^data_', '', dir('./R', pattern = '^data_.*R')))
    datanames = datanames[!grepl('aaaaaaaaaaaaa|xxxxxxxxxx', datanames)]
    # 2. Add names of functions to the WORDLIST
    funcnames = EJAM:::pkg_functions_and_data( 'EJAM' )$object
    wordlist = readLines('inst/WORDLIST')
    wordlist = sort(unique(union(wordlist, c(datanames, funcnames))))
    # Write the revised list of words to the file
    writeLines(wordlist, 'inst/WORDLIST')
```

### Website: web pages providing all of the above in HTML format

The web-based documentation pages should be updated right after any of the above
are edited or created. These pkgdown-based web pages should be updated by someone
who is managing the package, as follows:

-   The first step in updating those web pages is updating the YAML file. The
    web-based documentation pages are organized by a file called `_pkgdown.yml`
    -- see `usethis::edit_pkgdown_config()` -- which specifies the contents of
    the documentation web pages, such as the names of any vignette files to
    be shown in the list of articles, which functions are included in the
    reference pages index, and related settings. Any new, removed, or renamed
    functions, vignettes, datasets, or changes in what is exported may require
    edits to `_pkgdown.yml`.

-   The pkgdown-based web pages should be updated by someone who is managing the
    package, and they should use the `EJAM:::pkgdown_update()` function.
    Use `pkgdown_update()` so that the pages get rebuilt by the `{pkgdown}` R
    package, based on the above files and all other documentation from `.R` and
    `.md` files. For more
    information about using the `{pkgdown}` package to create documentation in
    the form of web pages, see [pkgdown
    articles](https://pkgdown.r-lib.org/articles/pkgdown.html). The
    documentation is created and updated using the `{roxygen2}` and `{pkgdown}` R
    packages, which read information from the `.R` files in the R folder, from
    `.Rmd` file vignettes (also called articles) in the vignettes folder, and from the
    `_pkgdown.yml` file in the root folder of the source package. The
    `{roxygen2}` package reads the `.R` files to create the `.Rd` files in the
    `man/` directory. The `{pkgdown}` package uses all those files as well as
    `DESCRIPTION` and others, to generate a website for the package, first
    created as files in the `docs/` folder, which are then published via GitHub
    Actions and the GitHub repository settings to be hosted on GitHub Pages.
    The packages `{devtools}` and `{usethis}` are also relevant.

-   The package's GitHub repository settings define which branch and folder
    hold the HTML files for deployment. See the GitHub Pages and pkgdown
    documentation for details.

-   The package's GitHub Actions can be used to deploy the web pages when
    triggered by an event such as a push to the main branch.

-   Depending on which folder you publish/deploy from, you might have to remove
    that folder from `.gitignore` in case it was listed there via some
    `{pkgdown}` or `{usethis}` function.

-   The URL where documentation is published at any given time should be
    recorded in the `URL` field of the `DESCRIPTION` file in the source package
    root. That URL can be read with `EJAM::url_package("docs")`, which
    currently returns `r docs_reponame_url`. Prior to mid-2025, documentation
    web pages had been on GitHub Pages at URLs related to the `USEPA/EJAM` and
    `USEPA/EJAM-open` repositories, but those pages may now be archived,
    unpublished, or otherwise obsolete.
