---
title: "Updating Documentation"
description: "Steps for updating package documentation and the pkgdown site"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Updating Documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

```{r, eval=TRUE, echo=FALSE, include=FALSE}

code_reponame_url        <- EJAM::url_package(type = "code", get_full_url = TRUE)  # "https://github.com/OWNER/REPONAME"
code_reponame_with_owner <- EJAM::url_package(type = "code", get_full_url = FALSE) # "OWNER/REPONAME"
code_reponame <- gsub(".*/", "", code_reponame_with_owner)    #       "REPONAME"

data_reponame_url        <- EJAM::url_package(type = "data", get_full_url = TRUE)  # "https://github.com/OWNER/REPONAME"
data_reponame_with_owner <- EJAM::url_package(type = "data", get_full_url = FALSE) # "OWNER/REPONAME"
data_reponame <- gsub(".*/", "", data_reponame_with_owner)    #       "REPONAME"

docs_reponame_url        <- EJAM::url_package(type = "docs", get_full_url = TRUE)  # "https://OWNER.github.io/REPONAME"
# only the full URL is useful in this case

```

This document describes how to update the documentation that supports the
package, including R help pages, dataset documentation, articles, app help
materials, and the pkgdown website.

## Updating package documentation

The R package is documented in web pages and in standard R help files viewable
in RStudio. Documentation generally needs to be updated before a new package
release. Updates may be needed in the following places:

1.  [Package: overall package-wide information and documents]

2.  [Functions: reference docs explaining the R functions]

3.  [Datasets: reference docs explaining the datasets] (in EJAM/R/data\_\*.R
    files)

4.  [Articles: vignettes helping R users with the package] (in
    `EJAM/vignettes/*.Rmd`)

5.  [User Guide: help for people using the web app] (in `EJAM/inst/app/www/` as
    a PDF file)

6.  [Spell-checking package-wide]

7.  [Website: web pages providing all of the above in HTML format], via the
    `{pkgdown}` package

Each is explained below:

### Package: overall package-wide information and documents

-   Package-wide documentation files can be edited directly and include the
    following: `DESCRIPTION`, `README.Rmd` (for the README, **edit
    `README.Rmd`, not the generated `README.md` file**),
    `R/EJAM-package.R` (for `?EJAM-package`),
    `NEWS.md` used to make [NEWS](../news/index.html), 
    `CONTRIBUTING.md` used to make [CONTRIBUTING](../CONTRIBUTING.html),
    [LICENSE](../LICENSE.html) based on `inst/LICENSE.md`,
    [CITATION](../authors.html#citation) based on `inst/CITATION`, 
    CITATION.cff, and others.

-   Some of these are used by and shown by the GitHub repository, notably
    the `README.md` that is generated from the `README.Rmd` file and appears on
    the GitHub repository as
    [README](`r paste0(code_reponame_url, "?tab=readme-ov-file#readme")`)
    and in the pkgdown site as [README](../index.html), and the `DESCRIPTION`
    file that stores the package version number, repository URLs, and related
    package metadata.

### Functions: reference docs explaining the R functions

-   `{roxygen2}` package tags define the documentation in the `R/*.R` files that
    contain the source code and documentation of functions and datasets.

-   While many functions or objects have their own `.R` files, some `.R` files
    contain code and documentation for multiple functions, so not every function
    has a file of its own.

-   Some utility/helper functions, mostly not exported, are stored in files
    prefixed with `utils_`, such as `R/utils_indexblocks.R` documenting `indexblocks()`
    (but not all helper functions are in files named that way).

-   It can be useful to check generated help files with
    `tools::checkDocFiles(".")` after running `roxygen2::roxygenise()`.

### Datasets: reference docs explaining the datasets

-   All datasets are documented in files named with a prefix of `data_`,
    such as `R/data_blockgroupstats.R` which documents the `?blockgroupstats`
    dataset.

-   Most dataset documentation is created or updated alongside the code that
    creates or updates those datasets. For annual EJScreen-style datasets, use
    the staged pipeline described in [Updating EJScreen Datasets Annually (via
    the Pipeline)](dev-update-ejscreen-datasets-yearly.html). For other
    package datasets, see [Updating and Managing the Datasets Used by
    EJAM](dev-update-datasets.html) and the relevant `data-raw/datacreate_*.R`
    scripts.

-   Many `R/*.R` files for datasets are created automatically, not edited by
    hand, through the `dataset_documenter()` function as used in the scripts
    within `data-raw/datacreate_*.R`, while others are updated by editing the `.R`
    file directly.

### Articles: vignettes helping R users with the package

-   Articles (called vignettes here) are written and edited as `.Rmd` files in
    the `vignettes` folder. Edit the `.Rmd` source file, not the generated HTML
    article.

-   Preview one article after editing it, before rebuilding the whole site:

```{r, eval=FALSE}
EJAM:::pkgdown_preview_article("installing")
EJAM:::pkgdown_preview_article("vignettes/dev-update-documentation.Rmd")
```

These preview helpers build into a temporary folder by default, so they do not
publish anything and do not require committing generated HTML. They also try to
use the Pandoc bundled with RStudio if a terminal R session cannot find Pandoc
on its own.

-   New articles or renamed files also require edits to `_pkgdown.yml`.

### User Guide: help for people using the web app

-   If you change the UI, new screenshots may be needed for a new version of
    the User Guide. The User Guide has historically been stored as a PDF
    document built from a Word document.

### Spell-checking package-wide

-   To run a spell check and review terms flagged as possible problems:

```{r, eval=FALSE}
    x <- spelling::spell_check_package()
    x <- data.frame(
      frq = sapply(x$found, function(z) {length(unlist(z))}), 
      word = x$word)
    x <- x[order(x$frq), ]
    rownames(x) <- NULL
    x
```

-   Then use Ctrl-Shift-F to search all files for a flagged term and check
    whether it is actually a problem.

-   To update `inst/WORDLIST`, the list of words ignored by spell checking:

```{r, eval=FALSE}
    # 1. Add obvious dataset names documented to the WORDLIST
    datanames = gsub('\\.R$', '', gsub('^data_', '', dir('./R', pattern = '^data_.*R')))
    datanames = datanames[!grepl('aaaaaaaaaaaaa|xxxxxxxxxx', datanames)]
    # 2. Add names of functions to the WORDLIST
    funcnames = EJAM:::pkg_functions_and_data( 'EJAM' )$object
    wordlist = readLines('inst/WORDLIST')
    wordlist = sort(unique(union(wordlist, c(datanames, funcnames))))
    # Write the revised list of words to the file
    writeLines(wordlist, 'inst/WORDLIST')
```

### Website: web pages providing all of the above in HTML format

The web-based documentation pages are built from source files by `{pkgdown}` and
published by GitHub Actions. The public site should be served from the root of
the `gh-pages` branch. Source branches such as `main`, `development`, and
release branches should not publish from `main/docs` and should not commit
generated `docs/` HTML.

In normal documentation work, edit and commit source files such as `.R`, `.Rmd`,
`README.Rmd`, `NEWS.md`, `DESCRIPTION`, `man/*.Rd` inputs generated from
roxygen comments, and `_pkgdown.yml`. Do not edit `docs/` or the `gh-pages`
branch by hand; those are generated deployment outputs.

#### Normal workflow for a vignette edit

1.  Edit the relevant `.Rmd` file in `vignettes/`.

2.  Preview just that article:

```{r, eval=FALSE}
EJAM:::pkgdown_preview_article("installing")
```

3.  If the article is new, renamed, removed, or moved in the navigation, update
    `_pkgdown.yml`.

4.  Commit and push the source files only. Do not add generated `docs/` files.

5.  Let GitHub Actions build and publish the public site after the source branch
    is pushed or merged into the branch that controls the desired documentation
    version.

-   The first step in updating those web pages is updating the YAML file. The
    web-based documentation pages are organized by a file called `_pkgdown.yml`
    -- see `usethis::edit_pkgdown_config()` -- which specifies the contents of
    the documentation web pages, such as the names of any vignette files to
    be shown in the list of articles, which functions are included in the
    reference pages index, and related settings. Any new, removed, or renamed
    functions, vignettes, datasets, or changes in what is exported may require
    edits to `_pkgdown.yml`.

-   For local checks, `EJAM:::pkgdown_update()` is still available as a
    maintainer utility. Treat it as a local preview/check workflow, not the
    publishing workflow. It can still be useful when you want broader local
    checks around roxygen documentation, README rendering, pkgdown site
    structure, or stale `.Rd` files.

-   For a faster local site preview without publishing:

```{r, eval=FALSE}
EJAM:::pkgdown_preview_site()
EJAM:::pkgdown_preview_site(pkgdown_dev_mode = "devel")
```

-   Public publishing is handled by `.github/workflows/pkgdown.yaml`. GitHub
    Pages should be configured to publish from the root of the `gh-pages`
    branch. The workflow builds the site in a runner and deploys generated HTML
    to `gh-pages`.

-   The normal publishing paths are:
    - `main` builds the current release documentation at the site root.
    - `development` builds development documentation under `dev/`.
    - version tags and GitHub releases build historical documentation under the
      tag name, such as `v3/`.
    - manual workflow dispatch can build a chosen ref into a chosen folder.

-   The workflow file must exist on the branch that GitHub Actions uses for the
    event being triggered. For normal public updates, make sure the workflow
    change has reached the default branch before expecting future `main` pushes
    to rebuild the root documentation site.

-   To publish historical docs for an older release, use a tag or release such
    as `v3` and deploy it to the same folder name. The resulting URLs should
    look like:

```{r, eval=FALSE}
https://public-environmental-data-partners.github.io/EJAM/
https://public-environmental-data-partners.github.io/EJAM/dev/
https://public-environmental-data-partners.github.io/EJAM/v3/
```

-   The documentation is created and updated using the `{roxygen2}` and
    `{pkgdown}` R packages, which read information from the `.R` files in the R
    folder, from `.Rmd` file vignettes (also called articles) in the `vignettes`
    folder, and from `_pkgdown.yml`. `{roxygen2}` creates `.Rd` files in `man/`;
    `{pkgdown}` creates HTML files for GitHub Pages.

-   The URL where documentation is published at any given time should be
    recorded in the standard `URL` field of the `DESCRIPTION` file in the
    source package root and in `Config/EJAM/url_ejamdocs`, which is the field
    read by `EJAM::url_package("docs")`. The helper currently returns
    `r docs_reponame_url`. Store stable alias and redirect URLs in
    `Config/EJAM/url_ejamdocs_alias` and
    `Config/EJAM/url_ejamdocs_redirect`. Prior to mid-2025, documentation web
    pages had been on GitHub Pages at URLs related to the `USEPA/EJAM` and
    `USEPA/EJAM-open` repositories, but those pages may now be archived,
    unpublished, or otherwise obsolete.
