---
title: "Deploying the Web App to AWS (Infrastructure & Procedure)"
description: "How the live EJAM Shiny app is hosted on AWS ECS Fargate, and the routine procedure for deploying updates"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Deploying the Web App to AWS (Infrastructure & Procedure)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 100
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
```

This article documents how the **live EJAM Shiny web app** is hosted on
**AWS ECS Fargate**, and the routine procedure for deploying updates. It is a
companion to [Deploying the Web App](dev-deploy-app.html), which covers the
higher-level *choices* (Docker/AWS vs. Posit Connect) and the app-configuration
options (`isPublic`, logos, titles); this article is the concrete, current
AWS procedure.

> **Not to be confused with the EJAM API.** This is the **Shiny app** (the
> interactive website), hosted on **AWS ECS Fargate**. The separate
> [EJAM REST API](dev-api.html) is hosted on **Google Cloud Run** and deployed
> from the [EJAM-API repository](https://github.com/Public-Environmental-Data-Partners/EJAM-API).

## Where the operational files live

The files that actually build and deploy the app are kept on the two
**deploy branches** (`dev-deploy` and `prod-deploy`) of the EJAM repository, **not
on `main`/`development`**. This keeps the branch-triggered deploy workflows and
AWS/Terraform config out of the package source (so they cannot misfire on a
`main` push and do not clutter the R package). This article is the human-readable
guide to those files; edit the files themselves on the deploy branches.

| File (on the deploy branches) | Purpose |
|---|---|
| [`ejam-infra/main.tf`](https://github.com/Public-Environmental-Data-Partners/EJAM/blob/dev-deploy/ejam-infra/main.tf) | Terraform — the full AWS infrastructure definition (VPC, ALB, ECR, ECS, IAM, logs) |
| [`ejam-infra/prod.tfvars`](https://github.com/Public-Environmental-Data-Partners/EJAM/blob/dev-deploy/ejam-infra/prod.tfvars), [`ejam-infra/dev.tfvars`](https://github.com/Public-Environmental-Data-Partners/EJAM/blob/dev-deploy/ejam-infra/dev.tfvars) | Per-environment Terraform variables (task size, retention, etc.) |
| `Dockerfile` | Container image build for the Shiny app. Chooses the EJAM version (`ARG EJAM_VERSION`) and the `ejamdata` release (`ARG EJAMDATA_VERSION`), and its `CMD` starts the app with `EJAM::ejamapp(isPublic = ...)` |
| `.dockerignore` | Files excluded from the Docker build context |
| `.github/workflows/deploy.yaml` | GitHub Actions — build + deploy to **prod** (on push to `prod-deploy`, or run manually) |
| `.github/workflows/deploy-dev.yaml` | GitHub Actions — build + deploy to **dev** (on push to `dev-deploy`, or run manually with an optional `ejam_version` input) |

The deploy branches also carry an older copy of the package source (including an
`app.R`, which is for Posit Connect), but the image does not use it: the
`Dockerfile` has no `COPY` step and installs EJAM from GitHub at the ref named by
`ARG EJAM_VERSION`. Only the files in the table above matter for a deploy.

## Architecture at a glance

| Layer | Technology |
|---|---|
| App | R Shiny (rocker/rstudio base image) |
| PDF generation | Google Chrome stable + chromote/pagedown |
| Containerization | Docker |
| Container registry | AWS ECR (shared repo `ejam`; the prod stack owns it, dev shares it) |
| Hosting | AWS ECS Fargate (region `us-east-1`) |
| Load balancing | AWS Application Load Balancer (ALB) |
| HTTPS / TLS | AWS ACM (DNS-validated) |
| Infrastructure-as-code | Terraform (S3 state backend, separate state per environment) |
| CI/CD | GitHub Actions |
| DNS | Squarespace (CNAME → ALB) |

```
  Each deploy branch triggers its own GitHub Actions workflow, which builds the
  image, pushes it to ECR, and updates that environment's ECS Fargate service:

    dev-deploy   ──push──► deploy-dev.yaml ──► ECR (ejam) ──► ECS Fargate: ejam-dev   ──► dev ALB
    prod-deploy  ──push──► deploy.yaml     ──► ECR (ejam) ──► ECS Fargate: ejam-prod  ──► prod ALB
                                               image tags:                                 │
                                               dev = dev-<sha>, prod = <sha>               ▼
                                                                          Squarespace CNAME ──► ejam.publicenvirodata.org

  Each image installs EJAM from GitHub at the ref pinned by ARG EJAM_VERSION in
  that branch's Dockerfile -- see "Branching and deploy model" below.
```

**URLs**

| Environment | URL |
|---|---|
| Production | <https://ejam.publicenvirodata.org> (Squarespace CNAME → prod ALB, ACM TLS) |
| Dev | the dev ALB DNS name printed by `terraform output` (an `…elb.amazonaws.com` address) |

**Task sizing** (set in the `.tfvars` files)

| | vCPU / memory | Task count | Logs |
|---|---|---|---|
| Prod | 2 vCPU / 7 GB | 2 | CloudWatch `/ecs/ejam-prod`, 30-day retention |
| Dev | 1 vCPU / 6 GB | 1 | CloudWatch `/ecs/ejam-dev`, 7-day retention |

The container serves the app on **port 2000** and a health-check endpoint on
**port 2001** (`app_port` / `health_check_port` in `ejam-infra/main.tf`, which
default to 2000/2001; the `Dockerfile` `EXPOSE`s `2000 2001`, and its `CMD` starts
a small health server on 2001 and then `EJAM::ejamapp(options = list(port = 2000))`).
The `CMD` also sets the app mode: dev runs `isPublic = FALSE` (the full app, so
testing exercises every feature) and prod runs `isPublic = TRUE` (the public app).

## Branching and deploy model

The deploy branches hold the **build files**, not the code that gets deployed.
Each branch's `Dockerfile` names the EJAM version to install, and the image
installs that version from GitHub:

```
dev-deploy/Dockerfile    ARG EJAM_VERSION=<tag, branch, or SHA>   ARG EJAMDATA_VERSION=vX.YYYY.Z
prod-deploy/Dockerfile   ARG EJAM_VERSION=<release tag>           ARG EJAMDATA_VERSION=vX.YYYY.Z
```

- **To change what is deployed, change those `ARG` lines** with a small pull
  request into the deploy branch. There is no need to merge `main` or
  `development` into a deploy branch; the package source on the branch is not
  installed.
- `ARG EJAM_VERSION` can be a release tag (e.g. `v3.2022.3`), a branch
  (e.g. `development`), or a commit SHA. **Production should always use a release
  tag.** Dev can use any ref, which lets you test a release candidate before it is
  tagged.
- `ARG EJAMDATA_VERSION` names the `ejamdata` release; see
  [Which EJAM version deploys](#which-ejam-version-deploys) below.
- **Merging a pull request into `dev-deploy` deploys dev; merging into
  `prod-deploy` deploys prod**, immediately (about 15–20 minutes). Each branch's
  workflow targets its own ECS cluster and service.
- The two `Dockerfile`s are meant to differ only in the pins and in `isPublic`
  (dev `FALSE`, prod `TRUE`). Make any other change (for example the font-cache
  line described [below](#which-ejam-version-deploys)) on **both** branches.
- Validating on dev and then setting the **same tag** on prod deploys the same EJAM
  code. The prod image is still a fresh build, though: it uses the current
  `rocker/rstudio:latest` base image and current CRAN packages, which may differ
  slightly from the dev image built earlier.
- Both `dev-deploy` and `prod-deploy` are **protected**: a PR is required (direct
  pushes are blocked for everyone, including admins); no approvals are required
  (a maintainer may merge their own PR).

**Running the dev deploy for another EJAM ref.** The dev workflow can also be
started by hand (GitHub → Actions → "Build & Deploy to ECS Fargate (dev)" → Run
workflow, or from the command line) with an `ejam_version` input, which overrides
`ARG EJAM_VERSION` for that one build without editing the branch:

```bash
gh workflow run "Build & Deploy to ECS Fargate (dev)" \
  --repo Public-Environmental-Data-Partners/EJAM --ref dev-deploy \
  -f ejam_version=development
```

Leave `ejam_version` blank to build the version pinned in the `Dockerfile`. The
dispatch **cannot override `EJAMDATA_VERSION`**: to test with a different data
release, change that pin in `dev-deploy` by pull request. The prod workflow can
also be run by hand, but it has no input; it rebuilds the pinned version.

> **Deploys don't queue.** Neither deploy workflow limits concurrent runs, so a
> deploy started by a merge and one started by hand (or two started by hand) run
> at the same time, each updating the same ECS service. Either one can end up
> live, and it may not be the one you meant. Start one deploy, wait for it to
> finish, and then confirm the version shown in the app footer. In particular,
> don't dispatch a dev deploy right after merging into `dev-deploy`, since the
> merge already started one.

## Routine deploy (GitHub Actions) — the normal path

1. **Release (or pick a candidate).** Deploy a release tag. To test a release
   candidate on dev before tagging, use a branch or commit SHA instead.
2. **Deploy to dev:** open a PR into `dev-deploy` that sets `ARG EJAM_VERSION`
   (and `ARG EJAMDATA_VERSION`, if the data release changed). On merge, GitHub
   Actions builds the image, pushes it to ECR, and updates the dev ECS service.
   For a quick test of an unreleased ref, you can dispatch the dev workflow with
   `ejam_version` instead (see above).
3. **Validate on dev** at the dev URL: the footer version, one report (including
   the PDF), and readable chart labels. A human must green-light before promoting.
4. **Deploy to prod:** open a PR into `prod-deploy` that sets the same
   `ARG EJAM_VERSION` tag (and `ARG EJAMDATA_VERSION`, if it changed). On merge,
   Actions deploys production. Then check the same things on the production app.

For the release steps around a deploy (tagging, the `ejamdata` release, EJAM-API
and EJSCREEN updates), see
[Releasing a New Version of EJAM](dev-update-package.html).

### PDF checks

PDF support is verified at three points because source-only testing does not
exercise the image's configuration:

1. **During the image build:** the package `configure` script performs a real
   `pagedown::chrome_print()` render whenever the build advertises Chrome through
   `CHROMOTE_CHROME` or `PAGEDOWN_CHROME` (or explicitly sets
   `EJAM_VERIFY_PDF_BUILD=true`). It verifies both the output size and `%PDF`
   magic bytes. A Chrome executable lookup alone is not treated as success.
2. **In the Shiny browser suite:** the full `latlon` case requests both PDF and
   HTML report downloads. Missing Pandoc is an error for this test, not a silent
   pass.
3. **After deployment:** `Verify deployed PDF generation` runs after either successful
   ECS deploy workflow. It performs a one-point analysis against the deployed
   development or production app, downloads the report, and verifies the `.pdf`
   filename and `%PDF` bytes. The workflow can also be dispatched manually.

The post-deployment verification workflow lives on the default branch, as required by
GitHub's `workflow_run` event, while the operational build/deploy workflows stay
on `dev-deploy` and `prod-deploy`.

**Required GitHub Actions secrets** (repo → Settings → Secrets and variables → Actions):

| Secret | Purpose |
|---|---|
| `AWS_ACCESS_KEY_ID`, `AWS_SECRET_ACCESS_KEY` | deploy credentials for ECR/ECS |

No GitHub token secret is needed. Both workflows pass the workflow run's own
short-lived token to the build (`--build-arg GITHUB_PAT=${{ github.token }}`),
which is enough to read the public EJAM and `ejamdata` repositories. (A stored
personal access token, `EJAMDATA_PAT`, was used until July 2026, when it expired
and broke the builds; it is no longer used.)

### Which EJAM version deploys

- **EJAM:** the `Dockerfile` installs EJAM with `remotes::install_github()` at
  `ARG EJAM_VERSION` (a tag, branch, or commit SHA; see
  [Branching and deploy model](#branching-and-deploy-model)).
- **`ejamdata`:** the 11 large `.arrow` files are downloaded separately, from the
  `ejamdata` GitHub release named by `ARG EJAMDATA_VERSION`. **It must equal the
  installed EJAM's `ejamdata_required_tag`** (in EJAM's `DESCRIPTION`; for
  EJAM v3.2022.3 that is `v3.2022.3`). If they differ, EJAM finds that the data in
  the image is not the release it requires and downloads the whole set (about 1 GB)
  again every time a container starts. (An explicit empty `EJAMDATA_VERSION` falls
  back to the "Latest" `ejamdata` release, which is not necessarily the one EJAM
  requires, so always set it.)
- A new `ejamdata` release, and so a new `EJAMDATA_VERSION`, is needed whenever
  `bgej`, the FRS files, or the block geography files change. A code-only EJAM
  release keeps the existing data tag. See
  [Releasing a New Version of EJAM](dev-update-package.html) for the full list of
  places that must name the same data tag.
- **Font cache:** the `Dockerfile` `CMD` must start with `system('fc-cache -r')`,
  before the app starts
  ([#621](https://github.com/Public-Environmental-Data-Partners/EJAM/issues/621)).
  On ECS Fargate the font cache built into the image did not match the fonts
  actually present, so R resolved "sans" to a decorative font from
  `texlive-fonts-extra` and chart text in reports came out as boxes (the same image
  looked fine under plain Docker). Rebuilding the cache at startup takes a few
  seconds. Keep this line on both deploy branches. (It was added to `dev-deploy` in
  [#623](https://github.com/Public-Environmental-Data-Partners/EJAM/pull/623); the
  matching `prod-deploy` change is
  [#624](https://github.com/Public-Environmental-Data-Partners/EJAM/pull/624).)

## First-time setup (manual / infrastructure)

Only needed when standing up the infrastructure or running Terraform locally;
day-to-day app deploys use GitHub Actions (above).

**Prerequisites** (macOS shown; use your platform's package manager):

```bash
# Homebrew, then:
brew install hashicorp/tap/terraform
brew install awscli
# plus Docker Desktop (docker.com), running before any Docker step
```

**AWS credentials.** Request credentials from the AWS account administrator, then:

```bash
aws configure          # region us-east-1, output json
aws sts get-caller-identity   # verify
```

Your IAM user needs the custom `ejam-terraform-deploy` policy to provision
infrastructure — request the policy JSON from the deployment manager and attach it
(AWS Console → IAM → Users → your user → Add permissions).

**Provision / update infrastructure with Terraform.** Run from `ejam-infra/`.
State lives in the S3 bucket `ejam-terraform-state-<ACCOUNT_ID>`, with a **separate
state key per environment**:

```bash
cd ejam-infra

# Prod
terraform init -backend-config="key=prod/terraform.tfstate"
terraform plan  -var-file=prod.tfvars -var="aws_account_id=<ACCOUNT_ID>"
terraform apply -var-file=prod.tfvars -var="aws_account_id=<ACCOUNT_ID>"

# Dev (separate state; -reconfigure switches the backend key)
terraform init -backend-config="key=dev/terraform.tfstate" -reconfigure
terraform apply -var-file=dev.tfvars -var="aws_account_id=<ACCOUNT_ID>"
```

Get your account ID with
`aws sts get-caller-identity --query Account --output text`.

### Manual Docker build & push (fallback)

Prefer GitHub Actions — the uncompressed image is large (~4 GB) and local pushes
are slow. If you must build/push by hand (from the deploy branch, which has the
`Dockerfile`):

```bash
# Authenticate Docker to ECR
aws ecr get-login-password --region us-east-1 \
  | docker login --username AWS --password-stdin \
    <ACCOUNT_ID>.dkr.ecr.us-east-1.amazonaws.com

# The ECR repo is created with image_tag_mutability = IMMUTABLE, so every push
# needs a UNIQUE tag -- re-pushing an existing tag (e.g. :latest) will fail.
# Match the workflows' commit-SHA convention:
TAG="manual-$(git rev-parse --short HEAD)"
# Pass a GitHub token as the workflows do (used to download the ejamdata files):
docker build --build-arg GITHUB_PAT="$(gh auth token)" -t ejam:$TAG .
docker tag  ejam:$TAG <ACCOUNT_ID>.dkr.ecr.us-east-1.amazonaws.com/ejam:$TAG
docker push <ACCOUNT_ID>.dkr.ecr.us-east-1.amazonaws.com/ejam:$TAG
```

The `.dockerignore` (on the deploy branch) excludes `.RData`, `.Rhistory`,
`.Rproj.user`, `.git`, `.github`, `ejam-infra/`, and several docs and test folders
from the build context. (The build does not copy anything from the context anyway;
see [Where the operational files live](#where-the-operational-files-live).)

## Infrastructure changes (Terraform)

App **code** deploys happen via GitHub Actions. AWS **infrastructure** changes
(resize a task, add HTTPS, change retention) are made by editing the `.tf`/`.tfvars`
files and running `terraform apply` locally from `ejam-infra/` as shown above.

**Custom domain / HTTPS:** set `domain_name = "ejam.yourdomain.com"` in the relevant
`.tfvars`, run `terraform apply`; Terraform outputs the CNAME records to add at the
DNS provider (Squarespace) for ACM certificate validation. Run `terraform apply`
once more after adding them — HTTP then redirects to HTTPS automatically.

## Rollback

Point the ECS service back at an earlier task-definition revision:

```bash
# List recent revisions
aws ecs list-task-definitions --family-prefix ejam --sort DESC \
  --query 'taskDefinitionArns[:5]' --output text

# Prod
aws ecs update-service --cluster ejam-prod-cluster \
  --service ejam-prod-service --task-definition ejam:<REVISION>

# Dev
aws ecs update-service --cluster ejam-dev-cluster \
  --service ejam-dev-service --task-definition ejam-dev:<REVISION>
```

## Logs and debugging

```bash
# Service health (running vs. desired)
aws ecs describe-services --cluster ejam-prod-cluster --services ejam-prod-service \
  --query 'services[0].{Status:status,Running:runningCount,Desired:desiredCount}'

# Recent errors in the last 30 minutes (prod). `date +%s` (current epoch) is
# portable across macOS and Linux; 1800 s = 30 min. CloudWatch wants milliseconds.
aws logs filter-log-events --log-group-name /ecs/ejam-prod \
  --filter-pattern "Error" \
  --start-time $(( ($(date +%s) - 1800) * 1000 )) \
  --query 'events[*].message' --output text
```

| Symptom | Likely fix |
|---|---|
| `UnauthorizedOperation` on an EC2/ECS/IAM action | add the missing action to the `ejam-terraform-deploy` IAM policy |
| Docker build fails | confirm Docker Desktop is running and `.dockerignore` is present |
| ECS tasks failing health checks | confirm the container serves HTTP 200 on the health-check port (2001) and the app port (2000) matches the task definition |
| Slow local Docker push | use GitHub Actions instead |
| Chart text in reports shows as boxes | the `Dockerfile` `CMD` is missing `system('fc-cache -r')` (see [Which EJAM version deploys](#which-ejam-version-deploys)) |
| Container logs show the `ejamdata` files downloading at every start | `ARG EJAMDATA_VERSION` does not equal the installed EJAM's `ejamdata_required_tag` |
| The app footer shows a different version than the one just deployed | two deploys ran at once (a merge and a manual run); run one again and let it finish alone |

## Testing against a local or draft API

The app computes analysis results in-process via `ejamit()` -- it does not call the
EJAM REST API for analysis. Where the API base URL matters is everywhere the app or
package **builds URLs that point at the API**: the per-site report links in results
tables and map popups (built by `url_ejamapi()`), `ejamapi()` calls, and the
EJScreen-to-EJAM token handoff. All of those read the base URL from one place,
`url_package("api")`, which normally comes from the `url_api` field of `DESCRIPTION`
(the production API, https://api.ejanalysis.com).

To test an app release candidate against a **different API** -- most usefully the local
API served by `EJAM:::ejamapi_local()`, which mirrors the latest EJAM-API code before it
is deployed anywhere (see the [API article](dev-api.html)) -- override that one lookup.
Precedence is: `options(ejam.api.baseurl=...)` first, then the environment variable
`EJAM_API_BASEURL`, then `DESCRIPTION`.

**Local app + local API** (no infrastructure needed):

```{r localapi, eval=FALSE}
apiproc <- EJAM:::ejamapi_local()               # local API at http://127.0.0.1:3035
Sys.setenv(EJAM_API_BASEURL = "http://127.0.0.1:3035")
# or, equivalently: options(ejam.api.baseurl = "http://127.0.0.1:3035")
EJAM::ejamapp()                                  # report links in tables/popups now
                                                 # hit the local API, end to end
# when done:
Sys.unsetenv("EJAM_API_BASEURL")
apiproc$kill()
```

The same override works for one-off calls without the app, e.g.
`ejamapi(fips = "10001", endpoint = "data")` or `url_ejamapi(lat = 34, lon = -118)`
will target whatever base is set.

**Deployed app (dev server) + draft API:** the AWS-hosted dev app cannot reach a
laptop's localhost, so point its `EJAM_API_BASEURL` (an environment variable in the
ECS task definition) at any reachable draft API deployment -- e.g. a future
apidev.ejanalysis.com staging service
([EJAM-API#47](https://github.com/Public-Environmental-Data-Partners/EJAM-API/issues/47)
tracks setting one up), or a temporary Cloudflare tunnel exposing a locally-run API.
Unsetting the variable (or omitting it) restores the production API from `DESCRIPTION`.

## Teardown

```bash
cd ejam-infra
terraform destroy -var-file=prod.tfvars -var="aws_account_id=<ACCOUNT_ID>"
```

The prod ALB has **deletion protection** enabled; disable it in the AWS Console
first (EC2 → Load Balancers → `ejam-prod-alb` → Edit attributes → Deletion
protection: off) before `terraform destroy` will succeed on prod.

## Future direction

The long-lived `dev-deploy` / `prod-deploy` branches are a workable pattern, but
their two `Dockerfile`s must be kept in step by hand, and the stale copy of the
package source on each branch is easy to mistake for what is deployed. The modern,
trunk-based alternative is to keep the infra + `Dockerfile` in subdirectories on
`main` (marked in `.Rbuildignore` so they don't affect the R package build) and
deploy via `workflow_dispatch`/tag triggers targeting **GitHub Environments**
(`dev`/`prod`) with approval gates — which removes the environment branches and
their divergence entirely. This is a larger change and is the deployment
maintainer's call; it is noted here as the recommended target state.

## About this document

This article consolidates two documents that previously lived only on the deploy
branches: `DEPLOY-GUIDE.md` (root) and `ejam-infra/README.md`. Where they
disagreed, it follows the **actual deploy-branch files** (`main.tf`, `Dockerfile`,
the deploy workflows): **GitHub Actions deploys are live** (the older
`DEPLOY-GUIDE.md` said "coming soon"), and the container serves on **ports
2000/2001** (as in `DEPLOY-GUIDE.md` and the current `main.tf`/`Dockerfile` — the
`ejam-infra/README.md` wireframe's "3838" was inaccurate). `DEPLOY-GUIDE.md` also
describes promoting code by merging `main` → `dev-deploy` → `prod-deploy`; that
no longer applies, because each image installs EJAM from GitHub at the ref in its
`Dockerfile` (see [Branching and deploy model](#branching-and-deploy-model)).
Concrete AWS account IDs and individual contact names have been replaced with
`<ACCOUNT_ID>` and role descriptions; substitute the real values from the deploy
branch files. If you own the deployment, treat this as the single source of truth
and retire the two originals (or replace them with a pointer here).
