---
title: "Future Plans and Ideas for EJAM"
description: "Future Plans and Ideas for EJAM"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Future Plans and Ideas for EJAM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

```{r developernote, eval=FALSE, echo= FALSE, include = FALSE}
#  *>>>>>>>>>> Developer note: vignettes need to be tested/edited/rebuilt regularly <<<<<<<<<<<*
#    - **See ?pkgdown::build_site** and script in EJAM/data-raw/- EJAM uses the pkgdown R package to build help and articles/ vignettes as web pages
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Status and Future Directions

This article highlights some key features or areas ripe for future enhancements. Additional specific feature requests or ideas are in github issues in the EJAM repository. Updates on the status of the repository, web app hosting, and open source package development are in the [README]() or at [ejanalysis.com/status](https://ejanalysis.com/status).


## Possible future enhancements

Although EPA stopped actively developing EJAM in 2025, it had already identified various options for future enhancements, including some of the ideas listed below. Subsequent non-EPA work on newer versions of EJAM has started to address some of these and is likely to continue to do so, in tandem with work on EJSCREEN itself. The open source EJAM R package is designed to be extensible, so that updated data, new kinds of indicators, new types of analyses, and new types of reports can be added over time by a community of contributors.

### API

In 2025, when the original EJSCREEN API was taken offline by EPA, a non-EPA effort replaced it with a new API ("EJAM API") that provides many of the original features. The new API became the tool through which EJSCREEN is able to provide a Community Report on a single location. It also allows the "EJSCREEN multisite tool" (aka EJAM's web app) to show links within tables of results and map popups, to link to a report on each site, for example. Some parallel work on an API was also drafted in the plumber folder of the EJAM source package. A future, expanded API is likely to provide the following: 
- reports in pdf format (not just html), for better pagination when printing
- better handling of polygons (ability to request a report on a large/complex polygon)
- ability to obtain a multisite summary report (not just a 1-site report) (via the sitenumber parameter)
- allowing the API to use most or all of the same parameters as the functions `ejamit()` and `ejam2report()`, providing more features and flexibility in customizing reports.
- data (an easy way to obtain subsets of data at the block or blockgroup level, for example, via an assets)
- enhancements and/or fixes as noted in [github issues for the EJAM-API repository](https://github.com/edgi-govdata-archiving/EJAM-API/issues)
- access to services such as fast identification of nearby blocks or other points (via `getblocksnearby()` `getblocksnearby_from_fips()` or `get_blockpoints_in_shape()`)
- summarization (aggregation over blocks or blockgroups), maybe for user-provided indicators


### User-specified Indicators

EJAM in the future will also be able to analyze, aggregate, and report on other residential population indicators (e.g., other Census ACS variables), to include user-selected EPA-hosted layers, or even user-provided data. In other words, an analysis would be able to include other layers on risks or concentrations (at blockgroup resolution), or user-provided scores for each blockgroup. This would first be available for use in R as one or more functions and would subsequently be made available as a set of features through the web app interface. That could separately include the ability to upload a user-provided blockgroup indicator for analysis in existing or distinct reports, and/or the ability to specify ACS variables and formulas to obtain and calculate new indicators on the fly. This could first be a set of functions available for use in R and could subsequently be made available for use through the web app interface. See preliminary work such as the following: `ACSdownload::get_acs_new()` as a way to obtain Census ACS demographics, `acs_bybg()` examples of another way to obtain data and then using ACS raw data to calculate indicators, `calctype()` for metadata on how a given indicator should get aggregated over blocks or blockgroups, `doaggregate()` parameters like wtdmeancols, and exploratory early drafts of possible approaches to creating new indicators or aggregating them for a report, in`calc_ejam()`, `custom_doaggregate()`, `doaggregate_newscores()`, etc.


### Counts of nearby points of interest

Features were drafted that will be able to report on the number of user-specified points of interest (POI) at each location. These POI could be amenities (e.g., parks, grocery stores), sources of risk or disamenities (e.g., diesel emissions sources, landfills), or potentially vulnerable locations (e.g., elementary schools, daycare, hospitals, nursing homes). The tool would report on counts of POI, in an analysis of circular buffers (e.g., \# of parks near each facility analyzed), but also for FIPS Census units (e.g., \# of parks in each County) or polygons/shapefiles (e.g., \# of parks in each high-risk zone). See `countpoints_nearby()`, and related early draft work in EJAM/R/PROXIMITY_FUNCTIONS.R. Note certain POI are already tallied in the reports, such as hospitals, schools, TRI sites, etc. and more information is available using `ejam2areafeatures()`.

### Facility Density Scores for all EPA-regulated Facilities

A facility density indicator may be explored and pre-calculated, that would summarize the proximity of all types of EPA-regulated facilities, or certain subcategories, for every community in the US or within one State of interest. This would be similar to the existing "proximity score" indicators such as the RMP score in EJSCREEN, but cover a larger range of facilities. See early draft work in `EJAM/R/PROXIMITY_FUNCTIONS.R`. This could first be a dataset available for use in R and could subsequently be made available for reporting through the web app interface.

### Proximity Scores for User-Specified Sites

A feature is planned that will be able to quickly calculate a new proximity indicator for every blockgroup in the US, to represent counts of and distances to user-specified points such as facilities. This would be in addition to or instead of a pre-calculated facility density score, and similar to the existing proximity indicators such as the RMP score in EJSCREEN, but it would be calculated on demand, when needed, for any user-specified set of points. See early draft work in `EJAM/R/PROXIMITY_FUNCTIONS.R`. This could first be a function available for use in R and could subsequently be made available for use through the web app interface.


### Written Long Report

EJAM is designed in a way that will allow a future update to provide a default, standard written report (e.g., such as a 10-page report). EJAM will provide a written report with text, graphics, and maps, ready for a user to print, download, share, and use. Because the report is a Word document, one can further edit it offline as needed. Users also can choose to use just selected tables or graphics or data files (such as for individual sites and summary statistics). See early draft work in `system.file("report/written_report", package = "EJAM")` and in R/app_ui.R and R/app_server.R.

### Methods for Identifying and Focusing on Key Findings

Communicating key findings is challenging when such a large number of metrics are available across multiple locations with multiple residential population groups. Some simple tools may be able to help with initial filtering of results, to suggest what might be highlighted. These could be used for a revised summary report, but especially will be explored for the long-form written report, particularly for ways to provide an Executive Summary. See early draft work in functions such as `count_sites_with_n_high_scores()` and the report/written_report folder.

### Visualization Tools

A variety of plots and maps will be explored as ways to communicate the rich data results calculated by EJAM. For example, EJAM R functions can calculate and display detailed estimates of the range of residential distances to facilities within each residential population subgroup - this type of visualization could be incorporated into the web app if there appears to be interest in it. Another possibility is that plots, maps, and tables could be made to interact in sync with each other, where that supports useful data exploration. See examples of plots in the vignettes, for example, and [the many plot_xyz functions and mapping functions](../reference/index.html#viewing-results).

### Spatial Resolution

EJAM is designed with growth in mind, assuming that the specific indicators of interest will expand or change over time. It is even possible that multiple resolutions of data will need to be analyzed, such as block, blockgroup, and tract data for different indicators. A subsequent refinement might even use a high-resolution raster grid of population estimates rather than the Census Block counts currently used for buffering and weighting blockgroup scores for partially included blockgroups.
