---
title: "Updating EJScreen Datasets Annually (via the Pipeline)"
description: "Annual staged workflow for updating ACS-based EJScreen/EJAM datasets"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Updating EJScreen Datasets Annually (via the Pipeline)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

```{r, eval=TRUE, echo=FALSE, include=FALSE}

code_reponame_url        <- EJAM::url_package(type = "code", get_full_url = TRUE)
code_reponame_with_owner <- EJAM::url_package(type = "code", get_full_url = FALSE)
code_reponame <- gsub(".*/", "", code_reponame_with_owner)

data_reponame_url        <- EJAM::url_package(type = "data", get_full_url = TRUE)
data_reponame_with_owner <- EJAM::url_package(type = "data", get_full_url = FALSE)
data_reponame <- gsub(".*/", "", data_reponame_with_owner)

docs_reponame_url        <- EJAM::url_package(type = "docs", get_full_url = TRUE)

```

This document describes the annual staged pipeline for updating the
blockgroup-level EJScreen/EJAM datasets, especially the objects historically
called `blockgroupstats`, `bgej`, `usastats`, and `statestats`.

For general dataset maintenance outside this pipeline, such as FRS-related
tables, NAICS/SIC tables, block-level files, Arrow releases, and package-data
publication mechanics, see [Updating and Managing the Datasets used by
EJAM](dev-update-datasets.html).

## Scope

The pipeline is designed to make each major data step explicit, saved, and
rerunnable. The main entry point is `calc_ejscreen_dataset()`. The recommended
annual runner is:

```{r, eval=FALSE}
source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

The runner writes pipeline checkpoints such as CSV files. It can also write
secondary formats such as `.rda` files for the same table stages when
`EJAM_STAGE_FORMATS` includes those formats. It does not, by itself, replace
every installed package dataset in `data/*.rda`. Replacing package data objects
with `usethis::use_data()` or EJAM metadata helpers is a separate release step
after the pipeline outputs have been reviewed.

## Pipeline Stages

The annual workflow creates or reads these stages:

1. `bg_acs_raw`: raw ACS table-based summary file data downloaded from the
   Census Bureau. This is saved before EJAM renaming or formula calculations.
   The default raw ACS storage is a folder of per-table files, with a manifest.

2. `bg_acsdata`: ACS-derived blockgroup indicators calculated from
   `bg_acs_raw`, including demographic indicators and `pctpre1960`.

3. `bg_envirodata`: blockgroup environmental indicators used for EJ indexes.
   This normally comes from a separate environmental-data workflow. For draft
   builds, it can be provisionally reused from the current package data.

4. `bg_extra_indicators`: other blockgroup indicators that are not ACS and not
   the main EJ environmental indicators, such as health, life expectancy, and
   related context variables. These can also be provisionally reused from the
   current package data.

5. `bg_geodata`: Census/TIGER blockgroup geography attributes. This stage
   stores `bgfips`, square-meter `arealand` and `areawater`, optional internal
   point fields `intptlat` and `intptlon`, and a compatibility-only `area`
   column. The pipeline uses Census TIGER/Line blockgroup shapefiles by default
   because their `ALAND` and `AWATER` values best match the legacy EJScreen
   tables; Census TIGERweb remains available as a lighter fallback source.
   TIGER/Line zip files are cached locally using `EJAM_TIGER_BG_CACHE_DIR`, or
   the EJAM user cache folder when that variable is unset, so later reruns can
   reuse the state files.
   By default, the pipeline requests geography only for blockgroups found in
   the ACS tabulated rows for that vintage.

6. `blockgroupstats`: combined blockgroup table with ACS indicators,
   environmental indicators, extra indicators, and geography fields.

7. `usastats_acs`, `statestats_acs`, `usastats_envirodata`, and
   `statestats_envirodata`: percentile lookup tables for ACS and environmental
   inputs.

8. `bgej`: blockgroup EJ index values calculated from demographic indexes and
   environmental percentiles.

9. `usastats_ej` and `statestats_ej`: percentile lookup tables for EJ index
   columns.

10. `usastats` and `statestats`: combined lookup tables used by EJAM.

11. `ejscreen_export`: EJScreen-ready export combining `blockgroupstats` and
    `bgej`, applying EJScreen-style names from `map_headernames`, and adding
    map helper fields where possible.

The runner also writes `pipeline_validation_summary.csv` and
`pipeline_run_manifest.csv`. If prior-version validation is requested, it writes
`prior_validation_summary.csv` and per-stage prior-validation details. If the
EJScreen export is requested, it writes `ejscreen_export_schema_report.csv`.

The run manifest records the package version, Git branch and SHA, ACS vintage,
pipeline location, primary stage format, selected run settings, and whether
provisional environmental or extra-indicator inputs were reused.

## Related Data Update Groups

EJAM data objects do not all change on the same schedule. For annual work, use
these update groups to decide what must be rebuilt, validated, or only checked.

1. **Facility Data Updates** include FRS-related and facility-code datasets such
   as `frs`, `frs_by_programid`, `frs_by_naics`, `frs_by_sic`,
   `frs_by_mact`, `frsprogramcodes`, `epa_programs`, NAICS, SIC, and MACT
   lookup tables. These can be refreshed when EPA facility data are updated,
   independently of the annual EJScreen-style pipeline.

2. **EJSCREEN Annual Data Update** includes the main pipeline stages:
   `bg_acs_raw`, `bg_acsdata`, `bg_envirodata`, `bg_extra_indicators`,
   `bg_geodata`, `blockgroupstats`, `bgej`, `usastats`, `statestats`, and
   `ejscreen_export`. It also includes supporting objects that may need review
   or regeneration, such as `map_headernames`, `names_*`, `namez`,
   `tables_ejscreen_acs`, `formulas_ejscreen_acs`,
   `formulas_ejscreen_acs_disability`, `formulas_ejscreen_demog_index`,
   `avg.in.us`, `high_pctiles_tied_with_min`, and `testoutput*`.
   `bgej.arrow` belongs in this group and must be pinned to the EJAM/EJScreen
   annual release, not treated as independently refreshable "latest" data.
   In practice, `dataload_dynamic("bgej")` uses the installed EJAM package
   version, as reported by `packageVersion("EJAM")`, to find the matching
   `ejamdata` release tag, such as `v2.5.0`, and checks the loaded `bgej`
   against the installed `blockgroupstats` blockgroup universe.

3. **Blockgroup Geography Updates** include blockgroup-keyed geography files
   and crosswalks such as `bgid2fips`, `blockwts`, `bgpts`, and
   `bg_cenpop2020`. The annual pipeline always checks whether these are still
   compatible with the current blockgroup universe, but they only need to be
   regenerated when blockgroup FIPS, EJAM `bgid`, internal points, or
   blockgroup-to-block relationships change. `bg_cenpop2020` requires special
   care because it is tied to Census 2020 geography. Related state/place
   geography objects such as `states_shapefile`, `stateinfo2`, and
   `censusplaces` should also be checked when their FIPS or boundaries change.

4. **Block Geography Updates** include block-level geometry/index files such as
   `blockpoints`, `quaddata`, and `blockid2fips`. These change only when block
   FIPS or block internal-point geography changes.

The runner writes `dynamic_geography_arrow_report.csv` to summarize whether
category 3 and 4 Arrow files cover the current `blockgroupstats` blockgroups
and whether block-level IDs line up across `blockwts`, `blockpoints`,
`quaddata`, and `blockid2fips`.

## ACS Geography Universe

For a given ACS 5-year release, use the Census geography vintage associated
with the ACS end year. For example, ACS 2020-2024 should use 2024 Census/TIGER
or TIGERweb geography attributes. The release pipeline prefers the downloadable
Census TIGER/Line blockgroup shapefiles for `arealand` and `areawater`, with
TIGERweb as a fallback. Those geography sources can occasionally include
blockgroups that are valid geography features but that are not present in the
ACS tabulated summary-file tables used by EJAM.

That mismatch is unusual but real. In the draft ACS 2020-2024 build, the
Census/TIGER blockgroup geography source included 39 Suffolk County, New York
blockgroups that were not present in the relevant ACS blockgroup or tract
tables downloaded for the pipeline. Including those geography-only rows would
expand `blockgroupstats` beyond the ACS data universe and create rows with no
ACS-derived indicators.

For that reason, the default pipeline setting is
`EJAM_BLOCKGROUP_UNIVERSE_SOURCE = "acs"`. Under that setting, `bg_acsdata`
defines the final blockgroup universe, `bg_geodata` is downloaded or subset to
those `bgfips` values, and environmental or extra-indicator inputs cannot add
extra rows to the final `blockgroupstats`. The alternative setting
`EJAM_BLOCKGROUP_UNIVERSE_SOURCE = "union"` is kept only for diagnostic or
special-purpose runs where the maintainer intentionally wants to retain
blockgroups found only in other inputs.

The pipeline uses the packaged `formulas_ejscreen_acs` object for ACS-derived
indicator formulas. The old
`data-raw/archived_datacreate_formulas_ejscreen_acs_notes.R` file is reference
material only. It is not the current formula rebuild workflow.

## Storage

The default local pipeline folder is:

```{r, eval=FALSE}
data-raw/pipeline_outputs/ejscreen_acs_2024
```

`data-raw/pipeline_outputs/` is ignored by Git because checkpoint files can be
large. The repository also has build-ignore rules for pipeline outputs and
Arrow data files, so release artifacts should normally be stored outside the
package source tree.

The pipeline can also use AWS S3. S3 support uses the AWS CLI, so `aws` must be
installed and configured before running an S3-backed pipeline. By default the
runner uses `EJAM_STAGE_FORMAT = "csv"` for loading/review and
`EJAM_STAGE_FORMATS = "csv,rda"` for saving major table stages in both formats.
Small summary and manifest files are written as CSV.

```{r, eval=FALSE}
Sys.setenv(
  AWS_PROFILE = "ejam",
  AWS_REGION = "us-east-1",
  EJAM_PIPELINE_DIR = "s3://pedp-data-preserved/ejscreen-data-processing/pipeline/ejscreen_acs_2024",
  EJAM_PIPELINE_STORAGE = "s3"
)
```

For local testing, use a local directory:

```{r, eval=FALSE}
Sys.setenv(
  EJAM_PIPELINE_DIR = file.path(
    getwd(),
    "data-raw",
    "pipeline_outputs",
    "ejscreen_acs_2024"
  ),
  EJAM_PIPELINE_STORAGE = "local"
)
```

## Key Settings

The runner is controlled by environment variables:

| Variable | Purpose |
|---|---|
| `EJAM_PIPELINE_YR` | ACS 5-year end year, such as `"2024"` for ACS 2020-2024. |
| `EJAM_PIPELINE_DIR` | Local folder or `s3://...` pipeline location. |
| `EJAM_PIPELINE_STORAGE` | `"auto"`, `"local"`, or `"s3"`. |
| `EJAM_STAGE_FORMAT` | Primary stage format used for loading/validation, usually `"csv"`. |
| `EJAM_STAGE_FORMATS` | Comma-separated formats to save for major table stages, usually `"csv,rda"`. |
| `EJAM_BLOCKGROUP_UNIVERSE_SOURCE` | `"acs"` uses the ACS tabulated blockgroup rows as the final universe. `"union"` also keeps rows found only in environmental or extra-indicator inputs. |
| `EJAM_TRACT_WEIGHT_SOURCE` | `"decennial2020"` uses 2020 Decennial Census population weights to apportion tract-only ACS tables to blockgroups, matching legacy EJSCREEN. `"acs"` uses same-vintage ACS blockgroup population weights. |
| `EJAM_DECENNIAL_BGWTS_CACHE` | Optional local `.rds` cache path for 2020 Decennial blockgroup-to-tract weights. If unset, EJAM uses a user cache folder. |
| `EJAM_REFRESH_DECENNIAL_BGWTS` | `"TRUE"` to ignore and overwrite the cached decennial blockgroup weights. |
| `EJAM_TIGER_BG_CACHE_DIR` | Optional local folder for downloaded Census TIGER/Line blockgroup zip files. If unset, EJAM uses a durable user cache folder. |
| `AWS_PROFILE`, `AWS_REGION` | Used by the AWS CLI for S3-backed runs. |
| `CENSUS_API_KEY` | Used by ACS/Census download helpers where needed. |
| `EJAM_FORCE_ACS` | `"TRUE"` to redownload raw ACS and rebuild ACS stages. |
| `EJAM_FORCE_BG_ACSDATA` | `"TRUE"` to rebuild `bg_acsdata` from saved raw ACS. |
| `EJAM_FORCE_BG_GEODATA` | `"TRUE"` to redownload/rebuild the Census/TIGER `bg_geodata` stage. |
| `EJAM_ACS_DOWNLOAD_TIMEOUT` | Download timeout in seconds. Useful for large ACS tables. |
| `EJAM_ACS_DOWNLOAD_RETRIES` | Number of retry attempts for ACS downloads. |
| `EJAM_USE_PROVISIONAL_BG_ENVIRODATA` | `"FALSE"` to require a supplied `bg_envirodata.csv`. |
| `EJAM_INCLUDE_EJSCREEN_EXPORT` | `"TRUE"` to create `ejscreen_export.csv`. |
| `EJAM_VALIDATE_VS_PRIOR` | `"TRUE"` to write prior-version comparison files. |
| `EJAM_PRIOR_PIPELINE_DIR` | Prior pipeline folder/S3 prefix to compare against. |
| `EJAM_PRIOR_PACKAGE_REF` | Optional Git ref/tag/SHA for prior package data comparison. |

To see what the runner will use:

```{r, eval=FALSE}
Sys.getenv(c(
  "EJAM_PIPELINE_YR",
  "EJAM_PIPELINE_DIR",
  "EJAM_PIPELINE_STORAGE",
  "EJAM_STAGE_FORMAT",
  "EJAM_STAGE_FORMATS",
  "EJAM_BLOCKGROUP_UNIVERSE_SOURCE",
  "EJAM_TRACT_WEIGHT_SOURCE",
  "EJAM_DECENNIAL_BGWTS_CACHE",
  "EJAM_REFRESH_DECENNIAL_BGWTS",
  "EJAM_TIGER_BG_CACHE_DIR",
  "AWS_PROFILE",
  "AWS_REGION",
  # "CENSUS_API_KEY",
  "EJAM_FORCE_ACS",
  "EJAM_FORCE_BG_ACSDATA",
  "EJAM_FORCE_BG_GEODATA",
  "EJAM_ACS_DOWNLOAD_TIMEOUT",
  "EJAM_ACS_DOWNLOAD_RETRIES",
  "EJAM_USE_PROVISIONAL_BG_ENVIRODATA",
  "EJAM_INCLUDE_EJSCREEN_EXPORT",
  "EJAM_VALIDATE_VS_PRIOR",
  "EJAM_PRIOR_PIPELINE_YR",
  "EJAM_PRIOR_PIPELINE_DIR",
  "EJAM_PRIOR_PACKAGE_REF"
))
```

For ACS 2022 and later, Connecticut ACS tract FIPS use planning-region county
equivalents while 2020 Decennial blockgroup FIPS use the older county
equivalents. The pipeline detects that no Connecticut tract FIPS overlap in the
decennial weight table and uses same-vintage ACS blockgroup population weights
for Connecticut only. In normal package use, the decennial weights are created
from packaged `bg_cenpop2020` data. If that data is unavailable, EJAM falls back
to `tidycensus::get_decennial()` and caches the downloaded weights locally.

## New or renamed indicators

### map_headernames

If new indicators are being used compared with prior version of the datasets and package, 
note that map_headernames may need to be modified to have metadata on those new indicators, 
including the variable name (rname column of map_headernames), longname, calculation type, weight,
rounding information, and "varlist" which groups related indicators into, e.g., names_e, and
demographic key indicators into names_d, etc. 

### names_*

Much of the code depends on the varlist info,
so names_e, names_d, etc. need to be updated. The script in datacreate_names_of_indicators
uses map_headernames$varlist to update these data objects.


## Run a Fresh ACS Update

Start from a clean branch. For ACS 2020-2024 using local checkpoints:

```{r, eval=FALSE}
yr <- "2024"

Sys.setenv(
  EJAM_PIPELINE_YR = yr,
  EJAM_PIPELINE_DIR = file.path(
    getwd(),
    "data-raw",
    "pipeline_outputs",
    paste0("ejscreen_acs_", yr)
  ),
  EJAM_PIPELINE_STORAGE = "local",
  EJAM_STAGE_FORMAT = "csv",
  EJAM_STAGE_FORMATS = "csv,rda",
  EJAM_BLOCKGROUP_UNIVERSE_SOURCE = "acs",
  EJAM_TRACT_WEIGHT_SOURCE = "decennial2020",
  EJAM_TIGER_BG_CACHE_DIR = file.path(tools::R_user_dir("EJAM", "cache"), "tiger_bg"),
  EJAM_FORCE_ACS = "TRUE",
  EJAM_FORCE_BG_ACSDATA = "TRUE",
  EJAM_FORCE_BG_GEODATA = "TRUE",
  EJAM_ACS_DOWNLOAD_TIMEOUT = "3600",
  EJAM_ACS_DOWNLOAD_RETRIES = "2",
  EJAM_INCLUDE_EJSCREEN_EXPORT = "TRUE"
)

source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

For the S3-backed pipeline:

```{r, eval=FALSE}
yr <- "2024"

Sys.setenv(
  AWS_PROFILE = "ejam",
  AWS_REGION = "us-east-1",
  EJAM_PIPELINE_YR = yr,
  EJAM_PIPELINE_DIR = paste0(
    "s3://pedp-data-preserved/ejscreen-data-processing/pipeline/ejscreen_acs_",
    yr
  ),
  EJAM_PIPELINE_STORAGE = "s3",
  EJAM_STAGE_FORMAT = "csv",
  EJAM_STAGE_FORMATS = "csv,rda",
  EJAM_BLOCKGROUP_UNIVERSE_SOURCE = "acs",
  EJAM_TRACT_WEIGHT_SOURCE = "decennial2020",
  EJAM_TIGER_BG_CACHE_DIR = file.path(tools::R_user_dir("EJAM", "cache"), "tiger_bg"),
  EJAM_FORCE_ACS = "TRUE",
  EJAM_FORCE_BG_ACSDATA = "TRUE",
  EJAM_FORCE_BG_GEODATA = "TRUE",
  EJAM_ACS_DOWNLOAD_TIMEOUT = "3600",
  EJAM_ACS_DOWNLOAD_RETRIES = "2",
  EJAM_INCLUDE_EJSCREEN_EXPORT = "TRUE"
)

source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

The runner asks for confirmation in interactive sessions before continuing.

## Rerun From Saved ACS Data

If raw ACS has already been downloaded, rerun downstream ACS calculations
without redownloading:

```{r, eval=FALSE}
Sys.setenv(
  EJAM_FORCE_ACS = "FALSE",
  EJAM_FORCE_BG_ACSDATA = "TRUE"
)

source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

If both raw ACS and `bg_acsdata` should be reused, leave both force flags false:

```{r, eval=FALSE}
Sys.setenv(
  EJAM_FORCE_ACS = "FALSE",
  EJAM_FORCE_BG_ACSDATA = "FALSE"
)

source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

If `bg_geodata` has already been created for the same ACS/TIGER vintage and
same blockgroup universe, leave `EJAM_FORCE_BG_GEODATA` false. Set it to
`"TRUE"` when changing vintages or when you want to refresh the Census
TIGER/Line area and internal-point attributes.
Even with `EJAM_FORCE_BG_GEODATA = "TRUE"`, already-downloaded TIGER/Line state
zip files are reused from `EJAM_TIGER_BG_CACHE_DIR` when present and valid.

```{r, eval=FALSE}
Sys.setenv(EJAM_FORCE_BG_GEODATA = "TRUE")
source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

## Supplying Updated Environmental Data

The environmental stage is intentionally separate from the ACS stage. When
updated environmental indicators are available, save them as `bg_envirodata.csv`
in the pipeline folder.

For a local pipeline:

```{r, eval=FALSE}
file.path(pipeline_dir, "bg_envirodata.csv")
```

For an S3 pipeline:

```{r, eval=FALSE}
s3://pedp-data-preserved/ejscreen-data-processing/pipeline/ejscreen_acs_2024/bg_envirodata.csv
```

The file must include `bgfips` and the environmental indicators used for EJ
indexes. It should also include `pctpre1960`. The environmental-data workflow
may create `pctpre1960` by reading the saved `bg_acsdata.csv` stage.

To force the runner to stop unless `bg_envirodata.csv` has been supplied:

```{r, eval=FALSE}
Sys.setenv(EJAM_USE_PROVISIONAL_BG_ENVIRODATA = "FALSE")
source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

After replacing `bg_envirodata.csv`, rerun without forcing ACS:

```{r, eval=FALSE}
Sys.setenv(
  EJAM_FORCE_ACS = "FALSE",
  EJAM_FORCE_BG_ACSDATA = "FALSE",
  EJAM_USE_PROVISIONAL_BG_ENVIRODATA = "FALSE"
)

source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

This reuses the saved ACS stages and regenerates downstream
`blockgroupstats`, `bgej`, `usastats`, `statestats`, and `ejscreen_export`.

## Extra Indicators

Some `blockgroupstats` columns are not ACS indicators and are not the main
environmental indicators. Examples include life expectancy, health indicators,
facility-count context variables, climate-related fields, and other columns
grouped in `map_headernames$varlist`.

The pipeline makes these explicit in `bg_extra_indicators.csv`. If an updated
table is not supplied, the runner currently creates a provisional version from
the packaged `EJAM::blockgroupstats`. That is useful for testing the ACS update,
but it should be documented clearly as reuse of older non-ACS data.

## Provisional Draft Builds

For early pipeline testing, it is acceptable to reuse existing environmental and
extra indicators:

```{r, eval=FALSE}
Sys.setenv(EJAM_USE_PROVISIONAL_BG_ENVIRODATA = "TRUE")
source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

The runner writes source-note text files next to provisional stages, such as
`bg_envirodata_SOURCE.txt` and `bg_extra_indicators_SOURCE.txt`. Final release
review should confirm whether any provisional stage remains.

## Reviewing Outputs

Start with the validation summary:

```{r, eval=FALSE}
library(data.table)

pipeline_dir <- "data-raw/pipeline_outputs/ejscreen_acs_2024"

validation <- fread(file.path(pipeline_dir, "pipeline_validation_summary.csv"))
validation[, .(stage, rows, columns, errors, warnings)]
validation[nzchar(errors)]
validation[nzchar(warnings)]
```

There should be no validation errors. Warnings should be understood and either
fixed or explicitly accepted for a draft build.

Then inspect the main outputs:

```{r, eval=FALSE}
bg_acsdata      <- fread(file.path(pipeline_dir, "bg_acsdata.csv"))
blockgroupstats <- fread(file.path(pipeline_dir, "blockgroupstats.csv"))
bgej            <- fread(file.path(pipeline_dir, "bgej.csv"))
usastats        <- fread(file.path(pipeline_dir, "usastats.csv"))
statestats      <- fread(file.path(pipeline_dir, "statestats.csv"))
bg_geodata      <- fread(file.path(pipeline_dir, "bg_geodata.csv"))

nrow(blockgroupstats)
nrow(bgej)
names(blockgroupstats)
names(bgej)
```

Useful checks include:

- expected FIPS/geography columns are present and typed as character;
- row counts are plausible for the ACS vintage and geography coverage;
- under the default `EJAM_BLOCKGROUP_UNIVERSE_SOURCE = "acs"`,
  `blockgroupstats`, `bgej`, and `bg_geodata` have the same `bgfips` values as
  `bg_acsdata`;
- key ACS indicators are non-missing for most populated blockgroups;
- percentage/rate variables are in the expected range;
- `blockgroupstats` and `bgej` join cleanly by `bgfips`;
- `bg_geodata` has one row per `bgfips` and non-missing, nonnegative
  `arealand` and `areawater`;
- lookup tables include `REGION`, `PCTILE`, `0`, `100`, and `mean`;
- `usastats` has one region, `"USA"`;
- `statestats` has expected state/territory regions.

For example:

```{r, eval=FALSE}
stopifnot("0" %in% as.character(usastats$PCTILE))
stopifnot("100" %in% as.character(usastats$PCTILE))
stopifnot("mean" %in% as.character(usastats$PCTILE))

stopifnot("0" %in% as.character(statestats$PCTILE))
stopifnot("100" %in% as.character(statestats$PCTILE))
stopifnot("mean" %in% as.character(statestats$PCTILE))

stopifnot(!anyDuplicated(blockgroupstats$bgfips))
stopifnot(!anyDuplicated(bgej$bgfips))
stopifnot(!anyDuplicated(bg_geodata$bgfips))
stopifnot(setequal(blockgroupstats$bgfips, bg_acsdata$bgfips))
stopifnot(setequal(bgej$bgfips, bg_acsdata$bgfips))
stopifnot(setequal(bg_geodata$bgfips, bg_acsdata$bgfips))
stopifnot(all(bg_geodata$arealand >= 0, na.rm = TRUE))
stopifnot(all(bg_geodata$areawater >= 0, na.rm = TRUE))
```

Also review the run manifest:

```{r, eval=FALSE}
manifest <- fread(file.path(pipeline_dir, "pipeline_run_manifest.csv"))
manifest[key %in% c(
  "package_version",
  "git_sha",
  "git_branch",
  "git_dirty",
  "acs_version",
  "stage_format",
  "setting_EJAM_STAGE_FORMATS",
  "used_provisional_bg_envirodata",
  "used_provisional_bg_extra_indicators"
)]
```

For S3-backed runs, read the same files from S3 using EJAM's pipeline input
helpers or the AWS CLI.

## Slow Stages

The longest stages are usually:

- fresh `bg_acs_raw` downloads, because several ACS table-based summary files
  are large;
- S3 writes of large raw and derived stage files, especially when saving both
  `.csv` and `.rda`;
- `bg_acsdata`, because it applies ACS formulas and tract-to-blockgroup
  calculations;
- `bg_geodata`, because it downloads and reads Census TIGER/Line blockgroup
  shapefiles for every state and Puerto Rico, with TIGERweb available as a
  fallback. Reusing `EJAM_TIGER_BG_CACHE_DIR` makes later `bg_geodata` rebuilds
  much faster because the state zip files do not need to be downloaded again;
- final `blockgroupstats`/`bgej`/statistics/export calculations;
- prior-version validation when comparing large blockgroup tables.

When debugging pipeline speed, check the console timestamps and consider
rerunning without forcing earlier stages once their saved outputs are known to
be current.

## Reviewing the EJScreen Export Schema

If `EJAM_INCLUDE_EJSCREEN_EXPORT` is true, the runner writes:

```{r, eval=FALSE}
ejscreen_export.csv
ejscreen_export_schema_report.csv
```

Use the schema report as a field-by-field checklist:

```{r, eval=FALSE}
schema <- fread(file.path(pipeline_dir, "ejscreen_export_schema_report.csv"))

schema[, .N, by = status]
schema[status == "missing_expected"]
schema[status == "missing_expected", .N, by = field_type][order(-N)]
schema[status == "unexpected_output"]
```

Each missing expected field should be classified as one of:

1. a field EJScreen needs and the export must add;
2. a metadata mapping issue in `map_headernames`;
3. a deliberately deferred field that is not needed for the current export.

For release, the ideal schema report has no `missing_expected` rows for the
FeatureServer fields required by the EJScreen app.

## Replacing Package Data

Pipeline stage files are review artifacts. After they are accepted, update the
package data objects deliberately. The runner currently has an interactive
helper path for replacing `blockgroupstats`, but it does not automatically
replace every final package dataset.

A release update should explicitly replace at least:

- `blockgroupstats`;
- `bgej`;
- `usastats`;
- `statestats`;
- any related lookup or metadata objects that changed.

Use the established EJAM metadata helpers before saving package data. A typical
manual pattern is:

```{r, eval=FALSE}
blockgroupstats <- fread(file.path(pipeline_dir, "blockgroupstats.csv"))
bgej            <- fread(file.path(pipeline_dir, "bgej.csv"))
usastats        <- fread(file.path(pipeline_dir, "usastats.csv"))
statestats      <- fread(file.path(pipeline_dir, "statestats.csv"))

EJAM:::metadata_add_and_use_this("blockgroupstats")
EJAM:::metadata_add_and_use_this("bgej")
EJAM:::metadata_add_and_use_this("usastats")
EJAM:::metadata_add_and_use_this("statestats")
```

Confirm the exact metadata values before saving, especially ACS vintage,
Census vintage, EJSCREEN/EJAM version, data source notes, and provisional reuse
notes.

After those key datasets are updated, rerun the scripts that create and saves
testoutput_* files and datasets, especially `datacreate_testpoints_testoutputs.R`
and `datacreate_testoutput_ejamit_*`, etc.

After package data are updated, reinstall the package and rerun release-critical
tests. Also regenerate any Arrow-format files used outside the package build,
if needed.

## Release Checklist

Before releasing a new annual dataset build:

1. Run the pipeline with the intended ACS year and storage backend.
2. Confirm whether `bg_envirodata` and `bg_extra_indicators` are updated or
   provisional.
3. Review `pipeline_run_manifest.csv`, including package version, Git SHA, ACS
   vintage, run settings, and provisional-input flags.
4. Review `pipeline_validation_summary.csv`.
5. Review row counts, missingness, ranges, and joins for final tables,
   including the `bg_geodata` area fields.
6. Review `ejscreen_export_schema_report.csv`.
7. Replace package `.rda` datasets only after the stage files are accepted.
8. Update dataset metadata and documentation.
9. Rebuild documentation and pkgdown.
10. Reinstall the package and run focused tests plus package checks.
11. Publish large artifacts through the chosen storage path, such as S3 or the
    data repository release process.

The general release and large-data publication steps are covered in [Updating
and Managing the Datasets used by EJAM](dev-update-datasets.html) and [Updating
the Package as a New Release](dev-update-package.html).
