---
title: "Defaults and Custom Settings for the Web App"
description: "Defaults and Custom Settings for the Web App"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Defaults and Custom Settings for the Web App}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 100
---

```{r developernote, eval=FALSE, echo= FALSE, include = FALSE}
#  *>>>>>>>>>> Developer note: vignettes need to be tested/edited/rebuilt regularly <<<<<<<<<<<*
#    - **See ?pkgdown::build_site** and script in EJAM/data-raw/- EJAM uses the pkgdown R package to build help and articles/ vignettes as web pages
```

```{r SETUP_default_eval_or_not, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(eval = FALSE)
# https://r-pkgs.org/vignettes.html
```

```{r libraryEJAM, eval = TRUE, echo= FALSE, include= FALSE}
# rm(list = ls()); golem::detach_all_attached(); devtools::load_all()

if (!exists("blockgroupstats")) {library(EJAM)} # use installed version only if pkg not yet attached

# dataload_dynamic('all') # varnames = all  currently means all defined by .arrow_ds_names
# 
# indexblocks()
```


# Deep links: launch the EJAM app pre-loaded via URL parameters

Besides the deploy-time settings described below, an already-running EJAM app can be
**launched pre-loaded with a set of places** by opening it with launch query parameters in
the URL. This is how the EJScreen Report tool's "Send to EJAM" button opens EJAM ready to
run, and what `url_ejamapp()` builds as shareable deep links.

Deep linking, in other words, lets you send selected sites from the EJScreen app
to the EJAM app (multisite tool) via a "handoff"
that lets you continue in EJAM for further analysis.

EJAM app v3.x handles a few deeplink parameters.

The app reads these at startup (one place-type per launch -- points, else FIPS, else
polygons):

- `?lat=33,34&lon=-112,-114` -- one or more points, each treated as a site
- `?fips=10001,10003` -- one or more FIPS codes, each a separate site
- `?shape=<url-encoded GeoJSON>` -- a polygon or FeatureCollection
- `?radius=` (or `?buffer=`) -- analysis radius in miles
- `?handoff=<token>` -- a token minted by the EJAM API `POST /handoff` and fetched back at
  startup, for large polygon sets that exceed URL-length limits

Some examples of URLs:

The URL https://ejam.ejanalysis.com/  is an alias for
the production server hosting the live EJAM shiny app 
at https://ejam.publicenvirodata.org,
and it provides a way to pass through URL-encoded parameters.

The URL https://ejamdev.ejanalysis.com is an alias for
the development / staging server hosting a live EJAM shiny app.
(It redirects to an AWS load balancer whose literal hostname can change when the
infrastructure is rebuilt -- one reason the stable alias is the address to share --
and it also provides a way to pass through URL-encoded parameters.)

Deep links based on lat and lon and radius
for one or more points (staging examples; links for the production app are the
same but with the production base, which is `url_ejamapp()`'s default):

- https://ejamdev.ejanalysis.com/?lat=33&lon=-112&radius=2
- https://ejamdev.ejanalysis.com/?lat=33,33.1&lon=-112,-112.2&radius=2

Deep links based on FIPS code of Census unit like county:

- https://ejamdev.ejanalysis.com/?fips=24029&buffer=0
- https://ejamdev.ejanalysis.com/?fips=24029,24031

```{r url_ejamapp, eval=FALSE}
## EJAM app deep link (default base URL)
urlx <- url_ejamapp(fips = c(3650617, "0973000"))
urlx
# browseURL(urlx)

## For a non-default deployment, override the base URL:
# browseURL(url_ejamapp(fips = c(3650617, "0973000"), baseurl = "<staging-app-baseurl>"))
```

Build such a link with `url_ejamapp()`; its query vocabulary matches `url_ejamapi()` (which
builds *API* request URLs rather than *app* links). The default app base is
`https://ejamapp.ejanalysis.com/`, a Cloudflare-fronted shortcut on the `ejanalysis.com`
domain that forwards the query string to the live app (a plain redirect that dropped the
query would open the app empty). For the API side -- including the friendlier
`https://api.ejanalysis.com/` base and parameter details -- see the `dev-api` article, the
help for `ejamapi()`/`url_ejamapi()`, and the EJAM-API repository's README.

## Deep links into EJScreen (the mapping tool)

The PEDP EJScreen map app accepts a symmetric set of launch-URL parameters, so a link can
open EJScreen with place(s) already drawn and **selected for analysis** -- the reverse
direction of the "Send to EJAM" handoff described above. Build these links with
`url_ejscreenmap()`, whose default `baseurl=` points at the current live EJScreen
deployment (override `baseurl=` for a different one); they are plain URLs anyone can
construct:

- `?fips=10001` or `?fips=10001,10003,10001040100` -- one or more FIPS codes,
  comma-separated: 5-digit county, 11-digit tract, or 12-digit block group (mixes allowed).
  EJScreen queries its boundary services, draws each area, and selects it. State (2-digit)
  and city/CDP (7-digit) codes have no boundary service in the app, so use `wherestr=` with
  a place name for those -- `url_ejscreenmap()` does that fallback automatically via
  `fips2name()`.
- `?lat=39,39.7&lon=-75.5,-75.6&radius=3` -- one or more points (comma-separated, equal
  lengths); the optional `radius=` (miles) prefills the site report's buffer distance.
- `?polygon=39.1,-75.5;39.2,-75.4;39.0,-75.3` -- a polygon as semicolon-separated `lat,lon`
  vertices (at least 3); repeat the parameter for more than one polygon. `shape=`, `shp=`,
  and `shapefile=` are aliases (synonyms) for `polygon=`, matching the parameter aliases
  EJAM functions like `ejamit()` accept.
- `?zip=10001` or `?zip=10001,99501` -- the explicit way to pass zip code(s),
  comma-separated: each is geocoded as a zip (never read as a county FIPS) and pinned
  with the report popup; one zip centers the map, several fit the view to all of them.
- `?wherestr=Dover%2C%20DE` or `?wherestr=39.157,-75.52` -- the original EJScreen
  parameter: geocodes a place name / address / zipcode (or centers on a `lat,lon`) and
  drops a pin, without selecting a boundary -- except a bare 5-digit value, see below.

One kind of place is used per link: the EJScreen app reads these URL parameters in the
order above, so a link's `fips=` beats its `lat=`/`lon=`, which beat polygons, which
beat `zip=`, which beats `wherestr=`. (That describes how the app reads a URL -- the
`url_ejscreenmap()` R helper arbitrates its *arguments* in its own order when several
are passed, so give it only one input type per call.) A single place opens the site
report popup on it; several places are all drawn and (other than zips) added to
EJScreen's Multisite list, ready for a Multisite Report or the "Send to EJAM" handoff
back into EJAM.

**ZIP code vs county FIPS:** a bare 5-digit number is ambiguous -- 10001 is both a
Manhattan zip code and Kent County, DE's county FIPS. The deep links resolve it
explicitly: `fips=` is always county/tract/blockgroup FIPS, and `zip=` is always a zip
code. A bare 5-digit `wherestr=` follows EJAM's convention: it is tried as a county FIPS
first (drawing and selecting that county), and only geocoded as a zip code if no county
has that code. So prefer `fips=` and `zip=` -- the `fips` and `zip` arguments of
`url_ejscreenmap()` -- while adding context like `?wherestr=10001,NY` also still forces
the zip reading. The interactive search box inside the EJScreen app is unchanged:
typing a bare 5-digit number there still searches it as a zip code.

URL-encoding: percent-encode free-text values (a space becomes `%20`, a comma can be
`%2C`); the digits, commas, and semicolons in the numeric parameters can be left as is.
EJScreen decodes the values on arrival. `url_ejscreenmap()` takes care of the encoding,
and for polygons it also simplifies each outline enough to fit URL length limits (falling
back to a centroid `wherestr=` link if an outline cannot fit):

```{r url_ejscreenmap_examples}
# One county: boundary drawn & selected, site report popup open
# (10001 as a FIPS is Kent County, DE; as a zip code it is in Manhattan, NY)
url_ejscreenmap(fips = "10001")
#> "https://pedp-ejscreen.azurewebsites.net/index.html?fips=10001"

# The same 5 digits as a zip code: pinned in Manhattan, never read as a FIPS
url_ejscreenmap(zip = "10001")
#> "https://pedp-ejscreen.azurewebsites.net/index.html?zip=10001"

# One URL per site (the default; used for per-site link columns in EJAM result tables)
url_ejscreenmap(fips = c("10001", "10003"))

# ONE link that loads ALL the sites into EJScreen's Multisite list
url_ejscreenmap(fips = c("10001", "10003"), combined = TRUE)
#> "https://pedp-ejscreen.azurewebsites.net/index.html?fips=10001,10003"

# Points, with the site report buffer prefilled to 3 miles
url_ejscreenmap(lat = c(39, 39.7), lon = c(-75.5, -75.6), radius = 3, combined = TRUE)
#> "https://pedp-ejscreen.azurewebsites.net/index.html?lat=39,39.7&lon=-75.5,-75.6&radius=3"

# A polygon (outline simplified as needed to fit in a URL)
url_ejscreenmap(shapefile = testinput_shapes_2[1, ])
#> "https://pedp-ejscreen.azurewebsites.net/index.html?polygon=39.0375,-75.5519;39.0392,..."

# A place name via the legacy parameter (free text gets percent-encoded)
url_ejscreenmap(wherestr = "Dover, DE")
#> "https://pedp-ejscreen.azurewebsites.net/index.html?wherestr=Dover%2C%20DE"

# browseURL(url_ejscreenmap(fips = "10001"))  # open one in the browser
```

The selection-drawing parameters need the EJScreen release that added deep links (see
[Public-Environmental-Data-Partners/EJScreen#70](https://github.com/Public-Environmental-Data-Partners/EJScreen/pull/70));
older EJScreen deployments ignore them and just open the map, while the `wherestr=` links
work on all versions. The same parameter vocabulary is documented in the EJScreen
repository's README.

# NOTES ON GLOBAL DEFAULTS, SHINY INPUTS, BOOKMARKS, AND FUNCTION PARAMETERS

This document provides technical notes on how defaults and custom settings are defined for the EJAM web app.

Most of what you might want to do is shown in documentation showing examples of using the `ejamapp()` function. This article is only needed if you want even more detail. First reading the `ejamapp()` reference info is recommended before reading any of the details below.


## HOW SETTINGS GET PASSED TO THE WEB APP

Many defaults are defined in files like `global_defaults_*.R`. They can be changed there, but also can be passed to `ejamapp()` as parameters, to override those global default settings for the duration of the app. Using `ejamapp()` parameters allows experienced analysts to customize how the web app works without having to edit any files. It can be used on a laptop, for example interactively with custom settings. For example, one can change the radius, the type of sites to analyze, default name of the analysis, report title/logo, etc., or to override caps such as limits on file upload size or number of points:

```{r run-app1, eval=FALSE}
ejamapp(
  radius=3.14,
  default_max_miles=31,
  default_max_mb_upload=100
)
```

These custom parameters are "advanced" features for knowledgeable users, and not all the possible changes to these have been tested,
so anyone using them should be careful to understand the details of how this work and confirm it is doing what is expected. 

*For example, the radius is defined like this:* 

1. if a user in RStudio launches the app with something like `ejamapp(radius_default=3.1)`, the setting `radius_default` is defined, and will override what it says in `global_defaults_shiny.R`

2. the files `global_defaults_*.R` are where the settings like `radius_default` are defined and stored as appOptions.

3. that option/setting is checked in the advanced tab UI, to provide the initially selected value of `input$radius_default` in a bookmarkable input control found in the advanced tab. In this case the option and input used the same exact name, "radius_default" but that is not always the case for other settings.

4. in the advanced tab, `input$radius_default` can then be adjusted by a user or just left alone. It provides the initially selected value of `input$radius_now` (which is for the slider UI, but happens to be defined in the server not ui in this case).

5. the radius slider bar will change the value of `input$radius_now` when a user moves the radius slider bar in the UI.

6. the value of `input$radius_now` is what the app uses in calculations when the start analysis button is clicked.

*A setting can be named differently in three places:*

Sometimes there are two or three separate names that affect the same
setting. You only need to know about the first one unless
you are editing URL-encoded bookmarks or editing source code!

1) The global default constant that is the initial value of x,
and is normally defined in a `global_defaults_*.R` file installed with the EJAM package, but
also can be provided as a parameter to `ejamapp()`.

2) for the subset of settings that can be adjusted in the Advanced tab (not all can),
there is a second variable: the input$ one can adjust in that tab.

3) the name of the input$ finally used in the main part of the web app.

The first of these three cannot be saved as a bookmark, but the second or third
can be bookmarked since it is a shiny app input$ variable.

*Steps in EJAM shiny app launch, and the use of params/ defaults/ advanced tab/ bookmarked inputs/etc.:*

### app.R and isPublic=T

If the app is on a server or one uses the RStudio button "Run", app.R is used.
app.R is normally how the app is launched if it is a hosted shiny app on a server.
app.R mostly just does library(EJAM) and ejamapp(isPublic = TRUE).
isPublic is special because it is checked by the global_defaults_shiny_public.R file
and some settings depend on isPublic being T or F.

### ejamapp() parameters

If the app is launched directly via `ejamapp()` by a user in RStudio console, it skips app.R
and a user can provide arguments as parameters in ejamapp(...). See `?ejamapp`

### global_defaults_*.R files

Many defaults/ settings used by the web app are defined in the global_defaults_ files
(global_defaults_package.R, global_defaults_shiny.R, global_defaults_shiny_public.R).
Some settings are needed even if the shiny app is never launched, so they are stored for the package 
as soon as the package is attached, in a list object called `global_defaults_package` in the global environment, e.g., `global_defaults_package$report_title`. (When the app is launched, those and many other global_defaults from the files `global_defaults_*.R`, including radius_default, will be stored in the shiny app).
Note the parameter `isPublic` is special because it is checked by the global_defaults_shiny_public.R file
and some settings depend on `isPublic` being T or F.

### any `ejamapp()` parameters will override global_defaults_*.R settings

Just before launching the app, `ejamapp()` calls `get_global_defaults_or_user_options()` which
gets the global defaults and then overrides/replaces them with
any user-specified versions of those options given as arguments in `ejamapp()`.

### ejamapp() convenience aliases

`ejamapp()` also accepts a number of short **aliases** for parameters and, for site selection,
infers related defaults. For example `pts` (or `lat`+`lon`) is an alias for `sitepoints`, `shp` is
an alias for `shapefile`, and `naics`/`sic`/`mact` are aliases for `default_naics`/etc. Some of these
also set the method automatically -- e.g. passing `fips`, `shapefile`, or `sitepoints` sets
`default_site_method = "upload"` (and the upload sub-type), while `default_naics` sets
`default_site_method = "dropdown"`. See `?ejamapp` and the alias-handling block near the top of
`R/ejamapp.R`.

The aliases and the defaults they cross-set (kept in sync with the alias block near the top of `R/ejamapp.R`):

| You pass to `ejamapp()` | Canonical param | Also cross-sets |
|---|---|---|
| `pts`, or `lat`+`lon` | `sitepoints` | `default_site_method="upload"`, `default_selected_type_of_site_upload="latlon"` |
| `shp` or `shape` | `shapefile` | `default_site_method="upload"`, `default_selected_type_of_site_upload="SHP"` |
| `shapefile` | `shapefile` | `default_site_method="upload"`, `default_selected_type_of_site_upload="SHP"` |
| `fips` | `fips` | `default_site_method="upload"`, `default_selected_type_of_site_upload="FIPS"` |
| `naics` | `default_naics` | `default_site_method="dropdown"`, `default_selected_type_of_site_category="NAICS"`, `default_naics_digits_shown="detailed"` |
| `sic` | `default_sic` | `default_site_method="dropdown"`, `default_selected_type_of_site_category="SIC"` |
| `mact` | `default_mact` | `default_site_method="dropdown"`, `default_selected_type_of_site_category="MACT"` |

The site-selection method param `default_site_method` was formerly named `default_upload_dropdown`;
the old name still works as a back-compat alias (normalized to the new name inside `ejamapp()`).

### app launch

`ejamapp()` launches the app -- it uses a function called `golem::with_golem_options()`, 
which defines the `app` object using `shiny::shinyApp()`, but then
stores all the global defaults and any user-provided parameters as `app$appOptions$golem_options`, 
just before launching the app. Those options can be retrieved in app or server later, 
via `EJAM:::global_or_param()` which relies on `golem::get_golem_options()`.

### global env?

Then app_ui() and app_server() can check for defaults/params using `EJAM:::global_or_param("radius_default")` 
which uses `golem::get_golem_options()` to check for a named option stored in `app$appOptions$golem_options`.
Usually these are used as initial/default values for input$ variables in the app.
However, note if an expected setting is not found (ie not set by user in ejamapp() and not set in any global_defaults_ file)
`global_or_param()` also looks in global env, as a last resort!! That might be a problem if ui or server looks for a
setting that we failed to provide in global_defaults files and that variable name happens to be in the global env !

### Advanced Tab

SOME BUT NOT ALL global_defaults_ settings are in the Advanced Tab (in app_ui)
where they typically provide the default for an input$ but that default can be overridden
there by someone using the app's Advanced Tab,
so what ui or server ends up using could be an input$ that is adjustable in the Advanced Tab
and therefore can be bookmarked (since inputs can be bookmarked unless specifically excluded from that).

An Advanced-Tab setting that should change a *main* control (e.g. the "Site Selection Method"
radio `default_ss_choose_method` setting the main `ss_choose_method` radio) is applied from the
**server** via `updateRadioButtons()` -- you cannot reference `input$...` from `app_ui()`.

### Launch-URL parameters (in addition to the above)

Beyond `ejamapp()` params and the global_defaults_ files, the app can also read settings from the
**launch URL**, via two distinct mechanisms:

- Shiny's native **bookmarking** restores any `input$` value from a `?_inputs_&id=value...` URL
  (see the bookmarking notes above).
- The EJScreen / multisite handoff work adds **custom launch-URL parameters** (e.g. `?lat=&lon=`,
  `?fips=`, `?shape=`, `?handoff=`) that take precedence over the matching `ejamapp()`/global default
  via `url_x() %||% global_or_param("x")`. (See PR #413 on the `ejscreen-multisite-selection` branch.)

### how an input control actually gets its default

Two orthogonal things determine an input control: **(Axis 1) how the widget is built**, and
**(Axis 2) where its value comes from**.

**Axis 1 -- how the widget is built** (a UI-mechanics choice, almost always forced by the widget, not stylistic):

| Pattern | Use when | Example |
|---|---|---|
| **A. static UI default** -- `selected=`/`value=` set in `app_ui.R` to `global_or_param("...")` | the default is a plain value known at page-load, with a fixed choice set | most controls |
| **B. server `renderUI`** (`uiOutput()` in UI + `output$x <- renderUI({...})`) | the widget's *structure* (which choices exist, min/max, conditional sub-controls) depends on reactive state | radius slider (`radius_now`) |
| **C. static-empty UI + server `update*Input()`** | choices load server-side (large selectize lists) OR the value must change on a later event | NAICS/SIC/MACT selectize; advanced-tab -> main-control sync |

Because the UI is built once before any reactive `input$` exists, an Axis-1-A control's `selected=`/`value=`
must be a **non-reactive** value -- `global_or_param("...")` -- never `input$...` (that errors at UI build).
To make an advanced-tab setting change a main control, sync it server-side via `update*Input()` (Axis-1-C).

**Axis 2 -- where the value comes from** (the precedence stack, highest wins):

1. **Shiny bookmark** `?_inputs_&...` -- restores saved `input$` values on load (framework-level; orthogonal to the rest).
2. **Launch-URL custom params** `?lat=&lon=`, `?fips=`, `?shape=`, `?handoff=`, `?radius=` -- read via the `url_*`
   reactives and resolved by `global_or_shinyparam_or_urlparam(name)`, so a launch-URL value wins over the default.
3. **`ejamapp()` params** (after alias normalization) -- override the global-default files.
4. **`global_defaults_*.R`** files (package < shiny < shiny_public).

The advanced-tab -> main-control sync sits *alongside* this stack: it applies a user's runtime change after load
(e.g. the "Site Selection Method" radio updates `ss_choose_method` via `update*Input()`).

To view the Advanced tab in the web app, click the About tab and then the Show advanced button on that page. If that button is not visible, use ejamapp(default_can_show_advanced_settings=TRUE). To show the advanced tab at launch, use ejamapp(default_show_advanced_settings=TRUE).

### web app reactives and interactive inputs

The shiny app uses some settings as initial values for
parameters that can get changed by the user via sliders, radio buttons, text boxes, dropdowns, etc.
and some selections/inputs can get changed by reactive observers, etc.

### Bookmarks (URL or server-based) for saved input$ state

If the app is launched by a bookmark (e.g.URL-encoded bookmark),
any inputs encoded in the URL will override the above and the app will change to
a state where the inputs are as specified by the bookmark.

Bookmarks have limitations though: 
Notably, uploaded files can only be bookmarked
if enableBookmarking="server" not enableBookmarking="url".
Also, bookmarks normally just save the state
of UI inputs, not other reactive values (unless the server code specifies otherwise by 
using the `shiny::onBookmark()` and `shiny::onRestore()` functions to make 
bookmarks store other information, beyond just inputs).
Some important global defaults affect the app but not by providing
initial values for inputs in the shiny app, so some useful information
including some that is controlled by global defaults or user parameters in ejamapp()
cannot be bookmarked because it is stored in reactives
but not as shiny app inputs.
To pass a set of *sites* (points, FIPS codes, or polygons) into an already-deployed
app through a URL — which the file-upload inputs cannot do through a url-bookmark —
use the custom launch query parameters described next.

### Launch URL with pre-selected sites (handoff from an external app)

Separately from Shiny's `?_inputs_&...` bookmark mechanism, the app server reads a
small set of **custom launch query parameters** so an external app (e.g. EJScreen)
can open EJAM already pre-loaded with a set of places. This is the way to pass in
**points or polygons**, which are file uploads and therefore cannot travel in a
url-encoded bookmark. The vocabulary matches `url_ejamapi()`:

- `?lat=33,34&lon=-112,-114` — one or more points (comma-separated), each a site
- `?fips=10001,10003` — one or more FIPS codes, **each a separate site**
- `?shape=<url-encoded GeoJSON>` — a polygon / FeatureCollection (read by `shapefile_from_any()`)
- `?radius=3` (or `?buffer=3`) — analysis radius in miles
- `?handoff=<token>` — a token previously minted by the EJAM API `POST /handoff`;
  the app fetches `GET /handoff/<token>` and pre-loads whatever places it holds.
  Use this for large polygon sets that would not fit in a URL.

Only one place-type is loaded per launch (points, then FIPS, then polygons),
matching the one-method-per-analysis model. Internally the launch handler stores
the parsed places in three per-session reactives — `url_sitepoints`, `url_fips`,
and `url_shapefile` — which `data_up_tablepassed_latlon()`/`data_up_latlon()`,
`data_up_fips()`, and `data_up_shp()` prefer over the `ejamapp()`/global defaults.
It also updates `ss_choose_method` and `ss_choose_method_upload` so the matching
upload tab is shown, so the app opens with those sites selected and ready to run.
The handler is the `observe()` near the top of the SELECT SITES section in
`R/app_server.R`.

For example, this opens the hosted app pre-loaded with two counties (two sites):

    https://ejamapp.ejanalysis.com/?fips=10001,10003

and this pre-loads two points with a 3-mile radius:

    https://ejamapp.ejanalysis.com/?lat=33,34&lon=-112,-114&radius=3

Because the query vocabulary matches `url_ejamapi()` (which builds EJAM API report
URLs), the helper `url_ejamapp()` builds these app-launch URLs with the same
vocabulary -- for example `url_ejamapp(fips = c("10001","10003"))` or
`url_ejamapp(handoff = "<token>")`. The `?handoff=<token>` form is how EJScreen's
"Send to EJAM" button passes a large set of drawn polygons without hitting
URL-length limits:
EJScreen `POST`s the GeoJSON to the EJAM API `/handoff`, gets a token, and opens
the EJAM app at `?handoff=<token>` (e.g. `https://ejamapp.ejanalysis.com/?handoff=<token>`);
the app then fetches those places back from the API's `GET /handoff/<token>` at startup.

--------------
--------------

## EXAMPLES OF USING CUSTOM SETTINGS 


### Launch local app (the R user's computer) with custom settings

See the examples in the documentation of the function `ejamapp()`  

One such example:
```{r run-app, eval=FALSE}
ejamapp(
  default_standard_analysis_title="Custom NAICS Analysis",
  default_site_method="dropdown",
  default_selected_type_of_site_category="NAICS",
  default_naics_digits_shown="detailed",
  default_naics="562211",
  radius_default=3.1,
  default_show_advanced_settings=FALSE
)
```

<!--   not finished   - hidden from article until ready...


Functions to help launch with custom parameters

```{r run-app-with-params, eval=FALSE, }

# Define helper functions -   *************** UNFINISHED DRAFT  -   *************** 

# draft function generator that gives you a custom version of ejamapp()
make_run_custom = function(params = NULL) {
  if (is.list(params)) {params <- unlist(params)}
  txt <- paste0(
    "ejamapp(", paste0(paste0(names(params), " = '", params, "'"), collapse=", "), ")"
  )
  myapp = function() {
    EJAM:::source_this_codetext(txt)  ### ***(((((((((((((( ACTUALLY THIS HAS TO BE EVALUATED BEFOREHAND - FIX THIS  ))))))))))))))
  }
  return(myapp)
}

# function to launch ejamapp() with parameters saved as vector or list
run_with_paramlist = function(params = NULL) {
  fn = make_run_custom(params)
  fn()
}
```

Launch using parameters that are saved in a named vector or list

```{r run-with-paramlist, eval=FALSE}
myparams <- c(
  default_standard_analysis_title="Custom NAICS Analysis",
  default_site_method="dropdown",
  default_selected_type_of_site_category="NAICS",
  default_naics_digits_shown="detailed",
  default_naics="562211",
  radius_default=3.1,
  default_show_advanced_settings=TRUE
)

run_with_paramlist(myparams)
```

Save a custom reusable function you can use to launch with preset parameters

```{r custom-app, eval=FALSE}

myparams <- c(
  default_standard_analysis_title="Custom NAICS Analysis",
  default_site_method="dropdown",
  default_selected_type_of_site_category="NAICS",
  default_naics_digits_shown="detailed",
  default_naics="562211",
  radius_default=3.1,
  default_show_advanced_settings=TRUE
)

# create your reusable customized app launcher
myapp = make_run_custom(myparams)
# save(myapp, file = "myapp.rda")

# use that later to launch app
# load("myapp.rda")
myapp()
```

-->




### Launch hosted app (online, URL) with custom settings

If the web app is running locally or hosted on a server,
launching it via URL-encoded bookmark can provide it any of the input$ variables as a starting state.

This approach cannot provide settings that are not input$ reactives, like some
of the defaults in global_defaults_ that are not turned into inputs in the app,
such as "app_logo_aboutpage" set in global_defaults_package.R and used in app_ui.R
as the src parameter for img().

Also note that url-encoded bookmarked inputs can launch the app
and pass various input$ settings to the app, which would override
the default values of those input$ settings provided by the global_defaults_*.R
and/or `ejamapp()` parameters handled by the utility function
get_global_defaults_or_user_options().

**Example that will launch app with a custom title and radius**

Note this tries to specify the NAICS method, but at least initially the draft code handling this could not specify the NAICS code itself or start analysis via URL bookmark:

Here is an example using the URL of one copy of the hosted app:

Locally hosted app example

```{r, eval=FALSE}
# In 1 R session, launch locally with advanced tab disabled
myport <- 5432
ejamapp(isPublic=TRUE, options = list(port=myport))

# Then you can open a second R session and in that new R session
# paste/use the code below to relaunch the app
# with a parameter that enables the advanced tab:

params <- c('ui_show_advanced_settings=1')
urlbase <- paste0("http://127.0.0.1:", myport, "/?_inputs_&")
urlx <- URLencode(paste0(urlbase, paste0(params, collapse = "&")))
browseURL(urlx)
```


Using a custom URL to go to a web app with custom settings

```{r browseURL, eval=FALSE}
urlbase <- "https://ejam.publicenvirodata.org/?_inputs_&"

myparams <- c(
  standard_analysis_title="Custom NAICS Analysis",
  default_ss_choose_method="dropdown", 
  ss_choose_method_drop="NAICS", 
  naics_digits_shown="detailed", 
  # ss_select_naics=313,  
  radius_default=3.1,
  ui_show_advanced_settings=1
)

params <- c('standard_analysis_title="Custom NAICS Analysis"',
            'ss_choose_method="dropdown"',   # like ejamapp( default_site_method="dropdown" )
            'ss_choose_method_drop="NAICS"', # see ?ejamapp()
            'naics_digits_shown="detailed"', # like ejamapp( default_naics_digits_shown = "detailed")
            # 'ss_select_naics=313',           # like ejamapp(default_naics=313)
            'radius_default=3.1',             # like ejamapp(radius_default = 3.1)
            'ui_show_advanced_settings=1')

urlx <- URLencode(paste0(urlbase, paste0(params, collapse = "&")))
browseURL(urlx)

# URL tried, with fips picked
# http://127.0.0.1:4200/?_inputs_&standard_analysis_title=%22Custom%20COUNTIES%20Analysis%22&pickermoduleid-all_regions_button=%22TRUE%22&pickermoduleid-all_states_button=%22TRUE%22&ui_show_advanced_settings=1&ss_choose_method=%22dropdown%22&fipspicker_done_button=2&ss_choose_method_drop=%22FIPS_PLACE%22&pickermoduleid-fips_type2pick=%22Counties%22&pickermoduleid-counties_picked=%2202110%22
```

--------------
--------------

## COMPILED LISTS OF ALL SETTINGS (FUNCTION PARAMETERS, GLOBAL_DEFAULTS_, AND SHINY INPUT$ VARS)

### > global_defaults_xyz - compiled full list

```{r global-d, eval=TRUE}
gdefs = EJAM:::get_global_defaults_or_user_options()
gdefnames = sort(unique(names(gdefs)))
cbind(global_defaults = gdefnames)
```

### > function params - compiled full list
```{r, eval=TRUE}
params = formals(ejamit) 
# params = c(formals(getblocksnearby), formals(getblocksnearbyviaQuadTree), formals(get_blockpoints_in_shape), formals(getblocksnearby_from_fips), formals(ejamit), formals(doaggregate))
paramnames = sort(unique(names(params)))
cbind(params = paramnames)
```

### > shiny inputs - compiled full list

Compilation of input$  variables used in the shiny app


<!--   needs updating   and  compare to  EJAM:::input_names_listing()    -->


```{r inputs, eval=FALSE}
# getwd() must be the source pkg root folder
getwd()
lines_with_inputnames = trimws(grep("input\\$", gsub("^(.*)(input\\$.*\\))(.*)", "\\2", readLines("./R/app_server.R") ), value=T))
x = gsub("^(.*)(input\\$.*)", "\\2", lines_with_inputnames)
x = gsub("]|)|,|}", " ", x)
x = gsub("(input\\$)([^ ]*)(.*)", "\\1\\2", x)
# cbind(x, lines_with_inputnames)
inputnames = sort(unique(x)); rm(x,lines_with_inputnames)
inputnames = grep("\\.$|\\$$", inputnames, value = TRUE, invert = TRUE)
# cbind(inputnames)
# cbind(part_of_a_named_list = inputnames[grepl("\\$.*\\$", inputnames)])
inputnames =  inputnames[!grepl("\\$.*\\$", inputnames)] # they all also showed up without the extra $xyz part

# What names are used in the shiny app input$ reactives?

cbind(shiny_inputs = inputnames)
inputnames = gsub("^input\\$", "", inputnames)
dput(inputnames)
```

```{r inputnames, eval=TRUE}

inputnames = c("add_naics_subcategories", "all_tabs", "allow_median_in_barplot_indicators", 
"an_map_clusters", "an_thresh_comp2", "an_threshgroup1", "an_threshgroup2", 
"an_threshnames2", "analysis_title", "avoidorphans", "back_to_site_sel", 
"back_to_site_sel2", "bt_get_results", "bysite_webtable_colnames", 
"calculate_ratios", "can_show_advanced_settings", "checkbox1", 
"coauthor_emails", "coauthor_names", "conclusion1", "conclusion2", 
"conclusion3", "Custom_title_for_bar_plot_of_indicators", "default_ss_choose_method", 
"demog_high_at_what_share_of_sites", "demog_how_elevated", "envt_high_at_what_share_of_sites", 
"envt_how_elevated", "epa_program_help", "extra_demog", "extratable_hide_missing_rows_for", 
"extratable_show_ratios_in_report", "extratable_title", "extratable_title_top_row", 
"facilities_studied", "facilities_studied_enter", "fips_help", 
"fipspicker_done_button", "format1pager", "frs_help", "fundingsource", 
"go", "in_areas_where_enter", "in_the_x_zone", "in_the_x_zone_enter", 
"include_ejindexes", "latlon_help", "latlontypedin_submit_button", 
"max_mb_upload", "max_miles", "max_pts_click", "max_pts_map", "max_pts_run", "max_pts_select",
"max_pts_upload", "max_shapes_map", "maxradius", "naics_digits_shown",
"need_proximityscore", "ok2plot", "plotkind_1pager", "radius_default", 
"radius_default_shapefile", "radius_now", "radius_units", "results_tabs", 
"return_to_results", "rg_author_email", "rg_author_name", "rg_enter_miles", 
"rg_enter_sites", "rg_zonetype", "risks_are_x", "shiny.testmode", 
"show_advanced_settings", "show_data_preview", "show_ratios_in_report", 
"shp_help", "sitereport_download_buttons_show", "source_of_latlons", 
"ss_choose_method", "ss_choose_method_drop", "ss_choose_method_upload", 
"ss_select_mact", "ss_select_naics", "ss_select_program", "ss_select_sic", 
"ss_upload_fips", "ss_upload_frs", "ss_upload_latlon", "ss_upload_program", 
"ss_upload_shp", "standard_analysis_title", "subgroups_type", 
"summ_bar_data", "summ_bar_ind", "summ_bar_stat", "summ_hist_bins", 
"summ_hist_data", "summ_hist_distn", "summ_hist_ind", "testing", 
"ui_hide_advanced_settings", "ui_show_advanced_settings", "y"
)
```

### > ALL OVERLAPS/ DIFFS (compiled func params, global defaults, shiny inputs)

```{r overlaps, echo=TRUE}

gdefnames_noprefix = gsub("default_", "", gdefnames)

cat("How many are named the same way in shiny inputs, global_defaults_, and function parameters?\n\n")

length(
  intersect(inputnames, gdefnames_noprefix) # about 28
  )
length(
  intersect(paramnames, gdefnames_noprefix)
  )
length(
  intersect(paramnames, inputnames)
  )
# in all 3 lists
print(
  intersect(paramnames, intersect(inputnames, gdefnames_noprefix))
)

cat("Unique names in shiny inputs, as named \n")
length(
  setdiff(inputnames, c(paramnames, gdefnames_noprefix))
)

cat("Unique names in global_defaults_ (not function parameters or inputs, as named) \n")
length(
  setdiff(gdefnames_noprefix, c(paramnames, inputnames))
)

cat("Unique names in function parameters \n")
length(
  setdiff(paramnames, c(gdefnames_noprefix, inputnames))
)

```

--------------
--------------

## ANNOTATED LISTS OF SOME SETTINGS


### > function params - grouped by function ####

Function parameters whose defaults are or could be included in advanced tab, and allowed by `ejamapp()`.

```{r func-params-by-func, eval=TRUE}

EJAM:::args2(getblocksnearby)

EJAM:::args2(getblocksnearbyviaQuadTree)

EJAM:::args2(get_blockpoints_in_shape)

EJAM:::args2(getblocksnearby_from_fips)

# sitepoints can be a param in ejamapp()
# shapefile  can be a param in ejamapp()
# fips

# radius_donut_lower_edge
# quiet = FALSE

```

### > function params - annotated list of most

```{r funcparams-annotated, eval=TRUE}
# > cbind(formals(ejamit))
EJAM:::args2(ejamit)

# sitepoints               NULL            -- Now can be passed to ejamapp()
# shapefile                NULL           ----   shapefile now can be  passed to ejamapp()
# fips                     NULL            ----   fips  will be enabled as param

# radius                   3   for getblocksnearby(), shiny radius_default is in global_defaults_
# radius_donut_lower_edge  0         na
# maxradius                31.07     for getblocksnearby(), shiny default is set in global
# avoidorphans             FALSE     for getblocksnearby(), shiny default is set in global
# quadtree                 NULL      na

# countcols                NULL      for doaggregate() ** shiny default NOT specified here
# popmeancols              NULL      for doaggregate() ** shiny default NOT specified here
# calculatedcols           NULL      for doaggregate() ** shiny default NOT specified here
# calctype_maxbg           NULL      for doaggregate() ** shiny default NOT specified here
# calctype_minbg           NULL      for doaggregate() ** shiny default NOT specified here

# subgroups_type           "nh"      for doaggregate(), shiny default is set in global
# include_ejindexes        TRUE      for doaggregate(), shiny default is set in global
# calculate_ratios         TRUE      for doaggregate(), shiny default is set in global
# extra_demog              TRUE      for doaggregate(), shiny default is set in global
# need_proximityscore      FALSE     for doaggregate(), shiny default is set in global
# infer_sitepoints         FALSE     for doaggregate() ** shiny default NOT specified here
# need_blockwt             TRUE      for getblocksnearby_from_fips(), shiny default not specified here, but by function defaults

# thresholds          expression  for batch.summarize() - set in global_ but named differently
# threshnames         expression  for batch.summarize() - set in global_ but named differently
# threshgroups        expression  for batch.summarize() - set in global_ but named differently

# updateProgress           NULL      ** shiny default NOT specified here
# updateProgress_getblocks NULL      ** shiny default NOT specified here
# in_shiny                 FALSE     for ejamit(), internal report builders, and related functions
# quiet                    TRUE      for getblocksnearby() and batch.summarize(), na

# testing                  FALSE    for doaggregate(), shiny default is set in global
# silentinteractive        FALSE    for doaggregate() etc. na
# called_by_ejamit         TRUE     for doaggregate(), na
# showdrinkingwater        TRUE     for doaggregate(), na
# showpctowned             TRUE     for doaggregate(), na
# download_city_fips_bounds = TRUE,
# download_noncity_fips_bounds = FALSE,
# ...                      ?   for getblocksnearby(), like use_unadjusted_distance ** shiny default NOT specified here
```


### > shiny inputs - categorized (partial) ####

EJAM shiny app inputs that can be URL-encoded in a bookmark to reopen the app with those settings
Note the URL cuts off  at about 4096 characters

#### inputs: tabs and nav and help buttons

```{r, eval=FALSE}

all_tabs="See Results"   # or # all_tabs="Advanced Settings"
results_tabs="Community Report"
details_subtabs="Site-by-Site Table"

show_advanced_settings="TRUE" 
can_show_advanced_settings="TRUE"

# buttons:
ui_show_advanced_settings=1 #button
ui_hide_advanced_settings=0 #button
show_data_preview=0
back_to_site_sel2=0
return_to_results=0
latlon_help=0
shp_help=0
fips_help=0
frs_help=0
epa_program_help=0
```

#### inputs: start analysis button

```{r, eval=FALSE}
bt_get_results=0
```

#### inputs: site selection (buttons/dropdowns)

```{r, eval=FALSE}

default_ss_choose_method="upload" # or "dropdown" or "mapclick" (click/draw on map)
ss_choose_method="upload"         # or "dropdown" or "mapclick"
ss_choose_method_upload="latlon"   # see global_defaults_*.R for options
mapclick_clear=0                   # "Clear all points" button for the mapclick method
ss_choose_method_drop="FIPS_PLACE" # see global_defaults_*.R for options
ss_upload_shp=null
ss_upload_fips=null
ss_upload_frs=null
ss_upload_program=null
ss_select_program="CAMDBS"
ss_select_naics="313"  # see default_naics
ss_select_sic="2015"   # see default_sic
ss_select_mact="AA" # see default_mact 
naics_digits_shown="basic"
add_naics_subcategories="TRUE"

shinyjs-resettable-ss_upload_latlon={}
shinyjs-resettable-ss_upload_shp={}
shinyjs-resettable-ss_upload_frs={}
shinyjs-resettable-ss_upload_program={}
shinyjs-resettable-ss_upload_fips={}

pickermoduleid-fips_type2pick="Counties"
pickermoduleid-states_picked=null
pickermoduleid-counties_picked=null
pickermoduleid-cities_picked=null
pickermoduleid-regions_picked=["1","2","3","4","5","6","7","8","9","10"]
pickermoduleid-all_regions_button=true
pickermoduleid-all_states_button=true
pickermoduleid-all_counties_button=false
pickermoduleid-all_cities_button=false
pickermoduleid-reset_button=0
fipspicker_done_button=0
```

#### inputs: radius and caps

```{r, eval=FALSE}
radius_now=3.1  # input from slider but not the way to bookmark /url-encode the radius
radius_default=3.1  # works as parameter to ejamapp() or as bookmarked input$ 
radius_default_shapefile=0
max_miles=10
maxradius=31.06856
max_pts_upload=5000
max_pts_select=5000
max_pts_click=30
max_pts_map=5000
max_pts_showtable=1000
max_pts_run=10000
max_shapes_map=159
max_mb_upload=50
```

#### inputs: more ejamit() parameters, summary report and excel output settings, etc.

```{r, eval=FALSE}
standard_analysis_title="CUSTOM Analysis" # shown on webpage of app and on reports
analysis_title="testname"
Custom_title_for_bar_plot_of_indicators=""
include_ejindexes="TRUE"
calculate_ratios=true
include_averages=true
show_ratios_in_report="TRUE"
extratable_show_ratios_in_report="TRUE"
more3="a"
avoidorphans="FALSE"
need_proximityscore="FALSE"
plotkind_1pager="bar"
format1pager="html"
ok2plot=true
allow_median_in_barplot_indicators="FALSE"
summ_hist_distn="People" # histograms
summ_hist_data="pctile"
summ_hist_bins=10

extra_demog="TRUE"
include_extraindicators=true
extratable_title=""
extratable_title_top_row="ADDITIONAL INFORMATION"
extratable_list_of_sections="pctpoor"
# extratable_hide_missing_rows_for=["pcthisp","pctnhba","pctnhaa","pctnhaiana","pctnhnhpia","pctnhotheralone","pctnhmulti","pctnhwa","pctlan_ie","pctlan_api","pctlan_other","pctlan_english","pctlan_spanish","pctlan_french",
# "pctlan_rus_pol_slav","p_chinese","p_korean","pctlan_other_ie","pctlan_vietnamese","pctlan_other_asian","pctlan_arabic","pctlan_nonenglish",
# "pctspanish_li","pctie_li","pctapi_li","pctother_li","pctmale","pctfemale","pctdisability","lowlifex","rateheartdisease","rateasthma","ratecancer","pctunder5","pctunder18","pctover64","occupiedunits","lifexyears",
# "percapincome","pctownedunits","pctpoor","sitecount_avg","sitecount_unique","sitecount_max","distance_min_avgperson","distance_min","count.NPL","count.TSDF","num_waterdis","num_airpoll","num_brownfield","num_tri","num_school",
# "num_hospital","num_church","yesno_tribal","yesno_airnonatt","yesno_impwaters","yesno_cejstdis","yesno_iradis","pctflood","pctfire","pctfire30","pctflood30","yesno_houseburden","yesno_transdis","yesno_fooddesert","pctnobroadband",
# "pctnohealthinsurance","pop","nonmins","age25up","hhlds","unemployedbase","pre1960","builtunits","povknownratio"]

# thresholds - counting how many of a certain indicator type are at/above some threshold
# see help docs on ejamapp() for examples of using these (but they are named a bit differently as parameters there)
an_thresh_comp1=80
an_thresh_comp2=80
an_threshgroup1="EJ-US-or-ST"
an_threshgroup2="Supp-US-or-ST"
an_threshnames1=["pctile.EJ.DISPARITY.dpm.eo","pctile.EJ.DISPARITY.drinking.eo","pctile.EJ.DISPARITY.no2.eo","pctile.EJ.DISPARITY.o3.eo","pctile.EJ.DISPARITY.pctpre1960.eo","pctile.EJ.DISPARITY.pm.eo","pctile.EJ.DISPARITY.proximity.npdes.eo",
                 "pctile.EJ.DISPARITY.proximity.npl.eo","pctile.EJ.DISPARITY.proximity.rmp.eo","pctile.EJ.DISPARITY.proximity.tsdf.eo","pctile.EJ.DISPARITY.rsei.eo","pctile.EJ.DISPARITY.traffic.score.eo","pctile.EJ.DISPARITY.ust.eo",
                 "state.pctile.EJ.DISPARITY.dpm.eo","state.pctile.EJ.DISPARITY.drinking.eo","state.pctile.EJ.DISPARITY.no2.eo","state.pctile.EJ.DISPARITY.o3.eo","state.pctile.EJ.DISPARITY.pctpre1960.eo","state.pctile.EJ.DISPARITY.pm.eo","state.pctile.EJ.DISPARITY.proximity.npdes.eo",
                 "state.pctile.EJ.DISPARITY.proximity.npl.eo","state.pctile.EJ.DISPARITY.proximity.rmp.eo","state.pctile.EJ.DISPARITY.proximity.tsdf.eo","state.pctile.EJ.DISPARITY.rsei.eo","state.pctile.EJ.DISPARITY.traffic.score.eo","state.pctile.EJ.DISPARITY.ust.eo"]
an_threshnames2=["pctile.EJ.DISPARITY.dpm.supp","pctile.EJ.DISPARITY.drinking.supp","pctile.EJ.DISPARITY.no2.supp","pctile.EJ.DISPARITY.o3.supp","pctile.EJ.DISPARITY.pctpre1960.supp","pctile.EJ.DISPARITY.pm.supp","pctile.EJ.DISPARITY.proximity.npdes.supp",
                 "pctile.EJ.DISPARITY.proximity.npl.supp","pctile.EJ.DISPARITY.proximity.rmp.supp","pctile.EJ.DISPARITY.proximity.tsdf.supp","pctile.EJ.DISPARITY.rsei.supp","pctile.EJ.DISPARITY.traffic.score.supp","pctile.EJ.DISPARITY.ust.supp","state.pctile.EJ.DISPARITY.dpm.supp",
                 "state.pctile.EJ.DISPARITY.drinking.supp","state.pctile.EJ.DISPARITY.no2.supp","state.pctile.EJ.DISPARITY.o3.supp","state.pctile.EJ.DISPARITY.pctpre1960.supp","state.pctile.EJ.DISPARITY.pm.supp","state.pctile.EJ.DISPARITY.proximity.npdes.supp",
                 "state.pctile.EJ.DISPARITY.proximity.npl.supp","state.pctile.EJ.DISPARITY.proximity.rmp.supp","state.pctile.EJ.DISPARITY.proximity.tsdf.supp","state.pctile.EJ.DISPARITY.rsei.supp","state.pctile.EJ.DISPARITY.traffic.score.supp","state.pctile.EJ.DISPARITY.ust.supp"]

```

#### inputs: maps

```{r, eval=FALSE}
an_leaf_map_zoom=5
an_leaf_map_bounds={"north":45.61403741135093,"east":-68.02734375000001,"south":28.18824364185031,"west":-123.2666015625}
an_leaf_map_center={"lng":-95.64786455000001,"lat":37.40467238129763}
an_leaf_map_groups="circles"
circleweight_in=4
```

#### inputs: test mode

```{r, eval=FALSE}
testing="FALSE"
shiny.testmode="FALSE"
print_uploaded_points_to_log=true
```

#### inputs: draft / reserved for a long word/pdf report

```{r, eval=FALSE}

# authors

rg_author_name="FirstName LastName"
rg_author_email="author@email.org"
rg_add_coauthors=false
coauthor_names=""
coauthor_emails=""
fundingsource=""

# methods & data

acs_version=""
ejscreen_version=""
subgroups_type="nh"

## sites selected for analysis

prefix_filenames=""
in_the_x_zone="area"
facilities_studied="rule"
facilities_studied_enter=""
rg_enter_sites="facilities in the _____ source category"
source_of_latlons=""
in_areas_where_enter=""
in_the_x_zone_enter="in "
rg_zonetype="zone_is_named_x"
within_x_miles_of="near the"
in_areas_where="in areas with"
risks_are_x="risk is at or above 1 per million (lifetime individual cancer risk due to inhalation of air toxics from this source category)"

# results

demog_how_elevated=""
envt_how_elevated=""
demog_high_at_what_share_of_sites="some of these sites, just as it varies nationwide"
envt_high_at_what_share_of_sites="some of these sites, just as it varies nationwide"
conclusion1=""
conclusion2=""
conclusion3=""

```

--------------
--------------
