---
title: "Updating EJScreen Datasets Annually (via the Pipeline)"
description: "Annual staged workflow for updating ACS-based EJScreen/EJAM datasets"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Updating EJScreen Datasets Annually (via the Pipeline)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

```{r, eval=TRUE, echo=FALSE, include=FALSE}

code_reponame_url        <- EJAM::url_package(type = "code", get_full_url = TRUE)
code_reponame_with_owner <- EJAM::url_package(type = "code", get_full_url = FALSE)
code_reponame <- gsub(".*/", "", code_reponame_with_owner)

data_reponame_url        <- EJAM::url_package(type = "data", get_full_url = TRUE)
data_reponame_with_owner <- EJAM::url_package(type = "data", get_full_url = FALSE)
data_reponame <- gsub(".*/", "", data_reponame_with_owner)

docs_reponame_url        <- EJAM::url_package(type = "docs", get_full_url = TRUE)

```

This document describes the annual staged pipeline for updating the
blockgroup-level EJScreen/EJAM datasets, especially the objects historically
called `blockgroupstats`, `bgej`, `usastats`, and `statestats`.

For general dataset maintenance outside this pipeline, such as FRS-related
tables, NAICS/SIC tables, block-level files, Arrow releases, and package-data
publication mechanics, see [Updating and Managing the Datasets used by
EJAM](dev-update-datasets.html).

## Scope

The pipeline is designed to make each major data step explicit, saved, and
rerunnable. The main entry point is `calc_ejscreen_dataset()`. The recommended
annual runner is:

```{r, eval=FALSE}
source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

The runner writes pipeline checkpoints such as CSV files. It does not, by
itself, replace every installed package dataset in `data/*.rda`. Replacing
package data objects with `usethis::use_data()` or EJAM metadata helpers is a
separate release step after the pipeline outputs have been reviewed.

## Pipeline Stages

The annual workflow creates or reads these stages:

1. `bg_acs_raw`: raw ACS table-based summary file data downloaded from the
   Census Bureau. This is saved before EJAM renaming or formula calculations.
   The default raw ACS storage is a folder of per-table files, with a manifest.

2. `bg_acsdata`: ACS-derived blockgroup indicators calculated from
   `bg_acs_raw`, including demographic indicators and `pctpre1960`.

3. `bg_envirodata`: blockgroup environmental indicators used for EJ indexes.
   This normally comes from a separate environmental-data workflow. For draft
   builds, it can be provisionally reused from the current package data.

4. `bg_extra_indicators`: other blockgroup indicators that are not ACS and not
   the main EJ environmental indicators, such as health, life expectancy, and
   related context variables. These can also be provisionally reused from the
   current package data.

5. `blockgroupstats`: combined blockgroup table with ACS indicators,
   environmental indicators, extra indicators, and geography fields.

6. `usastats_acs`, `statestats_acs`, `usastats_envirodata`, and
   `statestats_envirodata`: percentile lookup tables for ACS and environmental
   inputs.

7. `bgej`: blockgroup EJ index values calculated from demographic indexes and
   environmental percentiles.

8. `usastats_ej` and `statestats_ej`: percentile lookup tables for EJ index
   columns.

9. `usastats` and `statestats`: combined lookup tables used by EJAM.

10. `ejscreen_export`: EJScreen-ready export combining `blockgroupstats` and
    `bgej`, applying EJScreen-style names from `map_headernames`, and adding
    map helper fields where possible.

The runner also writes `pipeline_validation_summary.csv`. If the EJScreen
export is requested, it writes `ejscreen_export_schema_report.csv`.

The pipeline uses the packaged `formulas_ejscreen_acs` object for ACS-derived
indicator formulas. The old
`data-raw/archived_datacreate_formulas_ejscreen_acs_notes.R` file is reference
material only. It is not the current formula rebuild workflow.

## Storage

The default local pipeline folder is:

```{r, eval=FALSE}
data-raw/pipeline_outputs/ejscreen_acs_2024
```

`data-raw/pipeline_outputs/` is ignored by Git because checkpoint files can be
large. The repository also has build-ignore rules for pipeline outputs and
Arrow data files, so release artifacts should normally be stored outside the
package source tree.

The pipeline can also use AWS S3. S3 support uses the AWS CLI, so `aws` must be
installed and configured before running an S3-backed pipeline.

```{r, eval=FALSE}
Sys.setenv(
  AWS_PROFILE = "ejam",
  AWS_REGION = "us-east-1",
  EJAM_PIPELINE_DIR = "s3://pedp-data-preserved/ejscreen-data-processing/pipeline/ejscreen_acs_2024",
  EJAM_PIPELINE_STORAGE = "s3"
)
```

For local testing, use a local directory:

```{r, eval=FALSE}
Sys.setenv(
  EJAM_PIPELINE_DIR = file.path(
    getwd(),
    "data-raw",
    "pipeline_outputs",
    "ejscreen_acs_2024"
  ),
  EJAM_PIPELINE_STORAGE = "local"
)
```

## Key Settings

The runner is controlled by environment variables:

| Variable | Purpose |
|---|---|
| `EJAM_PIPELINE_YR` | ACS 5-year end year, such as `"2024"` for ACS 2020-2024. |
| `EJAM_PIPELINE_DIR` | Local folder or `s3://...` pipeline location. |
| `EJAM_PIPELINE_STORAGE` | `"auto"`, `"local"`, or `"s3"`. |
| `AWS_PROFILE`, `AWS_REGION` | Used by the AWS CLI for S3-backed runs. |
| `CENSUS_API_KEY` | Used by ACS/Census download helpers where needed. |
| `EJAM_FORCE_ACS` | `"TRUE"` to redownload raw ACS and rebuild ACS stages. |
| `EJAM_FORCE_BG_ACSDATA` | `"TRUE"` to rebuild `bg_acsdata` from saved raw ACS. |
| `EJAM_ACS_DOWNLOAD_TIMEOUT` | Download timeout in seconds. Useful for large ACS tables. |
| `EJAM_ACS_DOWNLOAD_RETRIES` | Number of retry attempts for ACS downloads. |
| `EJAM_USE_PROVISIONAL_BG_ENVIRODATA` | `"FALSE"` to require a supplied `bg_envirodata.csv`. |
| `EJAM_INCLUDE_EJSCREEN_EXPORT` | `"TRUE"` to create `ejscreen_export.csv`. |

To see what the runner will use:

```{r, eval=FALSE}
Sys.getenv(c(
  "EJAM_PIPELINE_YR",
  "EJAM_PIPELINE_DIR",
  "EJAM_PIPELINE_STORAGE",
  "AWS_PROFILE",
  "AWS_REGION",
  "CENSUS_API_KEY",
  "EJAM_FORCE_ACS",
  "EJAM_FORCE_BG_ACSDATA",
  "EJAM_ACS_DOWNLOAD_TIMEOUT",
  "EJAM_ACS_DOWNLOAD_RETRIES",
  "EJAM_USE_PROVISIONAL_BG_ENVIRODATA",
  "EJAM_INCLUDE_EJSCREEN_EXPORT"
))
```

## New or renamed indicators

### map_headernames

If new indicators are being used compared with prior version of the datasets and package, 
note that map_headernames may need to be modified to have metadata on those new indicators, 
including the variable name (rname column of map_headernames), longname, calculation type, weight,
rounding information, and "varlist" which groups related indicators into, e.g., names_e, and
demographic key indicators into names_d, etc. 

### names_*

Much of the code depends on the varlist info,
so names_e, names_d, etc. need to be updated. The script in datacreate_names_of_indicators
uses map_headernames$varlist to update these data objects.


## Run a Fresh ACS Update

Start from a clean branch. For ACS 2020-2024 using local checkpoints:

```{r, eval=FALSE}
yr <- "2024"

Sys.setenv(
  EJAM_PIPELINE_YR = yr,
  EJAM_PIPELINE_DIR = file.path(
    getwd(),
    "data-raw",
    "pipeline_outputs",
    paste0("ejscreen_acs_", yr)
  ),
  EJAM_PIPELINE_STORAGE = "local",
  EJAM_FORCE_ACS = "TRUE",
  EJAM_FORCE_BG_ACSDATA = "TRUE",
  EJAM_ACS_DOWNLOAD_TIMEOUT = "3600",
  EJAM_ACS_DOWNLOAD_RETRIES = "2",
  EJAM_INCLUDE_EJSCREEN_EXPORT = "TRUE"
)

source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

For the S3-backed pipeline:

```{r, eval=FALSE}
yr <- "2024"

Sys.setenv(
  AWS_PROFILE = "ejam",
  AWS_REGION = "us-east-1",
  EJAM_PIPELINE_YR = yr,
  EJAM_PIPELINE_DIR = paste0(
    "s3://pedp-data-preserved/ejscreen-data-processing/pipeline/ejscreen_acs_",
    yr
  ),
  EJAM_PIPELINE_STORAGE = "s3",
  EJAM_FORCE_ACS = "TRUE",
  EJAM_FORCE_BG_ACSDATA = "TRUE",
  EJAM_ACS_DOWNLOAD_TIMEOUT = "3600",
  EJAM_ACS_DOWNLOAD_RETRIES = "2",
  EJAM_INCLUDE_EJSCREEN_EXPORT = "TRUE"
)

source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

The runner asks for confirmation in interactive sessions before continuing.

## Rerun From Saved ACS Data

If raw ACS has already been downloaded, rerun downstream ACS calculations
without redownloading:

```{r, eval=FALSE}
Sys.setenv(
  EJAM_FORCE_ACS = "FALSE",
  EJAM_FORCE_BG_ACSDATA = "TRUE"
)

source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

If both raw ACS and `bg_acsdata` should be reused, leave both force flags false:

```{r, eval=FALSE}
Sys.setenv(
  EJAM_FORCE_ACS = "FALSE",
  EJAM_FORCE_BG_ACSDATA = "FALSE"
)

source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

## Supplying Updated Environmental Data

The environmental stage is intentionally separate from the ACS stage. When
updated environmental indicators are available, save them as `bg_envirodata.csv`
in the pipeline folder.

For a local pipeline:

```{r, eval=FALSE}
file.path(pipeline_dir, "bg_envirodata.csv")
```

For an S3 pipeline:

```{r, eval=FALSE}
s3://pedp-data-preserved/ejscreen-data-processing/pipeline/ejscreen_acs_2024/bg_envirodata.csv
```

The file must include `bgfips` and the environmental indicators used for EJ
indexes. It should also include `pctpre1960`. The environmental-data workflow
may create `pctpre1960` by reading the saved `bg_acsdata.csv` stage.

To force the runner to stop unless `bg_envirodata.csv` has been supplied:

```{r, eval=FALSE}
Sys.setenv(EJAM_USE_PROVISIONAL_BG_ENVIRODATA = "FALSE")
source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

After replacing `bg_envirodata.csv`, rerun without forcing ACS:

```{r, eval=FALSE}
Sys.setenv(
  EJAM_FORCE_ACS = "FALSE",
  EJAM_FORCE_BG_ACSDATA = "FALSE",
  EJAM_USE_PROVISIONAL_BG_ENVIRODATA = "FALSE"
)

source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

This reuses the saved ACS stages and regenerates downstream
`blockgroupstats`, `bgej`, `usastats`, `statestats`, and `ejscreen_export`.

## Extra Indicators

Some `blockgroupstats` columns are not ACS indicators and are not the main
environmental indicators. Examples include life expectancy, health indicators,
facility-count context variables, climate-related fields, and other columns
grouped in `map_headernames$varlist`.

The pipeline makes these explicit in `bg_extra_indicators.csv`. If an updated
table is not supplied, the runner currently creates a provisional version from
the packaged `EJAM::blockgroupstats`. That is useful for testing the ACS update,
but it should be documented clearly as reuse of older non-ACS data.

## Provisional Draft Builds

For early pipeline testing, it is acceptable to reuse existing environmental and
extra indicators:

```{r, eval=FALSE}
Sys.setenv(EJAM_USE_PROVISIONAL_BG_ENVIRODATA = "TRUE")
source("data-raw/run_ejscreen_acs2024_pipeline.R")
```

The runner writes source-note text files next to provisional stages, such as
`bg_envirodata_SOURCE.txt` and `bg_extra_indicators_SOURCE.txt`. Final release
review should confirm whether any provisional stage remains.

## Reviewing Outputs

Start with the validation summary:

```{r, eval=FALSE}
library(data.table)

pipeline_dir <- "data-raw/pipeline_outputs/ejscreen_acs_2024"

validation <- fread(file.path(pipeline_dir, "pipeline_validation_summary.csv"))
validation[, .(stage, rows, columns, errors, warnings)]
validation[nzchar(errors)]
validation[nzchar(warnings)]
```

There should be no validation errors. Warnings should be understood and either
fixed or explicitly accepted for a draft build.

Then inspect the main outputs:

```{r, eval=FALSE}
blockgroupstats <- fread(file.path(pipeline_dir, "blockgroupstats.csv"))
bgej            <- fread(file.path(pipeline_dir, "bgej.csv"))
usastats        <- fread(file.path(pipeline_dir, "usastats.csv"))
statestats      <- fread(file.path(pipeline_dir, "statestats.csv"))

nrow(blockgroupstats)
nrow(bgej)
names(blockgroupstats)
names(bgej)
```

Useful checks include:

- expected FIPS/geography columns are present and typed as character;
- row counts are plausible for the ACS vintage and geography coverage;
- key ACS indicators are non-missing for most populated blockgroups;
- percentage/rate variables are in the expected range;
- `blockgroupstats` and `bgej` join cleanly by `bgfips`;
- lookup tables include `REGION`, `PCTILE`, `0`, `100`, and `mean`;
- `usastats` has one region, `"USA"`;
- `statestats` has expected state/territory regions.

For example:

```{r, eval=FALSE}
stopifnot("0" %in% as.character(usastats$PCTILE))
stopifnot("100" %in% as.character(usastats$PCTILE))
stopifnot("mean" %in% as.character(usastats$PCTILE))

stopifnot("0" %in% as.character(statestats$PCTILE))
stopifnot("100" %in% as.character(statestats$PCTILE))
stopifnot("mean" %in% as.character(statestats$PCTILE))

stopifnot(!anyDuplicated(blockgroupstats$bgfips))
stopifnot(!anyDuplicated(bgej$bgfips))
```

## Reviewing the EJScreen Export Schema

If `EJAM_INCLUDE_EJSCREEN_EXPORT` is true, the runner writes:

```{r, eval=FALSE}
ejscreen_export.csv
ejscreen_export_schema_report.csv
```

Use the schema report as a field-by-field checklist:

```{r, eval=FALSE}
schema <- fread(file.path(pipeline_dir, "ejscreen_export_schema_report.csv"))

schema[, .N, by = status]
schema[status == "missing_expected"]
schema[status == "missing_expected", .N, by = field_type][order(-N)]
schema[status == "unexpected_output"]
```

Each missing expected field should be classified as one of:

1. a field EJScreen needs and the export must add;
2. a metadata mapping issue in `map_headernames`;
3. a deliberately deferred field that is not needed for the current export.

For release, the ideal schema report has no `missing_expected` rows for the
FeatureServer fields required by the EJScreen app.

## Replacing Package Data

Pipeline stage files are review artifacts. After they are accepted, update the
package data objects deliberately. The runner currently has an interactive
helper path for replacing `blockgroupstats`, but it does not automatically
replace every final package dataset.

A release update should explicitly replace at least:

- `blockgroupstats`;
- `bgej`;
- `usastats`;
- `statestats`;
- any related lookup or metadata objects that changed.

Use the established EJAM metadata helpers before saving package data. A typical
manual pattern is:

```{r, eval=FALSE}
blockgroupstats <- fread(file.path(pipeline_dir, "blockgroupstats.csv"))
bgej            <- fread(file.path(pipeline_dir, "bgej.csv"))
usastats        <- fread(file.path(pipeline_dir, "usastats.csv"))
statestats      <- fread(file.path(pipeline_dir, "statestats.csv"))

EJAM:::metadata_add_and_use_this("blockgroupstats")
EJAM:::metadata_add_and_use_this("bgej")
EJAM:::metadata_add_and_use_this("usastats")
EJAM:::metadata_add_and_use_this("statestats")
```

Confirm the exact metadata values before saving, especially ACS vintage,
Census vintage, EJSCREEN/EJAM version, data source notes, and provisional reuse
notes.

After those key datasets are updated, rerun the scripts that create and saves
testoutput_* files and datasets, especially `datacreate_testpoints_testoutputs.R`
and `datacreate_testoutput_ejamit_*`, etc.

After package data are updated, reinstall the package and rerun release-critical
tests. Also regenerate any Arrow-format files used outside the package build,
if needed.

## Release Checklist

Before releasing a new annual dataset build:

1. Run the pipeline with the intended ACS year and storage backend.
2. Confirm whether `bg_envirodata` and `bg_extra_indicators` are updated or
   provisional.
3. Review `pipeline_validation_summary.csv`.
4. Review row counts, missingness, ranges, and joins for final tables.
5. Review `ejscreen_export_schema_report.csv`.
6. Replace package `.rda` datasets only after the stage files are accepted.
7. Update dataset metadata and documentation.
8. Rebuild documentation and pkgdown.
9. Reinstall the package and run focused tests plus package checks.
10. Publish large artifacts through the chosen storage path, such as S3 or the
    data repository release process.

The general release and large-data publication steps are covered in [Updating
and Managing the Datasets used by EJAM](dev-update-datasets.html) and [Updating
the Package as a New Release](dev-update-package.html).
