---
title: "Using the EJAM API"
description: "A brief overview of the hosted EJAM REST API and how to reach it from R"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using the EJAM API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 100
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
```

This is a short overview. For the authoritative and most current details, see:

- the **[EJAM-API repository](https://github.com/Public-Environmental-Data-Partners/EJAM-API)**
  and its [README](https://github.com/Public-Environmental-Data-Partners/EJAM-API?tab=readme-ov-file#ejam-api) -- it documents every endpoint, parameter, and example;
- the help for **`?ejamapi`** and **`?url_ejamapi`**; and
- the [Defaults and Custom Settings for the Web App](dev-app-settings.html) article, for the
  *deep links* (launch-URL parameters + token handoff) that pre-load the EJAM web app.

## What the API is

Besides the R package and the web app, EJAM is available as a hosted **REST API** that
returns EJScreen-style community reports (and the underlying data) for a point, a set of
points, a polygon, or US Census areas (FIPS). It runs the same engine as the package, so an
API report matches what `ejamit()` / `ejam2report()` would produce locally.

## Reaching it from R

```{r api, eval=TRUE}
## API example: build a multisite report URL
urlx = EJAM::url_ejamapi(fips=c(3650617,'0973000'), sitenumber = 0)
urlx

# browseURL(urlx)
```

Two functions wrap the API:

- **`ejamapi()`** -- builds the request, calls the API, and returns the report or data.
- **`url_ejamapi()`** -- builds the API request URL(s) *without* calling them (useful for a
  table of shareable links, or to open one in a browser).

```r
ejamapi(lat = 33, lon = -112, radius = 3)              # get a report or data back
url_ejamapi(lat = c(33, 34), lon = c(-112, -114))      # just build the URL(s)
```

The API base URL -- and a friendlier branded alias such as `https://api.ejanalysis.com` --
are described in the
[EJAM-API README](https://github.com/Public-Environmental-Data-Partners/EJAM-API). You do not
normally need to set the base URL yourself.

## Deep linking: launch the app pre-loaded

Separately from the data API, the EJAM web app can be **launched pre-loaded with a set of
places** by opening it with launch query parameters in the URL -- the basis for the EJScreen
"Send to EJAM" hand-off, which moves a selection straight into the full app. Points and FIPS
codes travel directly in the URL; larger sets and drawn polygons use a token **handoff**
(`POST /handoff` returns a token, and the app fetches the places back from
`GET /handoff/<token>` at startup) so they are not limited by URL length. Use
`url_ejamapp()` to build these launch URLs from R.

For the launch-URL vocabulary and the handoff endpoints, see the
[Defaults and Custom Settings for the Web App](dev-app-settings.html) article and the
[EJAM-API README](https://github.com/Public-Environmental-Data-Partners/EJAM-API).

## A few notes on parameters

- Provide one input type per request: points (`lat`/`lon`), **or** `fips`, **or** `shape`.
- `buffer` is a synonym for `radius`; `fileextension` is `html` or `pdf`.
- For an aggregate **multisite** report over several places at once, pass `sitenumber=0`
  (the default for a single site is `sitenumber=1`; a positive `N` reports on just the Nth
  site). Many or large polygons that would exceed a GET URL's length can be sent via a
  `POST /report` request. See the parameter notes in the
  [EJAM-API README](https://github.com/Public-Environmental-Data-Partners/EJAM-API).

For anything not covered here, the
[EJAM-API README](https://github.com/Public-Environmental-Data-Partners/EJAM-API) is the
source of truth, and `?ejamapi` / `?url_ejamapi` document the R interface.
