---
title: "Updating and Managing the Datasets Used by EJAM"
description: "Managing datasets used by EJAM, including Arrow files"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Updating and Managing the Datasets Used by EJAM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

```{r, eval=TRUE, echo=FALSE, include=FALSE}

code_reponame_url        <- EJAM::url_package(type = "code", get_full_url = TRUE)  # "https://github.com/OWNER/REPONAME"
code_reponame_with_owner <- EJAM::url_package(type = "code", get_full_url = FALSE) # "OWNER/REPONAME"
code_reponame <- gsub(".*/", "", code_reponame_with_owner)    #       "REPONAME"

data_reponame_url        <- EJAM::url_package(type = "data", get_full_url = TRUE)  # "https://github.com/OWNER/REPONAME"
data_reponame_with_owner <- EJAM::url_package(type = "data", get_full_url = FALSE) # "OWNER/REPONAME"
data_reponame <- gsub(".*/", "", data_reponame_with_owner)    #       "REPONAME"

docs_reponame_url        <- EJAM::url_package(type = "docs", get_full_url = TRUE)  # "https://OWNER.github.io/REPONAME"
# only the full URL is useful in this case

```

The EJAM package and [Shiny](https://shiny.posit.co) app use many data
objects, including numerous datasets stored in the package's `/data/` folder
and several large tables stored in a separate data repository. Those large
tables contain information on Census block groups, Census block internal
points, Census block population weights, and EPA FRS facilities.

## How to Update Datasets in EJAM

The process begins from within the EJAM code repo. Historically, most data
updates were coordinated from the overarching notes and script file
`data-raw/datacreate_0_UPDATE_ALL_DATASETS.R`. That file is still useful as an
index of older maintainer workflows, but it is no longer the primary path for
the annual EJScreen-style blockgroup update. For `blockgroupstats`,
`usastats`, `statestats`, `bgej`, and related annual pipeline checkpoints, use
the staged pipeline documented in [Updating EJScreen Datasets Annually (via the
Pipeline)](dev-update-ejscreen-datasets-yearly.html). For other datasets, the
focused `datacreate_*` scripts in `data-raw/` remain the usual starting point.
Documentation of datasets via `/R/data_*.R` files is generally handled by those
same scripts while creating/updating the datasets.

That file covers not only the large Arrow datasets that are stored in a separate
repository, but also many smaller data objects that are installed along with the
package in the `/data/` folder. Updating all the package's data objects can be
complicated because there are many different data objects of various types and
formats and locations.

The various data objects need to be updated at various frequencies -- some only
yearly, some as part of the broader EJSCREEN Annual Data Update of demographic,
environmental, and other indicators, and others when facility IDs and locations
change (as often as possible, as when [EPA's FRS](https://www.epa.gov/frs) is
updated). Some need to be updated only when the package features/code changes,
such as the important data object called `map_headernames` (which in turn is
used to update objects such as `names_e`).

See the draft utility `EJAM:::pkg_data()` for a dataset inventory:

```{r}
x <- EJAM:::pkg_data()
x$Item[!grepl("names_|^test", x$Item)]
```

### Where the datasets are stored

EJAM relies on datasets mostly stored in the package itself or in a separate,
data-related repository:

-   Datasets stored within the EJAM package (`.rda` files): [Documentation](../reference/index.html#datasets-with-indicators-raw-data-means-percentiles-) and
[access to package data files](`r paste0(code_reponame_url, "/tree/main/data")`)

-   Datasets used by EJAM but stored separately (large `.arrow` files): [Documentation](../reference/index.html#datasets-with-indicators-raw-data-means-percentiles-) and
[access to the large data files as GitHub release assets](`r paste0(data_reponame_url, "/releases")`)

### Why the large datasets are put into the data repository using piggyback instead of committed using Git

As explained in the documentation for the [piggyback R package](https://docs.ropensci.org/piggyback/):

"Because larger (> 50 MB) data files cannot easily be committed to git, a different approach is required to manage data associated with an analysis in a GitHub repository. This package provides a simple work-around by allowing larger (up to 2 GB) data files to piggyback on a repository as assets attached to individual GitHub releases. These files are not handled by git in any way, but instead are uploaded, downloaded, or edited directly by calls through the GitHub API. These data files can be versioned manually by creating different releases. This approach works equally well with public or private repositories. Data can be uploaded and downloaded programmatically from scripts. No authentication is required to download data from public repositories."

### Key datasets

Some notable data files, code details, and other objects that may need to be
changed ANNUALLY or more often:

-   ***Blockgroup Datasets (Demographic and Environmental Data)***: These
include datasets included with the package `?blockgroupstats`,
[usastats](../reference/usastats.html), `?statestats`, and `?bgej`. The
annual staged workflow for updating these ACS/EJScreen-style blockgroup
datasets is now documented separately in [Updating EJScreen Datasets Annually
(via the Pipeline)](dev-update-ejscreen-datasets-yearly.html). That pipeline
covers `bg_acs_raw`, `bg_acsdata`, optional Island Areas checkpoints,
`bg_envirodata`, `bg_extra_indicators`, `bg_geodata`, `blockgroupstats`,
`bgej`, `usastats`, `statestats`, `ejscreen_export`, and the optional
`ejscreen_dataset_creator_input` stage. This more general vignette focuses on
the other datasets and storage/release mechanics used by EJAM. For EJAM
v2.5.0, AS/GU/MP/VI are included at the blockgroup dataset, EJSCREEN export,
and map-data visibility level with demographic fields kept as `NA` and partial
EPA environmental fields where available. See
[Island Areas in EJAM v2.5.0](island-areas.html) for the user-facing coverage
notes and live EJSCREEN layer inventory observed in May 2026.

-   ***Block Datasets***: The *block* (not blockgroup) tables might be updated
less often, but Census FIPS codes do change yearly so the `?blockwts`,
`?blockpoints`, `?quaddata`, `?blockid2fips`, and related additional
data tables should be updated as needed. This is also done from within
*`/data-raw/datacreate_0_UPDATE_ALL_DATASETS.R`*. See the
`census2020download` package on GitHub for the function `census2020_get_data()` that
may be useful.

```{r notes, eval=FALSE, include=FALSE}
# note that any URLs created in this .Rmd file that point to
# something like articles/ejscreen-map-descriptions.html
# could be created here either using the function EJAM::url_package(type = "docs", get_full_url = TRUE) which defined docs_reponame_url above, like this:
#  (`r paste0(docs_reponame_url, "/articles/ejscreen-map-descriptions.html" )`)
# or as simply the relative path from this article to the other article like this:
#  (ejscreen-map-descriptions.html)
```

-   ***Facilities Datasets for creating updated proximity scores each year***:
Facility (and roadway) locations for key types of sites were used once a
year to update several [environmental indicators that are proximity
scores in EJSCREEN](`r paste0(docs_reponame_url, "/articles/ejscreen-map-descriptions.html#environmental-burden-indicators")`).
The resulting environmental indicators are stored with EJAM, but these
facility location datasets are not stored in EJAM. EJSCREEN obtains their
locations for mapping purposes, via an API accessing [hosted datasets with
facility locations](https://geopub.epa.gov/arcgis/rest/services/EMEF/efpoints/MapServer).
In general, scripts for updating environmental indicators (including
documentation of sources of facility location data, etc.) [were stored by
EPA](https://github.com/Public-Environmental-Data-Partners/EJSCREEN-Data-Processing).
After 2025, new code for updating indicators may be found in this package's
`data-raw/` folder or in related non-EPA source repositories. Proximity scores in EJSCREEN as of
2024-2026 were calculated based on the locations of these types of sites:

-   [Major roadways (traffic)](`r                         paste0(docs_reponame_url, "/articles/ejscreen-map-descriptions.html#traffic-proximity-and-volume"                                  )`)
-   [Superfund NPL sites](`r                              paste0(docs_reponame_url, "/articles/ejscreen-map-descriptions.html#environmental-burden-indicators"                               )`)
-   [Facilities with hazardous waste (TSDF)](`r           paste0(docs_reponame_url, "/articles/ejscreen-map-descriptions.html#hazardous-waste-proximity"                                     )`)
-   [Water bodies downstream of wastewater discharges](`r paste0(docs_reponame_url, "/articles/ejscreen-map-descriptions.html#wastewater-discharge-stream-proximity-and-toxic-concentration" )`)
-   [Risk management plan (RMP) facilities](`r            paste0(docs_reponame_url, "/articles/ejscreen-map-descriptions.html#risk-management-program-rmp-facility-proximity"                )`)
-   [Underground storage tanks (UST)](`r                  paste0(docs_reponame_url, "/articles/ejscreen-map-descriptions.html#underground-storage-tanks-ust"                                 )`)
(for a facility density indicator, similar to a proximity indicator)

-   ***Facilities Datasets for a user to specify places to analyze/report on***:

Facility locations and categories are used in EJAM to help a user specify
sets of EPA-regulated facilities or other types of sites to analyze and
report on in EJSCREEN reports, using their *NAICS/SIC/MACT/program*
information and coordinates. All of that information may need frequent
updates because facilities open, close, relocate, or have their
information corrected or otherwise updated. EPA's FRS is the source for much
of this information and the FRS is updated by EPA frequently and is
available via an API. Through at least v2.32.8, EJAM (and therefore the
community reports in EJSCREEN) used a snapshot of the EPA FRS data rather
than using an API to obtain the latest info on demand -- that is something
that could be changed in a future version. Facility-related info is stored
in tables EJAM uses, such as these: `?frs`, `?frs_by_programid`,
`?frs_by_naics`, `?frs_by_sic`, `?frs_by_mact`, `?NAICS`, `?SIC`,
`?naics_counts`, `?naicstable`, `?SIC`, `?sictable`, `?mact_table`, and
`?epa_programs`, `?frsprogramcodes`, etc. These FRS, MACT, and Program info
tables of EPA-relevant data have been updated in the EJAM package from scripts
within *`/data-raw/datacreate_0_UPDATE_ALL_DATASETS.R`*. The `?NAICS`,
`?naicstable`, and `?sictable` objects (viewable using `naics_categories()`
and `sic_categories()` utilities) have no EPA-specific data so they do not
need frequent updates. The NAICS data object stores just the name of each
NAICS code number, and new codes/names are published every five years, such
as in 2017 and 2022, so a new version would typically be expected in 2027.
The tables called `?SIC` (unlike the NAICS table) and `?naics_counts` (which
has no analogous sic version), however, contain counts of EPA FRS
facilities, so they need updates when FRS data are updated. The
inconsistency in how NAICS vs SIC tables and the naics_counts table were
named and defined was by historical accident, not intentional, so it would
be OK if refactoring later made them consistent or even switched entirely to
more frequent automated updates or even reliance on the FRS API.

-   `?map_headernames` stores critical metadata. This needs to be updated
especially if indicator names change or are added. `?map_headernames` holds
most of the useful metadata about each variable (each indicator, like % low
income) -- e.g., how many digits to use in rounding, units, long and short
indicator names, EJAM and EJScreen field names, the type or category of
indicator, sort order to use in reports, and the method of calculating
aggregations of the indicator over blockgroups. The editable source is now
`data-raw/map_headernames.csv`. If metadata rows or values need to change, edit
that CSV directly, then source `data-raw/datacreate_map_headernames.R` to
validate the CSV and save `data/map_headernames.rda`. Older `.xlsx` workflows
are obsolete and should not be used to regenerate this object.

-   [Test data (inputs) and examples of outputs](testdata.Rmd) may have to be
updated (every time parameters change & when outputs returned change). Those
are generated by scripts/functions referred to from
*`/data-raw/datacreate_0_UPDATE_ALL_DATASETS.R`*

-   A default year is used in various functions, such as for the last year of
the 5-year ACS dataset. These defaults like yr or year should be updated via
global searches where relevant.

-   Metadata about vintage/version is stored in attributes of many datasets.
That metadata is updated via scripts/functions that call helpers such as
`metadata_add()`, `metadata_add_and_use_this()`, `metadata_check()`, and
`metadata_mapping.R`. For staged EJScreen annual outputs, the pipeline save
helpers add the relevant metadata based on the requested pipeline year. After
package data are replaced, run `EJAM:::metadata_check()` and
`EJAM:::metadata_check_print()` to find stale attributes. Atomic name-vector
objects such as many `names_*` datasets do not need metadata attributes.

-   Version numbering is recorded primarily in the DESCRIPTION file, release
tags, and the NEWS file. The `ejamdata_required_tag` field in DESCRIPTION
records which `ejamdata` release EJAM should use. The
`ejamdata_version.txt` marker records which `ejamdata` release tag is actually
saved in the local data folder.

-   [Updating documentation](dev-update-documentation.html) - updates may be
needed for the README, vignettes, and possibly examples in some functions in
case updates to datasets alter how the examples would work.

Again, for non-pipeline datasets it is useful to understand
`data-raw/datacreate_0_UPDATE_ALL_DATASETS.R`, because that script still points
to many older focused data-creation scripts. For annual EJScreen-style
blockgroup outputs, use the pipeline vignette and runner script as the current
maintainer workflow.

The information below focuses on the other type of data objects -- the set of
large `{arrow}` files that are stored outside the package code repository.

### Repository that stores the large arrow file release assets

Several large `{data.table}` files are not installed as part of the R package in
the typical `/data/` folder that contains `.rda` files lazy-loaded by the
package. Instead, they are kept as release assets in a separate GitHub
repository that we refer to here as the data repository. The release assets
are the authoritative copies used by installed EJAM packages; committed files
in a repository `data/` folder should not be treated as the source used by EJAM
installs.

*IMPORTANT:* The name of the *data* repository (as distinct from the *package
code* repository) must be recorded/updated in the EJAM package `DESCRIPTION`
file, so that the package knows where to look for the data files if the
datasets are moved to a new repository. The current data repository for the
installed or loaded source version is `r data_reponame_url`, which can be
checked with `url_package(type = "data", get_full_url = TRUE)`.

### Arrow Package and Arrow File Format

To store the large files needed by the EJAM package, we use the Apache arrow
file format through the `{arrow}` R package, with file extension `.arrow`. This
allows us to work with larger-than-memory data and store it outside of the EJAM
package itself.

Earlier versions of EJAM used the `.arrow` filename more loosely. Current
dynamic datasets should be real Arrow IPC files. For example, the object called
`frs_arrow` is the Arrow-backed version of what had been called the `?frs`
dataset.

The names of these tables should be listed in `R/arrow_ds_names.R` and in the
global variable called `.arrow_ds_names`, which is used by functions like
`dataload_dynamic()` and `dataload_from_local()`.

These are the Arrow files used by EJAM:

### Arrow file update groups

Arrow files do not all change on the same schedule. Use these groups when
planning updates:

1. **Facility Data Updates** include `frs`, `frs_by_programid`,
   `frs_by_naics`, `frs_by_sic`, and `frs_by_mact`. These may be refreshed
   when EPA FRS/facility data are updated.

2. **EJSCREEN Annual Data Update** currently includes `bgej.arrow`. It is
   calculated from the annual EJScreen/EJAM demographic and environmental
   pipeline and must match the installed package's `blockgroupstats`,
   `usastats`, and `statestats`.

3. **Blockgroup Geography Updates** include `bgid2fips` and `blockwts`, and
   related `.rda` objects such as `bgpts` and `bg_cenpop2020`. These need
   review during each annual update and regeneration when blockgroup FIPS,
   EJAM `bgid`, internal points, or blockgroup-to-block relationships change.

4. **Block Geography Updates** include `blockpoints`, `quaddata`, and
   `blockid2fips`. These need regeneration only when block-level FIPS or block
   internal-point geography changes.

For EJAM v2.5.0, the block and blockgroup helper files are intentionally
carried forward without Island Area blocks. They should not be used to promise
point-buffer/radius or block-weighted polygon analysis for AS/GU/MP/VI; those
analyses should return no-data results rather than block-weighted estimates.

Use `EJAM:::dynamic_geography_arrow_report()` to check whether the current
blockgroup and block geography Arrow files are compatible with the installed
`blockgroupstats` blockgroup universe.

### Blockgroup and block-level arrow files

-   `?bgid2fips`.arrow: crosswalk of EJAM blockgroup IDs (1-n) with 12-digit
blockgroup FIPS codes
-   `?blockid2fips`.arrow: crosswalk of EJAM block IDs (1-n) with 15-digit block
FIPS codes
-   `?blockpoints`.arrow: Census block internal points lat-lon coordinates, EJAM
block ID
-   `?blockwts`.arrow: Census block population weight as share of blockgroup
population, EJAM block and blockgroup ID
-   `?bgej`.arrow: blockgroup-level statistics of EJ variables. This is part
of the EJSCREEN Annual Data Update group and must match the package's
`blockgroupstats`
-   `?quaddata`.arrow: 3D spherical coordinates of Census block internal points,
with EJAM block ID

### FRS/facility-related arrow files

-   `?frs`.arrow: data.table of EPA Facility Registry Service (FRS) regulated
sites
-   `?frs_by_naics`.arrow: data.table of NAICS industry code(s) for each
EPA-regulated site in Facility Registry Service
-   `?frs_by_sic`.arrow: data.table of SIC industry code(s) for each
EPA-regulated site in Facility Registry Service
-   `?frs_by_programid`.arrow: data.table of Program System ID code(s) for each
EPA-regulated site in the Facility Registry Service
-   `?frs_by_mact`.arrow: data.table of [MACT
NESHAP](https://en.wikipedia.org/wiki/National_Emissions_Standards_for_Hazardous_Air_Pollutants)
codes for sites, indicating the [subpart(s) that categorize relevant
EPA-regulated
sites](https://www.epa.gov/stationary-sources-air-pollution/national-emission-standards-hazardous-air-pollutants-neshap-8)

## Development/Setup

1.  The Arrow files are stored as release assets in a separate public GitHub
repository (referred to here as `ejamdata`). The owner/repository name must be
recorded/updated in the `DESCRIPTION` field called `ejam_data_repo`, which can
be checked with `url_package(type = "data", get_full_url = TRUE)`. EJAM uses
that information to find the dynamic data files.

2.  Any time the Arrow datasets are updated, create or update an `ejamdata`
release and upload the `.arrow` files as release assets. Use the maintainer
helper described below rather than relying on an automatic GitHub Actions
workflow.

3.  EJAM's `download_latest_arrow_data()` function does the following:

<!-- -->

a.  Resolves the package-compatible `ejamdata` release tag from the
`DESCRIPTION` field `ejamdata_required_tag`, unless a maintainer explicitly
passes a different `piggybacktag`. This lets a patch release of EJAM keep using
an earlier compatible `ejamdata` release if the Arrow files have not changed.
b.  Checks the user's locally installed Arrow data release tag, which is stored
in `data/ejamdata_version.txt`.
c.  If the `data/ejamdata_version.txt` file doesn't exist, for example on the
first EJAM install, it will be created at the end of the script.
d.  If the versions are different, downloads Arrow files from the matching
`ejamdata` release with `piggyback::pb_download()`.
e.  When `dataload_dynamic("bgej")` loads `bgej`, the local `bgej.arrow` must
also match the installed package's `blockgroupstats`; if it does not, EJAM
tries to replace it from the package-compatible `ejamdata` release tag. See how
this function works for details:

```{r, eval=FALSE}
download_latest_arrow_data()
```

4.  EJAM calls this logic from the attach/startup path through
`dataload_dynamic()` so the needed Arrow files are available when a user loads
EJAM or starts the app.

## How it Works for the User

1.  User installs EJAM

<!-- -->

-   `pak::pkg_install("`r code_reponame_with_owner`")` (or as adjusted depending on the actual repository owner and name)

<!-- -->

2.  User loads EJAM as usual

<!-- -->

-   `library(EJAM)`. This triggers the dynamic-data checks needed for startup.

<!-- -->

3.  User runs EJAM as usual

<!-- -->

-   The `dataload_dynamic()` function will work as usual because the needed
`.arrow` files are cached locally after they are downloaded.

## How New Versions of Arrow Datasets Are Republished / Released

First, create the key Arrow files locally or from the relevant pipeline output,
as explained above.

For future updates, the package may be modified to publish these files
via the update pipeline or related script such as `run_arrow_publish_v2.5.0.R`
but the information below is to describe the functions if an update is done manually.

As mentioned above, we use [the piggyback package](https://docs.ropensci.org/piggyback/index.html) to place large datasets in the assets of a new release on the `r data_reponame_url` repository, rather than committing them with Git. The current maintainer path is to call `datasets_arrow_publish()` with explicit local `.arrow` file paths.

The helper is intentionally conservative. It defaults to `dry_run = TRUE`,
`overwrite = FALSE`, and `mark_latest = FALSE`. The default release note is
`"Updated datasets for EJScreen/EJAM updated as of "` plus the `release_date`
parameter.

Make sure the intended new data objects are available as `.arrow` files. For an
annual EJSCREEN data release, `bgej.arrow` is the critical package-coupled asset
and must match the package version/release tag. Facility and geography Arrow
files may be carried forward unchanged if they are still compatible. If block or
blockgroup helper Arrow files such as `blockwts.arrow`, `blockpoints.arrow`,
`blockid2fips.arrow`, `bgid2fips.arrow`, or `quaddata.arrow` are intentionally
regenerated in a future geography update, publish those files with the same
helper after a dry-run review. For v2.5.0, Island Areas are handled only at the
blockgroup dataset/export/map-data level; Island Area blocks are not added to
the block-helper universe for this release path.

Example dry-run for a manual publish:
 (also see run_arrow_publish_v2.5.0.R to publish all .arrow files in 1 step)

```{r, eval=FALSE}
release_number <- EJAM:::ejamdata_required_tag()
new_datasets_folder <- "~/Downloads"
filepaths_arrow <- file.path(new_datasets_folder, "bgej.arrow")

EJAM:::datasets_arrow_publish(
  files = filepaths_arrow,
  tag = release_number,
  release_date = Sys.Date(),
  dry_run = TRUE,
  overwrite = FALSE,
  mark_latest = FALSE
)
```

After reviewing the dry-run output and the intended release tag, rerun with
`dry_run = FALSE` only when ready to create/update the release assets. Use
`overwrite = TRUE` only after confirming existing assets with the same names
should be replaced. Use `mark_latest = TRUE` only when this release should be
shown by GitHub as the latest release.

Open a browser to confirm they are there.
```{r, eval=FALSE}
browseURL(paste0(EJAM:::url_package("data", get_full_url = T), "/releases"))
```

Reload EJAM so it can get the updates. It should detect that new versions are
available and cache them for the installed package.

```{r, eval=FALSE}
rm(list=ls())
require(EJAM)

# Confirm they all can be opened as Arrow-backed objects
# as arrow files:
dataload_dynamic("all", return_data_table = FALSE)
# or read into memory as data.table/data.frame objects:
dataload_dynamic("all", return_data_table = TRUE)
```

This previously had been handled with a GitHub Actions workflow that tried to
use Git LFS. That automatic workflow is no longer used and should not be
restored.


## Potential Improvements

### Making More of the Code More Arrow-Friendly

Problem: loading the data as tibbles/data frames takes a long time.

Solution: We may be able to modify more of our code to be more Arrow-friendly. This
essentially keeps the analysis code as a sort of query, and only actually loads
the results into memory when requested (e.g., via `dplyr::collect()`). This
dramatically reduces memory usage, which would speed up processing times and
avoid potential crashes resulting from insufficient memory. However, this would
require a decent lift to update the code in all places.

Pros: processing efficiency and significantly reduced memory usage.

Implementation: This has been enabled by the `dataload_dynamic()`
function, which contains a `return_data_table` parameter. If `FALSE`, the Arrow
file is opened as an Arrow-backed object rather than read fully into a
data.table/data.frame.
