---
title: "Testing EJAM App with shinytest2"
description: "How to run {shinytest2} suite for app-based functionality testing."
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Testing EJAM App with shinytest2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

This document covers the UI-related automated tests.

## Dev Environment

If you are successfully running the app, you should have all the necessary packages, or at least those in Imports or Suggests of the `DESCRIPTION` file -- the diffviewer package is not there, e.g. Some dev-related packages to note:

-   **`{shinytest2}`** - [shinytest2](https://rstudio.github.io/shinytest2/) is the key R package helping to test shiny web app functionality 
-   **`{diffviewer}`** - [diffviewer](https://diffviewer.r-lib.org) helps visually compare 2 files

-   **Pandoc** and **knitr** – The Pandoc software comes bundled with RStudio, and is a "swiss-army knife" for document conversion. There also is a function called [knitr::pandoc()] that is a wrapper that calls Pandoc to convert docs to HTML, PDF, etc.
-   **PhantomJS** maybe – This used to be needed for downloads (and was installed with `webshot::install_phantomjs()`). But note this newer info:

NOTE FROM https://rstudio.github.io/shinytest2/articles/z-migration.html: 
-  `{shinytest2}` is the successor to `{shinytest}`. `{shinytest}` was implemented using `{webdriver}` which uses [PhantomJS](https://phantomjs.org/api/). PhantomJS has been unsupported since 2017 and does not support displaying `{bslib}`'s Bootstrap v5. `{shinytest2}` uses `{chromote}` to connect to your locally installed Chrome or Chromium application, allowing `{shinytest2}` to display `{bslib}`'s Bootstrap v5.


## How shinytest2 Works

`{shinytest2}` (here referred to henceforth as "shinytest" not to be confused with the older R package that was named shinytest!) automates shiny web app functionality testing, so we can determine if code updates break or unexpectedly modify parts of an application.

It runs the installed version of the app in a headless Chromium browser, simulating user interactions and taking snapshots. These snapshots are stored as JSON files with accompanying PNG images. If differences arise from code updates, the test fails, indicating which files changed.

### Key Features:

-   Compares `.json`, `.html`, `.xlsx` files to a baseline
-   Snapshots include inputs, outputs, and exported values
-   `.png` files provide visual confirmation (but do not cause test failures)
-   Developers can update snapshots to set a new baseline

## EJAM's shinytest2 Folder Structure

``` plaintext
tests/
  ├── testthat.R (modified to launch the shinytest2 web app functionality tests)
  └── testthat/
      ├── setup-shinytest2.R
      ├── test-webapp-[DATA TYPE]-functionality.R (e.g. test-webapp-FIPS-functionality.R)
      └── _snaps/
          ├── [OS, e.g. linux]-[R Version, e.g. 4.5]/
          │   ├── FIPS-functionality/
          │   │   ├── .json, .png, .xlsx, .html files
          │   ├── latlon-functionality/
          │   │   ├── .json, .png, .xlsx, .html files
          │   ├── NAICS-functionality/
          │   │   ├── .json, .png, .xlsx, .html files
```

### File Descriptions

-   **`testthat.R`** – Calls `shinytest2::test_app()` to run all tests. Can filter specific tests using `filter="test-name"`. This file is work in progress -- it may need more work for the shiny app testing to work correctly using the tests in the `test-ui_and_server.R` file.
-   **`testthat/setup-shinytest2.R`** – Loads `global.R` and app scripts into the testing environment.  Does basic setup including defining shinytest2_webapp_functionality() which has a script of web app UI interactions to test, using the app to upload or select, and analyze, multiple data types (FIPS, shapefile, latlon, NAICS, etc.), and save results like reports or spreadsheets.
-   **`testthat/test-webapp-[DATA TYPE]-functionality.R`** – Simple call to the main app functionality function, specifying the data type to test with.
-   **`testthat/_snaps/`** – Stores snapshots categorized by OS, R version, and data type.
    -   **`.json` files** – Capture app snapshots.
    -   **`.png` files** – Screenshots (do not trigger failures).
    -   **`.xlsx` & `.html` files** – Download files. Compared via content hashing to prevent false failures.

## Updating Tests

You may wish to modify the shinytest scripts, either to add new interactions with the application or to modify existing ones, such as in the case of an app component that can no longer be interacted with. Here are some methods and tips for updating the shinytest script accordingly.

### Direct Updates

Modify `shinytest2_webapp_functionality()` in setup-shinytest2.R to add new interactions with the app for the shinytest to test. You would update that file by coding new UI interactions directly.

### Using `shinytest2::record_test()` to generate testing code

If you're not sure how to code new UI interactions directly, run `shinytest2::record_test()` to test the app interactively and record your actions, which can then be copied into test scripts.

### Using `shiny::exportTestValues(name = value)`

Throughout the app code, `shiny::exportTestValues()` can be used to store values from reactive expressions or other _items that are not inputs or outputs_ and therefore may not be included in the standard snapshots. Then, in the shinytests, you can specify `export=[name]` to include in the snapshot the export named "name" that you specified in the code, or `export=TRUE` to include all exports.

## Running Tests Locally

The primary method for running the shinytests (other than doing so via GitHub Actions) is:

``` r
shinytest2::test_app(".", filter="-functionality")
```

However, it is strongly recommended during development to source() the entire `testthat.R` which runs `remotes::install_local()` to ensure your development code is the one tested. This is because shinytest2 automatically references the installed version of a package.


## GitHub Actions Integration

Using GitHub Actions (GHA) we can have GitHub run our shiny web app UI tests prior to merging a Pull Request, to give us peace of mind that the app will still work with the merged code.

### Workflow

-   The GHA sets up R, installs dependencies, runs tests, and compares snapshots.
-   The workflow is stored in `.github/workflows/test-webapp-functionality.yml`.
-   PRs to a specified branch such as  `development`, `main` can trigger GHA workflows (as specified in the workflow yml file).

### Speed Optimization

-   If GHA takes too long, cache dependencies by temporarily disabling steps after setup.
-   If snapshots fail, merge the base branch into the feature branch before updating snapshots.

## Reviewing & Updating Snapshots (Saved Results of Testing)

### Reviewing Snapshots 

```{r, eval = FALSE}
testthat::snapshot_review()
```

Optionally, can filter to review specific files or folders of snapshots.

### Accepting New Snapshots

```{r, eval=FALSE}
testthat::snapshot_accept()
```

Optionally, can accept them interactively when reviewing.

## Debugging Tests & GitHub Actions

### Debugging shinytest2

-   Use `save_log()` to inspect logs.
-   Add `print()`, `message()`, or `warning()` statements in shinytest2_webapp_functionality().
-   Run, line-by line or in chunks, the main shinytest code in shinytest2_webapp_functionality()

Then, after running lines or chunks, run `app$get_log()` to view the log.

### Debugging GHA

Generally, if you test locally and update snapshots accordingly, GHA should pass. However, the tests do sometimes fail due to OS differences, R version differences, or even package differences. Here are some tips for debugging these issues:

-   Inspect the log in the GitHub repo, under the Actions tab or the Checks tab of the PR.
-   Inspect artifacts (zipped test outputs) after a failed run and compare snapshots in a diff viewer to identify discrepancies.

## Current State of Tests

-   If the shinytests are failing, it is likely because snapshots have not been updated locally and pushed after recent changes.
-   The versions of R and shiny and shinytest2 used for testing also should be noted as potentially affecting tests.

## Resources

[RStudio shinytest2 Documentation](https://rstudio.github.io/shinytest2/reference/AppDriver.html#method-AppDriver-expect_download)
