---
title: "EJSCREEN Documentation"
description: "EJSCREEN Documentation"
output:
  rmarkdown::html_vignette:
  toc: true
vignette: >
  %\VignetteIndexEntry{EJSCREEN Documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

```{r note, eval=FALSE, echo=FALSE}
# Some of this info was previously at
# http://htmlpreview.github.io/?https://github.com/ejanalysis/EJAM/blob/development/data-raw/EJSCREEN_archived_pages/EJSCREEN_archived_pages_and_docs.html
```

EJSCREEN originally provided extensive documentation and help through the
EJSCREEN website (e.g., overviews and how to interpret reports) and documents
available for download, such as the User Guide (how to use the app) and the
Technical Documentation (how indicators are calculated).

In 2025, before EPA took down the EJSCREEN website and documents, they were
archived by the Wayback Machine. The webpages and documents on archive.org load
slowly, so copies of some key documents are also made available here, for easy
access.

Below are links to archived copies of selected publicly-available EJSCREEN
webpages and documents from the U.S. Environmental Protection Agency (EPA) as of
2025, when the EPA-hosted open source EJAM project was archived by the [Wayback
Machine](https://web.archive.org) (an initiative of the Internet Archive) and by
EJAM-related repositories.

# Selected Documents

## Help with How to Use EJSCREEN

-   EJSCREEN User Guides (.pdf docs)

    -   [EJSCREEN Overall User
        Guide](https://github.com/ejanalysis/EJAM/blob/development/data-raw/user-guides/EJSCREEN-USER-GUIDE-2025-02.pdf){target="_blank"}
        (aka EJSCREEN walkthrough). Also [at
        archive.org](https://web.archive.org/web/20250203174905/https://ejscreen.epa.gov/mapper/help/ejscreen_help.pdf){target="_blank"}

    -   [EJSCREEN Multisite Tool User
        Guide](https://ejam.policyinnovation.info/www/user-guide-2025-02.pdf){target="_blank"}
        (aka EJAM walkthrough). Also [at
        archive.org](https://github.com/ejanalysis/EJAM/blob/main/inst/app/www/user-guide-2025-02.pdf){target="_blank"}

## Documentation of Data Sources and Technical Methods

-   Changes in 2025

    -   EJSCREEN / EJAM [developments in
        2025](https://ejanalysis.org/status){target="_blank"}

-   Data Sources (2024 through 2025)

    -   [**Full list of indicators, with
        definitions**](ejscreen-map-descriptions.html){target="_blank"}.
        Also [at
        archive.org](https://web.archive.org/web/20250123161322/https://www.epa.gov/ejscreen/ejscreen-map-descriptions){target="_blank"}

    -   [**Key environmental indicators with year and source of
        each**](http://htmlpreview.github.io/?https://github.com/ejanalysis/EJAM/blob/development/data-raw/EJSCREEN_archived_pages/overview-environmental-indicators-ejscreen.html){target="_blank"}.
        Also [at
        archive.org](https://web.archive.org/web/20250124214227/https://www.epa.gov/ejscreen/overview-environmental-indicators-ejscreen){target="_blank"}

    -   [Key socioeconomic
        indicators](http://htmlpreview.github.io/?https://github.com/ejanalysis/EJAM/blob/development/data-raw/EJSCREEN_archived_pages/overview-socioeconomic-indicators-ejscreen.html){target="_blank"}.
        Also [at
        archive.org](https://web.archive.org/web/20250102180705/https://www.epa.gov/ejscreen/overview-socioeconomic-indicators-ejscreen){target="_blank"}

    -   [EJSCREEN Change Log (newest indicators,
        etc.)](https://web.archive.org/web/20250124090116/https://www.epa.gov/ejscreen/ejscreen-change-log){target="_blank"}

    -   [Old API-provided list of
        indicators](https://web.archive.org/web/20250131005731/https://ejscreen.epa.gov/mapper/ejsoefielddesc1.html){target="_blank"}

    -   Raw Data Downloads -- EJSCREEN datasets as used by EJAM

        -   Datasets stored in the EJAM repository (.rda files):
            [Documentation](../reference/index.html#datasets-with-indicators-raw-data-means-percentiles-){target="_blank"}
            and [Data
            files](https://github.com/ejanalysis/EJAM/tree/main/data){target="_blank"}
        -   Datasets stored in the ejamdata repo (large .arrow files):
            [Documentation](../reference/index.html#datasets-with-indicators-raw-data-means-percentiles-){target="_blank"}
            and [Data
            files](https://github.com/ejanalysis/ejamdata/tree/main/data){target="_blank"}

-   Methods (development of indicators, calculations for community reports,
    etc.)

    -   [**EJSCREEN Technical Doc. (as saved for
        EJAM)**](`r EJAM::url_ejscreentechdoc()`){target="_blank"}:
        Detailed documentation of EJSCREEN methods (2024) for Version 2.3 (7/31/2024), .pdf manual (also 
        [at archive.org](https://web.archive.org/web/20250929044644/https://www.epa.gov/system/files/documents/2024-07/ejscreen-tech-doc-version-2-3.pdf){target="_blank"})

    ```{r notes, eval=FALSE, echo=FALSE, include=FALSE}
    # (`r EJAM::url_ejscreentechdoc()`)
    # (https://github.com/ejanalysis/EJAM/blob/development/data-raw/EJSCREEN_archived_pages/ejscreen-tech-doc-version-2-3.pdf)
    ```

    -   Documentation of EJAM software that generates EJSCREEN reports
        (from 2025 on):
        [ejanalysis.org/ejamdocs](https://ejanalysis.org/ejamdocs){target="_blank"}

--------------------------------------------------------------------------------

# *Original EJScreen Website Homepage*

## **EJScreen: Environmental Justice Screening and Mapping Tool**

EJScreen is EPA's environmental justice mapping and screening tool that provides
EPA with a nationally consistent dataset and approach for combining
environmental and socioeconomic indicators.

EPA is committed to protecting the environment and public health in low-income,
tribal and other overburdened communities by integrating environmental justice
in all programs, policies and activities.

[Launch the EJScreen Tool](https://ejanalysis.com/ejscreenapp){target="_blank"}

```{r echo = FALSE, results = 'asis'}
links_table <- cbind(
    aaa = c(
        "**Tool Basics and Uses**",
        "Explore these helpful resources to gain a solid understanding of EJScreen’s concepts and uses.",
        "[Explore concepts and uses](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/tool-basics-and-uses)"
    ),
    bbb = c(
        "**Learn to Use EJScreen**",
        "Explore these introductory videos, guides, and training sessions.",
        "[Learn about the tool](https://web.archive.org/web/20250121194750/https://www.epa.gov/ejscreen/learn-use-ejscreen)"
    ),
    ccc = c(
        "**Technical Information and Data Downloads**",
        "Download data and find detailed information about the technical aspects of EJScreen.",
        "[Dive into details](http://htmlpreview.github.io/?https://github.com/ejanalysis/EJAM/blob/development/data-raw/EJSCREEN_archived_pages/technical-information-and-data-downloads.html)"
    )
)
library(knitr)
kable(links_table,  col.names = NULL, label = NA)
```


# *Original EJScreen Website Outline*

**EJScreen: Environmental Justice Screening and Mapping Tool**

-   [Launch the EJScreen
    Tool](https://ejanalysis.com/ejscreenapp){target="_blank"}
-   [**Tool Basics and
    Uses**](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/tool-basics-and-uses){target="_blank"}
    -   **General**
        -   [What is
            EJScreen?](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/what-ejscreen){target="_blank"}
        -   [EJ and Supplemental
            Indexes](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/ej-and-supplemental-indexes-ejscreen){target="_blank"}
        -   [Environmental
            Indicators](http://htmlpreview.github.io/?https://github.com/ejanalysis/EJAM/blob/development/data-raw/EJSCREEN_archived_pages/overview-environmental-indicators-ejscreen.html){target="_blank"}
        -   [Socioeconomic
            Indicators](http://htmlpreview.github.io/?https://github.com/ejanalysis/EJAM/blob/development/data-raw/EJSCREEN_archived_pages/overview-socioeconomic-indicators-ejscreen.html){target="_blank"}
        -   [How was EJScreen
            Developed?](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/how-was-ejscreen-developed){target="_blank"}
        -   [Limitations and
            Caveats](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/limitations-and-caveats-using-ejscreen){target="_blank"}
    -   **Uses**
        -   [Purposes and Uses for
            EJScreen](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/purposes-and-uses-ejscreen){target="_blank"}
        -   [How Does EPA Use
            EJScreen?](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/how-does-epa-use-ejscreen){target="_blank"}
    -   **EJScreen Resources**
        -   [EJScreen Map
            Descriptions](ejscreen-map-descriptions.html){target="_blank"}
            [archived
            version](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/ejscreen-map-descriptions){target="_blank"}
        -   [Frequent Questions about
            EJScreen](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/frequent-questions-about-ejscreen){target="_blank"}
        -   [Other EJ Related Mapping
            Tools](https://web.archive.org/web/20250121194840/https://www.epa.gov/ejscreen/related-tools){target="_blank"}
        -   [EJScreen Fact Sheet
            (pdf)](https://web.archive.org/web/20250121194840/https://www.epa.gov/system/files/documents/2023-06/ejscreen-fact-sheet.pdf){target="_blank"}
-   [**Learn to Use
    EJScreen**](https://web.archive.org/web/20250121194750/https://www.epa.gov/ejscreen/learn-use-ejscreen){target="_blank"}
    -   [EJScreen Office Hours and Trainings (note: this is
        obsolete)](https://web.archive.org/web/20250121194750/https://www.epa.gov/ejscreen/ejscreen-office-hours-training){target="_blank"}
    -   [How to Interpret EJScreen
        Data](https://web.archive.org/web/20250121194816/https://www.epa.gov/ejscreen/how-interpret-ejscreen-data){target="_blank"}
    -   [Example of an EJScreen Community
        Report](https://web.archive.org/web/20250121194750/https://www.epa.gov/ejscreen/example-ejscreen-community-report){target="_blank"}
    -   [Threshold
        Maps](https://web.archive.org/web/20250121194750/https://www.epa.gov/ejscreen/threshold-maps-ejscreen){target="_blank"}
-   [**Technical Information and Data
    Downloads**](http://htmlpreview.github.io/?https://github.com/ejanalysis/EJAM/blob/development/data-raw/EJSCREEN_archived_pages/technical-information-and-data-downloads.html){target="_blank"}
    or [as
    archived](https://web.archive.org/web/20250121194640/https://www.epa.gov/ejscreen/technical-information-and-data-downloads){target="_blank"}
    -   [Download EJScreen
        Data](https://web.archive.org/web/20250121194640/https://www.epa.gov/ejscreen/download-ejscreen-data){target="_blank"}
    -   [EJScreen
        API](https://web.archive.org/web/20250121194640/https://www.epa.gov/ejscreen/ejscreen-api){target="_blank"}
    -   [Updates and
        Changes](https://web.archive.org/web/20250121194640/https://www.epa.gov/ejscreen/ejscreen-updates-and-changes){target="_blank"}
    -   [**EJSCREEN Technical Doc. (as saved for
        EJAM)**](`r EJAM::url_ejscreentechdoc()`){target="_blank"}:
        Detailed documentation of EJSCREEN methods (2024) - EJScreen Technical
        Documentation for Version 2.3 from 7/31/2024, pdf-format manual (also
        archived [at
        archive.org](https://web.archive.org/web/20250929044644/https://www.epa.gov/system/files/documents/2024-07/ejscreen-tech-doc-version-2-3.pdf){target="_blank"})
-   [Contact Us About EJScreen (note: this is
    obsolete)](https://web.archive.org/web/20250121194732/https://www.epa.gov/ejscreen/forms/contact-us-about-ejscreen){target="_blank"}

--------------------------------------------------------------------------------
