---
title: "Installing the EJAM R package"
description: "1. Installing the EJAM R package"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Installing the EJAM R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 80
---

```{r developernote, eval=FALSE, echo= FALSE, include = FALSE}
#  *>>>>>>>>>> Developer note: vignettes need to be tested/edited/rebuilt regularly <<<<<<<<<<<*
#    - **See ?pkgdown::build_site** and script in EJAM/data-raw/- EJAM uses the pkgdown R package to build help and articles/ vignettes as web pages
```

```{r SETUP_default_eval_or_not, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(eval = FALSE)
# https://r-pkgs.org/vignettes.html
```

```{r libraryEJAM, eval = FALSE, echo= FALSE, include= FALSE}
# rm(list = ls()); golem::detach_all_attached(); devtools::load_all()
 ## DONT NEED THESE FOR THIS VIGNETTE ON INSTALLATION STEPS
#if (!exists("blockgroupstats")) {library(EJAM)} # use installed version only if pkg not yet attached
#dataload_dynamic('all') # varnames = all  currently means all defined by .arrow_ds_names
#indexblocks()
```

```{r, eval=TRUE, include=FALSE}
owner_repo <- EJAM:::repo_from_desc()      # had been "USEPA/EJAM-open"
repo <- gsub(".*/", "", owner_repo)        # had been       "EJAM-open"
quoted_owner_repo <- paste0("'", owner_repo, "'")
```

EJAM is not only a web app, it is also an R package you can install. This is useful if you want to use the full set of EJAM R functions, many of which do things not available in the web app, to do customized analysis, to explore the datasets, or to reuse data or code in other applications.

This assumes you have installed and know how to use [R and RStudio](https://posit.co/download/rstudio-desktop/) (some functions assume you have RStudio), and [git](https://github.com/git-guides/install-git){.uri target="_blank" rel="noreferrer noopener"}. You might also want [GitHub Desktop](https://github.com/apps/desktop){.uri target="_blank" rel="noreferrer noopener"} for convenience.

## Options

The package is on GitHub.com, not CRAN (so `install.packages('EJAM')` would _not_ work).

To use EJAM in RStudio, you have two options:

  1. You can clone (or fork) the repo as a new RStudio Project, and then install from that source. Cloning (or forking) is easy and allows you to explore or contribute to source code. This is the recommended option. 
  2. Option 2 is to avoid cloning. That is easier if you only want to run the web app locally, but may not work if you want to use certain R functions directly.

## Option 1. Install _with_ cloning (preferred approach)

### Step 1. Clone EJAM & create a Project in RStudio

Here are three ways, but any of them should work -- The first one is probably the simplest:

method A. start in **RStudio**: 
Click **File**, **New Project**, **Version Control**, **Git**, and for the repository URL, enter ``r paste0("https://github.com/", owner_repo)``. For help: [How to create a new RStudio project by cloning a remote Git repo](https://docs.posit.co/ide/user/ide/guide/tools/version-control.html#creating-a-new-project-based-on-a-remote-git-or-subversion-repository){.uri target="_blank" rel="noreferrer noopener"}. Then add it to GitHub Desktop if that is a tool you use.

method B. start in **GitHub Desktop**: 
Click **File**, **Clone**, **URL**, and for the repository URL, enter ``r paste0("https://github.com/", owner_repo)``. For help: [How to clone a repo using GitHub Desktop](https://docs.github.com/en/desktop/adding-and-cloning-repositories/cloning-and-forking-repositories-from-github-desktop){.uri target="_blank" rel="noreferrer noopener"}. Then in **RStudio**: Click **File**, **New Project**, **Existing Directory**, and specify the folder you just cloned into.

method C. start at **GitHub.com**: 
Go to [the EJAM repository page](https://github.com/`r owner_repo` `r paste0("'github.com/", owner_repo, "'")`){.uri target="_blank" rel="noreferrer noopener"}, click the green **Code** button, **Download Zip**, then unzip the zip file that contains the package. For help: [How to clone a repo from GitHub.com](https://docs.github.com/en/repositories/creating-and-managing-repositories/cloning-a-repository){.uri target="_blank" rel="noreferrer noopener"}. Then in **RStudio**: Click **File**, **New Project**, **Existing Directory**, and specify the folder you just cloned into.

#### Configure Build Tools in RStudio (optional)

You can click Build, Configure Build Tools, and change settings. For example, try these settings:

- Project build tools: Package
- Package directory: (Project Root)
- Always use --preclean : checked (yes)
- Use devtools : checked (yes)
- Generate documentation with Roxygen: checked (yes)
  - Configure Roxygen options: all checked/yes, except for Vignettes and Build/Install
- Install package R CMD check: --no-multiarch --with-keep.source
- Build source & Build binary: leave both blank

For help: [How to configure build tools in RStudio](https://docs.posit.co/ide/user/ide/guide/pkg-devel/writing-packages.html#configuring-build-tools){.uri target="_blank" rel="noreferrer noopener"}.

------------------------------------------------------------------------

### Step 2. Install (after Cloning)

You can use the Build menu in RStudio, or just run the following (which installs all the Suggests, not just Imports)
```{r install-post-clone}
options(timeout=300)
if (!require(devtools)) {install.packages("devtools")}
if (!(basename(getwd()) == "EJAM" && file.exists("DESCRIPTION"))) {
  stop("must be in root folder of EJAM source package")}
devtools::install(dependencies=T, upgrade="always", build=F)
```


## Option 2. Install _without_ cloning (not the preferred approach)

You may be able to install the package to just run the web app locally in RStudio without cloning, but some of the code assumes a local source code copy is available (especially functions or scripts related to analysis the web app cannot do, or for package development).

### Using install_github()

If you *do have a PAT set up*, you can use `remotes::install_github()`. (For help on PAT setup, see the last section of this article).

```{r install-gh, eval=FALSE}
options(timeout=300); if (!require(remotes)) {install.packages("remotes")}
ref = github_release() # specifies you want the latest released version
#ref = "development"   # development branch version
#ref = "HEAD"          # main branch
#ref = "EJAM-v2.32.6"  # a specific branch
remotes::install_github(
  repo = "ejanalysis/EJAM", ref = ref, 
  dependencies=T, upgrade="always", build=F)
library(EJAM)
```

### Using install_url() or install_local()

If you *don't have a [personal access token](https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/managing-your-personal-access-tokens#about-personal-access-tokens){.uri target="_blank" rel="noreferrer noopener"} (PAT) and don't want to set one up*, you could [check the github repository for the latest release](`r paste0("https://github.com/", EJAM:::repo_from_desc(), "/releases/latest")`) and may be able to use `remotes::install_url()` something like this:

```{r install-url, eval=FALSE}
options(timeout=300); if (!require(remotes)) {install.packages("remotes")}
zipname = "v2.32.6.zip" # or "v2.32.5.zip", etc. - update as needed
remote_zip = paste0(
  "https://github.com/ejanalysis/EJAM/archive/refs/tags/", zipname)
remotes::install_url(url = remote_zip, 
                     dependencies=T, upgrade='always', build=F)
library(EJAM)
```

_or download the .zip file as a separate step, and install from that saved file:_

```{r install-local, eval=FALSE}
options(timeout=300); if (!require(remotes)) {install.packages("remotes")}
zipname = "v2.32.6.zip" # or "v2.32.5.zip", etc. - update as needed
remote_zip = paste0(
  "https://github.com/ejanalysis/EJAM/archive/refs/tags/", zipname)
local_zip = file.path(tempdir(), zipname)
download.file(remote_zip, destfile = local_zip)
remotes::install_local(path = local_zip, 
                       dependencies=T, upgrade="always", build=F)
library(EJAM)
```

```{r install-package-notes, eval=FALSE, echo=FALSE, include=FALSE}
## note the repo name would be owner-repo obtained near top of .Rmd via EJAM:::repo_from_desc()
```


------------------------------------------------------------------------

## Start Using EJAM

After installing, see the [Basics - Quick Start Guide](basics.html).

------------------------------------------------------------------------

------------------------------------------------------------------------

## Other Technical Details

Just in case you need more details on how installing and attaching the package works, the following describes the package dependencies and code used by the package to get the data and build an index.

### Setting up Personal Access Tokens (PATs)

You may need to set up a [personal access token](https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/managing-your-personal-access-tokens#about-personal-access-tokens){.uri target="_blank" rel="noreferrer noopener"} (PAT) for authentication to work when using a repository on GitHub.com. See more [about git credentials and the credential store](https://usethis.r-lib.org/articles/git-credentials.html#git-credential-helpers-and-the-credential-store). Note Windows takes care of most of this now, in conjunction with GitHub.
```{r PAT, echo=TRUE, eval=FALSE}
##  To check for existing PATs:
usethis::gh_token_help() # or
usethis::git_sitrep() # git situation report

#  To make a new PAT:
usethis::create_github_token()

#  To register a PAT:
credentials::set_github_pat()
```

### Details on CRAN packages needed (dependencies)

You should not have to do anything other than the instructions above, to handle package dependencies. EJAM needs dozens of other packages to be installed that are (almost all) available from [CRAN](https://cran.r-project.org). Installing the EJAM package as explained above (with dependencies=TRUE, upgrade="always") will handle obtaining those other packages. Cloning and building/installing and then trying to load/attach EJAM will also alert you to those other packages you need to install if you don't already have them. In case it is of interest, a list of packages needed is in the `DESCRIPTION` file in the R package source code root folder (as can be found in the code repository). Note some are in Suggests and you probably want to install those as well since some of them were actually used in some less-often-used features or functions. Using dependencies=T in `remotes::install_github()`, `remotes::install_url()`, install(), etc. will make sure they are all installed.

### Details on the automatic data downloads

To work in the RStudio console, EJAM needs some datasets not stored as part of the package. However, they already should be downloaded (right after you first install the package) and loaded into memory automatically (or be ready for lazy-loading or otherwise loading as needed) as soon as you do require(EJAM) or library(EJAM).

On first use, the package should automatically download some data files from a related repository. Each time the package is attached via library() or require(), `.onAttach()` will check for updates and also will build a spatial index of Census block points called `?localtree`, via `indexblocks()`.

Typically you would not need to download any datasets yourself, because EJAM just downloads these when the app starts (technically, when the R package is attached) (or only as needed in the case of certain datasets that are not always needed). Some datasets are installed along with the package, such as the [blockgroupstats](../reference/blockgroupstats.html) data. But large files like [blockpoints](../reference/blockpoints.html) are stored in a separate data repo, and EJAM downloads them from there. You might want your own local copies, though, for these reasons:
 
Attaching the package actually checks (using internal function `dataload_dynamic()`) for copies in memory first (e.g., `exists("quaddata", envir = globalenv())`), then local disk (using `dataload_from_local()` looking in the data folder of the (source or installed) package, as defined by `app_sys()` which is just a wrapper for `system.file()`), then finally tries to download any still needed, using internal functions.
